### 2024-05-01 _tex/index.tex
Built site for gh-pages

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{agujournal2019}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={La Palma Earthquakes},
  pdfauthor={Steve Purves; Rowan Cockett},
  pdfkeywords={La Palma, Earthquakes},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\journalname{Earth and Space Science}

\draftfalse

\begin{document}
\title{La Palma Earthquakes}

\authors{Steve Purves\affil{1}, Rowan Cockett\affil{1}}
\affiliation{1}{Curvenote, }
\correspondingauthor{Steve Purves}{steve@curvenote.com}

\begin{abstract}
In September 2021, a significant jump in seismic activity on the island
of La Palma (Canary Islands, Spain) signaled the start of a volcanic
crisis that still continues at the time of writing. Earthquake data is
continually collected and published by the Instituto Geográphico
Nacional (IGN). \ldots{}
\end{abstract}

\section*{Plain Language Summary}
Earthquake data for the island of La Palma from the September 2021
eruption is found \ldots{}

\section{Introduction}\label{introduction}

\textsubscript{Source:
\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-timeline}
\begin{figure}[H]

\centering{

\includegraphics{index_files/figure-pdf/fig-timeline-1.pdf}

}

\caption{\label{fig-timeline}Timeline of recent earthquakes on La Palma}

\end{figure}%

\textsubscript{Source:
\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
Notebook}}

\textsubscript{Source:
\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
Notebook}}

Based on data up to and including 1971, eruptions on La Palma happen
every 79.8 years on average.

Studies of the magma systems feeding the volcano, such as Marrero et al.
(2019), have proposed that there are two main magma reservoirs feeding
the Cumbre Vieja volcano; one in the mantle (30-40km depth) which
charges and in turn feeds a shallower crustal reservoir (10-20km depth).

Eight eruptions have been recorded since the late 1400s
(Figure~\ref{fig-timeline}).

Data and methods are discussed in Section~\ref{sec-data-methods}.

Let \(x\) denote the number of eruptions in a year. Then, \(x\) can be
modeled by a Poisson distribution

\begin{equation}\phantomsection\label{eq-poisson}{
p(x) = \frac{e^{-\lambda} \lambda^{x}}{x !}
}\end{equation}

where \(\lambda\) is the rate of eruptions per year. Using
Equation~\ref{eq-poisson}, the probability of an eruption in the next
\(t\) years can be calculated.

\begin{longtable}[]{@{}ll@{}}
\caption{Recent historic eruptions on La
Palma}\label{tbl-history}\tabularnewline
\toprule\noalign{}
Name & Year \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
Name & Year \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Current & 2021 \\
Teneguía & 1971 \\
Nambroque & 1949 \\
El Charco & 1712 \\
Volcán San Antonio & 1677 \\
Volcán San Martin & 1646 \\
Tajuya near El Paso & 1585 \\
Montaña Quemada & 1492 \\
\end{longtable}

Table~\ref{tbl-history} summarises the eruptions recorded since the
colonization of the islands by Europeans in the late 1400s.

\begin{figure}

\centering{

\includegraphics{images/la-palma-map.png}

}

\caption{\label{fig-map}Map of La Palma}

\end{figure}%

La Palma is one of the west most islands in the Volcanic Archipelago of
the Canary Islands (Figure~\ref{fig-map}).

\begin{figure}[H]

\centering{

\includegraphics{index_files/figure-latex/notebooks-explore-earthquakes-fig-spatial-plot-output-1.png}

}

\caption{\label{fig-spatial-plot}Locations of earthquakes on La Palma
since 2017}

\end{figure}%

\textsubscript{Source:
\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/explore-earthquakes-preview.html\#cell-fig-spatial-plot}{Explore
Earthquakes}}

Figure~\ref{fig-spatial-plot} shows the location of recent Earthquakes
on La Palma.

\section{Data \& Methods}\label{sec-data-methods}

\section{Conclusion}\label{conclusion}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\vspace{1em}

\bibitem[\citeproctext]{ref-marrero2019}
Marrero, J., García, A., Berrocoso, M., Llinares, Á., Rodríguez-Losada,
A., \& Ortiz, R. (2019). Strategies for the development of volcanic
hazard maps in monogenetic volcanic fields: The example of {La} {Palma}
({Canary} {Islands}). \emph{Journal of Applied Volcanology}, \emph{8}.
\url{https://doi.org/10.1186/s13617-019-0085-5}

\end{CSLReferences}

\end{document}

### 2024-05-24 index.qmd
updated figures, added methods

---
title: "ParaChiCSVO2"
titlerunning: "CSVO2"
author:
  - name: Thomas Gavin Carmichael
    orcid: 0009-0008-6849-5333
    corresponding: false
    email: tgcarmichael@outlook.com
    roles:
      - writing - original draft
      - formal analysis
      - methodology
      - validation
      - visualization
    affiliations:
      - ref: 1
      - ref: 2
    degrees:
      - HBSc
  - name: Alexander Rauscher
    orcid: 0000-0002-1961-8252
    email: rauscher@physics.ubc.ca
    corresponding: false
    roles:
      - writing - review & editing
    affiliations:
      - ref: 3
    degrees:
      - PhD
      - MSc
  - name: Ruth E Grunau
    orcid: 0000-0002-5428-9212
    corresponding: false
    email: rgrunau@mail.ubc.ca
    roles:
      - writing - review & editing
      - funding acquisition
    affiliations:
      - ref: 2
      - ref: 3
  - name: Alexander Mark Weber
    orcid: 0000-0001-7295-0775
    corresponding: true
    email: aweber@bcchr.ca
    roles:
      - project administration
      - supervision
      - validation
      - visualization
      - resources
      - methodology
      - formal analysis
      - funding acquisition
      - writing - review & editing
      - conceptualization
      - data curation
      - investigation
    affiliations:
      - ref: 2
      - ref: 3
    degrees:
      - PhD
      - MSc
affiliations:
  - id: 1
    name: The University of British Columbia
    department: Integrated Sciences
    address: 2329 West Mall
    city: Vancouver
    region: BC
    country: Canada
    postal-code: V6T 1Z4
  - id: 3
    name: The University of British Columbia
    department: Pediatrics
    address: 2329 West Mall
    city: Vancouver
    region: BC
    country: Canada
    postal-code: V6T 1Z4
  - id: 2
    department: BC Children's Hospital Research Institute
    name: The University of British Columbia
    address: 938 West 28th Avenue
    city: Vancouver
    state: BC
    country: Canada
    postal-code: V5Z 4H4

keywords:
  - Quantitative Susceptbility Mapping
  - Preterm
  - Newborn
  - Cerebral Venous Oxygen Saturation
abstract: |
  **Background**: Quantitative susceptibility mapping (QSM) is a magnetic resonance imaging (MRI) modality proposed to be a viable method of measuring cerebral oxygenation in neonates given its sensitivity to deoxyhemoglobin, a paramagnetic molecule. During QSM, however, paramagnetic sources can be obscured by opposing diamagnetic sources such as water and myelin. We sought to evaluate whether QSM images alone, or an algorithm that attempts to isolate their paramagnetic components, are more accurate in measuring oxygenation of the major cerebral veins in a cohort of neonates born preterm. Additionally, we aimed to determine whether a difference in oxygenation existed between the major cerebral veins.

  **Methods**: 19 neonates born preterm were scanned on a 3T research MRI at term equivalent age. The protocol included a multi-echo susceptibility-weighted imaging sequence. The acquired imaging data were processed as QSM images to obtain the susceptibility values of the superior sagittal sinus (SSS) and central cerebral veins (CCV). These values were used to calculate the oxygen saturation (SvO2) of the SSS and CCV.  QSM images were subsequently processed to isolate their paramagnetic components. SvO2 values of the SSS and CCV were calculated again from the paramagnetic components.

  **Results**: The mean SvO2 values of the SSS and CCV calculated from QSM images were found to be 72.4% (SD, 3.4%) and 68.7% (SD, 3.5%), respectively. The mean SvO2 values calculated from paramagnetic components were found to be 58.1% (SD, 7.3%) for the SSS and 57.7% (SD, 7.0%) for the CCV.

  **Conclusion**: SSS SvO2 values derived from paramagnetic components agreed well with the existing literature and were closer than the values derived from QSM, however, they displayed greater variability. Although the CCV SvO2 data from QSM aligns more closely with existing literature, it is important to note that the current literature on this topic remains relatively limited in the CCV. Thus, decomposing QSM images into paramagnetic components shows great promise as a method for more accurately measuring cerebral oxygenation in neonates but may require more research to improve precision. Notably, no significant difference in oxygenation was observed between the CCV and the SSS, contrasting with previous studies.

plain-language-summary: |

key-points:
  - 
date: last-modified
bibliography: [ Gavin_Thesis_Ref.bib ]
citation:
  container-title: Unpublished
number-sections: true
---

```{r setup, include=FALSE}
# options
knitr::opts_chunk$set(
  # fig.width=8, fig.height=5, 
  # out.width="50%",
  # fig.align="center",
  echo=FALSE, 
  message=FALSE, 
  warning=FALSE
  # cache=TRUE 
  )
set.seed(1234) # reproducible
options(knitr.kable.NA = '') # how kable handles NA
options(reticulate.repl.quiet = TRUE)
```

```{r libraries}
#libraries
library(tidyverse) # ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
theme_set( theme_minimal() ) # ggplot theme
library(broom) # for nice summaries
library(knitr) # 
library(kableExtra) # more tables options. Can cause problems
# library(reticulate) # incorporate Python
# use_virtualenv('./pythonvenv', required = TRUE) # load pythong venv from path
# library(neurobase)
```

# Introduction {#sec-intro}

<!-- should be around 5-6 paragraphs. Aim for 460 words -->

# Methods {#sec-data-methods}

The study was approved by the Clinical Research Ethics Board at the University of British Columbia and Children's & Women's Hospital (H21-00655) and written informed consent was obtained from the parent/guardian for each infant.

## Study population

Participant data comes from a previous study [@zhu-etal-cmro2]. Participants consisted of preterm neonates born between 25- and 31-weeks gestational age (GA) who were admitted to the level III NICU at BC Women’s Hospital. Recruitment took place over a span of one year, from February 2021 to January 2022, facilitated by a dedicated research nurse. Parents of eligible infants were approached by the research nurse prior to discharge from the NICU to explain the study objectives and seek their consent for participation. Infants meeting the criteria for inclusion were scanned for the study if they had already been discharged from the NICU, were in stable condition, and had reached a term equivalent age of 37 to 44 weeks GA. However, certain exclusion criteria were applied to ensure the homogeneity and integrity of the study sample. Infants were excluded if there was clinical evidence of a congenital malformation or syndrome, a TORCH infection, or ultrasound evidence of large parenchymal hemorrhagic infarction (\>2 cm, Grade 4 intraventricular hemorrhage).

## Image acquisition

MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750 scanner (scanner software version DV26.0_R03) equipped with a SREE Medical Systems single-channel neonatal head coil (@tbl-mri). The scans were conducted at the BC Children’s MRI Research Facility. Prior to the scanning procedure, subjects were carefully prepared by a research nurse. Swaddling and feeding were used to ensure the comfort and cooperation of the subjects during the scan. Importantly, no sedatives or invasive markers were utilized throughout the procedure. Subjects were placed within a specially designed SREE Medical Systems MRI compatible incubator, which facilitated both safety and motion minimization. Molded foam was strategically positioned around the head and body within the incubator to further restrict subject movement. To protect against potential hearing damage, ear plugs were employed during the scanning process. Additionally, a pulse oximeter was affixed to the subject’s foot to monitor arterial oxygen saturation and heart rate throughout the scan.

|                          | T1w          | T2w          | pcASL                                                               | SWI                                               |
|------------|------------|------------|----------------------|---------------|
| Sequence                 | 3D FSPGR     | 3D CUBE      | Multi-shot 3D \\\\ fast spin-echo                                   | 3D spoiled GRE flow-compensated                   |
| Phase-encoding direction | Coronal      | Sagittal     | Axial                                                               | Axial                                             |
| TR (ms)                  | 7.74         | 2,300        | 4,680                                                               | 30.9                                              |
| TE (ms)                  | 2.97         | 66.29        | 10.55                                                               | 5.24                                              |
| Flip angle               | 12$^{\circ}$ | 90$^{\circ}$ | 111$^{\circ}$                                                       | 20$^{\circ}$                                      |
| FOV (cm)                 | 20           | 20           | 24                                                                  | 25                                                |
| Acquisition Matrix       | 512 x 512    | 256 x 256    | 128 x 128                                                           | 256 x 256                                         |
| In-plane resolution (mm) | 0.39 x 0.39  | 0.78 x 0.78  | 1.875 x 1.875                                                       | 0.977 x 0.977                                     |
| Slice thickness (mm)     | 1            | 1            | 4                                                                   | 2, reconstructed to 1 mm with zero filling (ZIP2) |
| Number of slices         | 126          | 106          | 50                                                                  | 92                                                |
| Additional parameters    | n/a          | n/a          | 1,450 ms label period; 2,025 ms pulse lable; 24 control-label pairs | n/a                                               |
| Scan duration            | 4 min 39 s   | 5 min 1 s    | 5 min 26 s                                                          | 5 min 29 s                                        |

: Technical parameters for MR imaging protocol. {#tbl-mri}

The MRI scan protocol comprised several sequences, including a T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin labeling (ASL) scan, a multi-echo susceptibility-weighted imaging scan, and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging (EPI) sequence.

{{< embed notebooks/Figures.ipynb#fig-sample >}}

See @fig-graph

{{< embed notebooks/Figures.ipynb#fig-graph >}}

# Results {#sec-results}

[@ahmedDiamagneticComponentMap2023a]

# Conclusion {#sec-conclusion}

# References {.unnumbered}

::: {#refs}
:::

### 2024-05-24 _tex/index.tex
Built site for gh-pages

@@
 \documentclass[
-]{agujournal2019}
+sn-nature
+]{sn-jnl}
 
@@
 
-\usepackage{url} %this package should fix any errors with URLs in refs.
-\usepackage{lineno}
-\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
-\usepackage{soul}
-\linenumbers
+%%%% Standard Packages
+
+\usepackage{graphicx}%
+\usepackage{multirow}%
+\usepackage{amsmath,amssymb,amsfonts}%
+\usepackage{amsthm}%
+\usepackage{mathrsfs}%
+\usepackage[title]{appendix}%
+\usepackage{xcolor}%
+\usepackage{textcomp}%
+\usepackage{manyfoot}%
+\usepackage{booktabs}%
+\usepackage{algorithm}%
+\usepackage{algorithmicx}%
+\usepackage{algpseudocode}%
+\usepackage{listings}%
+
+%%%%
+
+\raggedbottom
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 \makeatletter
@@
 \hypersetup{
-  pdftitle={La Palma Earthquakes},
-  pdfauthor={Steve Purves; Rowan Cockett},
-  pdfkeywords={La Palma, Earthquakes},
+  pdftitle={ParaChiCSVO2},
+  pdfauthor={Thomas Gavin Carmichael; Alexander Rauscher; Ruth E Grunau; Alexander Mark Weber},
+  pdfkeywords={Quantitative Susceptbility
+Mapping, Preterm, Newborn, Cerebral Venous Oxygen Saturation},
   colorlinks=true,
@@
 
-\journalname{Earth and Space Science}
+\title[ParaChiCSVO2]{ParaChiCSVO2}
 
-\draftfalse
+% author setup
+\author[1,2]{\fnm{Thomas Gavin} \sur{Carmichael}}\email{tgcarmichael@outlook.com}\author[3]{\fnm{Alexander} \sur{Rauscher}}\email{rauscher@physics.ubc.ca}\author[2,3]{\fnm{Ruth E} \sur{Grunau}}\email{rgrunau@mail.ubc.ca}\author*[2,3]{\fnm{Alexander Mark} \sur{Weber}}\email{aweber@bcchr.ca}
+% affil setup
+\affil[1]{\orgdiv{Integrated Sciences}, \orgname{The University of
+British Columbia}, \orgaddress{\street{2329 West
+Mall}, \city{Vancouver}, \postcode{V6T 1Z4}, \country{Canada}}}
+\affil[3]{\orgdiv{Pediatrics}, \orgname{The University of British
+Columbia}, \orgaddress{\street{2329 West
+Mall}, \city{Vancouver}, \postcode{V6T 1Z4}, \country{Canada}}}
+\affil[2]{\orgdiv{BC Children's Hospital Research
+Institute}, \orgname{The University of British
+Columbia}, \orgaddress{\street{938 West 28th
+Avenue}, \city{Vancouver}, \postcode{V5Z 4H4}, \country{Canada}}}
+
+% abstract
+
+\abstract{\textbf{Background}: Quantitative susceptibility mapping (QSM)
+is a magnetic resonance imaging (MRI) modality proposed to be a viable
+method of measuring cerebral oxygenation in neonates given its
+sensitivity to deoxyhemoglobin, a paramagnetic molecule. During QSM,
+however, paramagnetic sources can be obscured by opposing diamagnetic
+sources such as water and myelin. We sought to evaluate whether QSM
+images alone, or an algorithm that attempts to isolate their
+paramagnetic components, are more accurate in measuring oxygenation of
+the major cerebral veins in a cohort of neonates born preterm.
+Additionally, we aimed to determine whether a difference in oxygenation
+existed between the major cerebral veins.
+
+\textbf{Methods}: 19 neonates born preterm were scanned on a 3T research
+MRI at term equivalent age. The protocol included a multi-echo
+susceptibility-weighted imaging sequence. The acquired imaging data were
+processed as QSM images to obtain the susceptibility values of the
+superior sagittal sinus (SSS) and central cerebral veins (CCV). These
+values were used to calculate the oxygen saturation (SvO2) of the SSS
+and CCV. QSM images were subsequently processed to isolate their
+paramagnetic components. SvO2 values of the SSS and CCV were calculated
+again from the paramagnetic components.
+
+\textbf{Results}: The mean SvO2 values of the SSS and CCV calculated
+from QSM images were found to be 72.4\% (SD, 3.4\%) and 68.7\% (SD,
+3.5\%), respectively. The mean SvO2 values calculated from paramagnetic
+components were found to be 58.1\% (SD, 7.3\%) for the SSS and 57.7\%
+(SD, 7.0\%) for the CCV.
+
+\textbf{Conclusion}: SSS SvO2 values derived from paramagnetic
+components agreed well with the existing literature and were closer than
+the values derived from QSM, however, they displayed greater
+variability. Although the CCV SvO2 data from QSM aligns more closely
+with existing literature, it is important to note that the current
+literature on this topic remains relatively limited in the CCV. Thus,
+decomposing QSM images into paramagnetic components shows great promise
+as a method for more accurately measuring cerebral oxygenation in
+neonates but may require more research to improve precision. Notably, no
+significant difference in oxygenation was observed between the CCV and
+the SSS, contrasting with previous studies.}
+
+% keywords
+\keywords{Quantitative Susceptbility
+Mapping,  Preterm,  Newborn,  Cerebral Venous Oxygen Saturation}
 
 \begin{document}
-\title{La Palma Earthquakes}
-
-\authors{Steve Purves\affil{1}, Rowan Cockett\affil{1}}
-\affiliation{1}{Curvenote, }
-\correspondingauthor{Steve Purves}{steve@curvenote.com}
-
-\begin{abstract}
-In September 2021, a significant jump in seismic activity on the island
-of La Palma (Canary Islands, Spain) signaled the start of a volcanic
-crisis that still continues at the time of writing. Earthquake data is
-continually collected and published by the Instituto Geográphico
-Nacional (IGN). \ldots{}
-\end{abstract}
-
-\section*{Plain Language Summary}
-Earthquake data for the island of La Palma from the September 2021
-eruption is found \ldots{}
-
-\section{Introduction}\label{introduction}
+\maketitle
 
@@
 
-\phantomsection\label{cell-fig-timeline}
-\begin{figure}[H]
+\section{Introduction}\label{sec-intro}
 
-\centering{
+\section{Methods}\label{sec-data-methods}
 
-\includegraphics{index_files/figure-pdf/fig-timeline-1.pdf}
+The study was approved by the Clinical Research Ethics Board at the
+University of British Columbia and Children's \& Women's Hospital
+(H21-00655) and written informed consent was obtained from the
+parent/guardian for each infant.
 
-}
+\subsection{Study population}\label{study-population}
 
-\caption{\label{fig-timeline}Timeline of recent earthquakes on La Palma}
+Participant data comes from a previous study (Zhu et al. UNDER REVIEW
+May 2024). Participants consisted of preterm neonates born between 25-
+and 31-weeks gestational age (GA) who were admitted to the level III
+NICU at BC Women's Hospital. Recruitment took place over a span of one
+year, from February 2021 to January 2022, facilitated by a dedicated
+research nurse. Parents of eligible infants were approached by the
+research nurse prior to discharge from the NICU to explain the study
+objectives and seek their consent for participation. Infants meeting the
+criteria for inclusion were scanned for the study if they had already
+been discharged from the NICU, were in stable condition, and had reached
+a term equivalent age of 37 to 44 weeks GA. However, certain exclusion
+criteria were applied to ensure the homogeneity and integrity of the
+study sample. Infants were excluded if there was clinical evidence of a
+congenital malformation or syndrome, a TORCH infection, or ultrasound
+evidence of large parenchymal hemorrhagic infarction (\textgreater2 cm,
+Grade 4 intraventricular hemorrhage).
 
-\end{figure}%
+\subsection{Image acquisition}\label{image-acquisition}
 
-\textsubscript{Source:
-\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
-Notebook}}
+MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750
+scanner (scanner software version DV26.0\_R03) equipped with a SREE
+Medical Systems single-channel neonatal head coil (Table~\ref{tbl-mri}).
+The scans were conducted at the BC Children's MRI Research Facility.
+Prior to the scanning procedure, subjects were carefully prepared by a
+research nurse. Swaddling and feeding were used to ensure the comfort
+and cooperation of the subjects during the scan. Importantly, no
+sedatives or invasive markers were utilized throughout the procedure.
+Subjects were placed within a specially designed SREE Medical Systems
+MRI compatible incubator, which facilitated both safety and motion
+minimization. Molded foam was strategically positioned around the head
+and body within the incubator to further restrict subject movement. To
+protect against potential hearing damage, ear plugs were employed during
+the scanning process. Additionally, a pulse oximeter was affixed to the
+subject's foot to monitor arterial oxygen saturation and heart rate
+throughout the scan.
 
-\textsubscript{Source:
-\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
-Notebook}}
-
-Based on data up to and including 1971, eruptions on La Palma happen
-every 79.8 years on average.
-
-Studies of the magma systems feeding the volcano, such as Marrero et al.
-(2019), have proposed that there are two main magma reservoirs feeding
-the Cumbre Vieja volcano; one in the mantle (30-40km depth) which
-charges and in turn feeds a shallower crustal reservoir (10-20km depth).
-
-Eight eruptions have been recorded since the late 1400s
-(Figure~\ref{fig-timeline}).
-
-Data and methods are discussed in Section~\ref{sec-data-methods}.
-
-Let \(x\) denote the number of eruptions in a year. Then, \(x\) can be
-modeled by a Poisson distribution
-
-\begin{equation}\phantomsection\label{eq-poisson}{
-p(x) = \frac{e^{-\lambda} \lambda^{x}}{x !}
-}\end{equation}
-
-where \(\lambda\) is the rate of eruptions per year. Using
-Equation~\ref{eq-poisson}, the probability of an eruption in the next
-\(t\) years can be calculated.
-
-\begin{longtable}[]{@{}ll@{}}
-\caption{Recent historic eruptions on La
-Palma}\label{tbl-history}\tabularnewline
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1644}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1644}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1644}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3014}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2055}}@{}}
+\caption{Technical parameters for MR imaging
+protocol.}\label{tbl-mri}\tabularnewline
 \toprule\noalign{}
-Name & Year \\
+\begin{minipage}[b]{\linewidth}\raggedright
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+T1w
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+T2w
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+pcASL
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+SWI
+\end{minipage} \\
 \midrule\noalign{}
@@
 \toprule\noalign{}
-Name & Year \\
+\begin{minipage}[b]{\linewidth}\raggedright
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+T1w
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+T2w
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+pcASL
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+SWI
+\end{minipage} \\
 \midrule\noalign{}
@@
 \endlastfoot
-Current & 2021 \\
-Teneguía & 1971 \\
-Nambroque & 1949 \\
-El Charco & 1712 \\
-Volcán San Antonio & 1677 \\
-Volcán San Martin & 1646 \\
-Tajuya near El Paso & 1585 \\
-Montaña Quemada & 1492 \\
+Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D
+\textbackslash\textbackslash{} fast spin-echo & 3D spoiled GRE
+flow-compensated \\
+Phase-encoding direction & Coronal & Sagittal & Axial & Axial \\
+TR (ms) & 7.74 & 2,300 & 4,680 & 30.9 \\
+TE (ms) & 2.97 & 66.29 & 10.55 & 5.24 \\
+Flip angle & 12\(^{\circ}\) & 90\(^{\circ}\) & 111\(^{\circ}\) &
+20\(^{\circ}\) \\
+FOV (cm) & 20 & 20 & 24 & 25 \\
+Acquisition Matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256 \\
+In-plane resolution (mm) & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 &
+0.977 x 0.977 \\
+Slice thickness (mm) & 1 & 1 & 4 & 2, reconstructed to 1 mm with zero
+filling (ZIP2) \\
+Number of slices & 126 & 106 & 50 & 92 \\
+Additional parameters & n/a & n/a & 1,450 ms label period; 2,025 ms
+pulse lable; 24 control-label pairs & n/a \\
+Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s \\
 \end{longtable}
 
-Table~\ref{tbl-history} summarises the eruptions recorded since the
-colonization of the islands by Europeans in the late 1400s.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{images/la-palma-map.png}
-
-}
-
-\caption{\label{fig-map}Map of La Palma}
-
-\end{figure}%
-
-La Palma is one of the west most islands in the Volcanic Archipelago of
-the Canary Islands (Figure~\ref{fig-map}).
+The MRI scan protocol comprised several sequences, including a
+T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin
+labeling (ASL) scan, a multi-echo susceptibility-weighted imaging scan,
+and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging
+(EPI) sequence.
 
@@
 
-\includegraphics{index_files/figure-latex/notebooks-explore-earthquakes-fig-spatial-plot-output-1.png}
+\includegraphics{index_files/figure-latex/notebooks-Figures-fig-sample-output-1.png}
 
@@
 
-\caption{\label{fig-spatial-plot}Locations of earthquakes on La Palma
-since 2017}
+\caption{\label{fig-sample}Sample images}
 
@@
 \textsubscript{Source:
-\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/explore-earthquakes-preview.html\#cell-fig-spatial-plot}{Explore
-Earthquakes}}
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures.ipynb.html\#cell-fig-sample}{Python
+Figures}}
 
-Figure~\ref{fig-spatial-plot} shows the location of recent Earthquakes
-on La Palma.
+See Figure~\ref{fig-graph}
 
-\section{Data \& Methods}\label{sec-data-methods}
+\begin{figure}[H]
 
-\section{Conclusion}\label{conclusion}
+\centering{
+
+\includegraphics{index_files/figure-latex/notebooks-Figures-fig-graph-output-1.png}
+
+}
+
+\caption{\label{fig-graph}A graph of the methods to create a mask}
+
+\end{figure}%
+
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures.ipynb.html\#cell-fig-graph}{Python
+Figures}}
+
+\section{Results}\label{sec-results}
+
+(Ahmed et al. 2023)
+
+\section{Conclusion}\label{sec-conclusion}
 
@@
 \begin{CSLReferences}{1}{0}
-\vspace{1em}
+\bibitem[\citeproctext]{ref-ahmedDiamagneticComponentMap2023a}
+Ahmed, Maruf, Jingjia Chen, Arvin Arani, Matthew L. Senjem, Petrice M.
+Cogswell, Clifford R. Jack, and Chunlei Liu. 2023. {``The Diamagnetic
+Component Map from Quantitative Susceptibility Mapping ({QSM}) Source
+Separation Reveals Pathological Alteration in {Alzheimer}'s
+Disease-Driven Neurodegeneration.''} \emph{NeuroImage} 280 (October):
+120357. \url{https://doi.org/10.1016/j.neuroimage.2023.120357}.
 
-\bibitem[\citeproctext]{ref-marrero2019}
-Marrero, J., García, A., Berrocoso, M., Llinares, Á., Rodríguez-Losada,
-A., \& Ortiz, R. (2019). Strategies for the development of volcanic
-hazard maps in monogenetic volcanic fields: The example of {La} {Palma}
-({Canary} {Islands}). \emph{Journal of Applied Volcanology}, \emph{8}.
-\url{https://doi.org/10.1186/s13617-019-0085-5}
+\bibitem[\citeproctext]{ref-zhu-etal-cmro2}
+Zhu, A., C. Chau, N. Chan, A. Chacko, L. Holsti, R. E. Grunau, and A. M.
+Weber. UNDER REVIEW May 2024. {``Regional Cerebral Metabolic Rate of
+Oxygen and Levels of Respiratory Support in Preterm Neonates.''}
+\emph{Pediatric Research}, UNDER REVIEW May 2024.
 

### 2024-05-24 index.tex
rendered

--- 2024-05-24 _tex/index.tex
@@
 % Make \paragraph and \subparagraph free-standing
-\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}{
-    \@ifstar
-      \xxxParagraphStar
-      \xxxParagraphNoStar
-  }
-  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
-  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}{
-    \@ifstar
-      \xxxSubParagraphStar
-      \xxxSubParagraphNoStar
-  }
-  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
-  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
-\makeatother
 
@@
 \endlastfoot
-Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D
-\textbackslash\textbackslash{} fast spin-echo & 3D spoiled GRE
-flow-compensated \\
+Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D fast spin-echo & 3D
+spoiled GRE flow-compensated \\
 Phase-encoding direction & Coronal & Sagittal & Axial & Axial \\

### 2024-05-24 index.qmd
Built site for gh-pages

@@
 |------------|------------|------------|----------------------|---------------|
-| Sequence                 | 3D FSPGR     | 3D CUBE      | Multi-shot 3D \\\\ fast spin-echo                                   | 3D spoiled GRE flow-compensated                   |
+| Sequence                 | 3D FSPGR     | 3D CUBE      | Multi-shot 3D fast spin-echo                                        | 3D spoiled GRE flow-compensated                   |
 | Phase-encoding direction | Coronal      | Sagittal     | Axial                                                               | Axial                                             |

### 2024-05-24 index.qmd
html

@@
 
-Participant data comes from a previous study [@zhu-etal-cmro2].
+Participant data comes from a previous study.
 Participants consisted of preterm neonates born between 25- and 31-weeks gestational age (GA) who were admitted to the level III NICU at BC Women’s Hospital.
@@
 |                          | T1w          | T2w          | pcASL                                                               | SWI                                               |
-|------------|------------|------------|----------------------|---------------|
+|--------------|--------------|--------------|-----------------|--------------|
 | Sequence                 | 3D FSPGR     | 3D CUBE      | Multi-shot 3D fast spin-echo                                        | 3D spoiled GRE flow-compensated                   |

### 2024-05-31 index.qmd
May31st, working version

--- 2024-05-24 index.qmd
@@
   container-title: Unpublished
-number-sections: true
+number-sections: false
+notebook-links: false
 ---
@@
 library(kableExtra) # more tables options. Can cause problems
+library(Rmpfr)
 # library(reticulate) # incorporate Python
@@
 # library(neurobase)
+```
+
+```{r}
+load("notebooks/results.RData")
 ```
@@
 
-MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750 scanner (scanner software version DV26.0_R03) equipped with a SREE Medical Systems single-channel neonatal head coil (@tbl-mri).
+MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750 scanner (scanner software version DV26.0_R03) equipped with a SREE Medical Systems single-channel neonatal head coil (Table \ref{tbl-mri}).
 The scans were conducted at the BC Children’s MRI Research Facility.
@@
 
-|                          | T1w          | T2w          | pcASL                                                               | SWI                                               |
-|------------|------------|------------|----------------------|---------------|
-| Sequence                 | 3D FSPGR     | 3D CUBE      | Multi-shot 3D fast spin-echo                                        | 3D spoiled GRE flow-compensated                   |
-| Phase-encoding direction | Coronal      | Sagittal     | Axial                                                               | Axial                                             |
-| TR (ms)                  | 7.74         | 2,300        | 4,680                                                               | 30.9                                              |
-| TE (ms)                  | 2.97         | 66.29        | 10.55                                                               | 5.24                                              |
-| Flip angle               | 12$^{\circ}$ | 90$^{\circ}$ | 111$^{\circ}$                                                       | 20$^{\circ}$                                      |
-| FOV (cm)                 | 20           | 20           | 24                                                                  | 25                                                |
-| Acquisition Matrix       | 512 x 512    | 256 x 256    | 128 x 128                                                           | 256 x 256                                         |
-| In-plane resolution (mm) | 0.39 x 0.39  | 0.78 x 0.78  | 1.875 x 1.875                                                       | 0.977 x 0.977                                     |
-| Slice thickness (mm)     | 1            | 1            | 4                                                                   | 2, reconstructed to 1 mm with zero filling (ZIP2) |
-| Number of slices         | 126          | 106          | 50                                                                  | 92                                                |
-| Additional parameters    | n/a          | n/a          | 1,450 ms label period; 2,025 ms pulse lable; 24 control-label pairs | n/a                                               |
-| Scan duration            | 4 min 39 s   | 5 min 1 s    | 5 min 26 s                                                          | 5 min 29 s                                        |
+```{=tex}
+\begin{table}[h]
+  \caption{Technical parameters for MR imaging pulse sequences}\label{tbl-mri}%
+    \centering
+    \begin{tabular*}{\textwidth}{@{}lllll@{}}
+    \toprule
+        ~ & T1w & T2w & pcASL & SWI \\ \midrule
+        Sequence & 3D FSPGR & 3D CUBE & \makecell[tl]{Multi-shot 3D fast \\ spin-echo} & \makecell[tl]{3D spoiled GRE \\ flow-compensated} \\ 
+        \makecell[tl]{Phase-encoding \\ direction} & Coronal & Sagittal & Axial & Axial \\ 
+        TR (ms) & 7.74 & 2,300 & 4,680 & 30.9 \\ 
+        TE (ms) & 2.97 & 66.29 & 10.55 & 5.24 \\ 
+        Flip angle & 12$^{\circ}$ & 90$^{\circ}$ & 111$^{\circ}$ & 20$^{\circ}$ \\ 
+        FOV (cm) & 20 & 20 & 24 & 25 \\ 
+        Acquisition Matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256 \\ 
+        \makecell[tl]{In-plane resolution \\ (mm)} & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 & 0.977 x 0.977 \\ 
+        \makecell[tl]{Slice thickness \\ (mm)} & 1 & 1 & 4 & \makecell[tl]{2, reconstructed \\ to 1 with zero \\ filling (ZIP2)} \\ 
+        Number of slices & 126 & 106 & 50 & 92 \\ 
+        Additional parameters & n/a & n/a & \makecell[tl]{1,450 ms label period; \\ 2,025 ms pulse lable; \\ 24 control-label pairs} & n/a \\ 
+        Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s \\ \botrule
+    \end{tabular*}
+    \footnotetext{T1w = T1-weighted; T2w = T2-weighted; pcASL = pseudo-continuous arterial spin labelling; SWI = susceptibility weighted imaging; FSPG = fast spoiled gradient echo; CUBE = General Electric name of sequence, not an acronym; GRE = gradient echo; ZIP2 = through-plane zero filling interpolation}
+\end{table}
+```
+The MRI scan protocol comprised of the following sequences: a T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin labeling (ASL) scan, a multi-echo susceptibility-weighted imaging scan, and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging (EPI) sequence.
+The DWI sequence was not used for the present study.
 
-: Technical parameters for MR imaging protocol. {#tbl-mri}
+## Image analysis
 
-The MRI scan protocol comprised several sequences, including a T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin labeling (ASL) scan, a multi-echo susceptibility-weighted imaging scan, and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging (EPI) sequence.
+The raw DICOM files acquired from the scanning procedure were converted to NIfTI (Neuroimaging Informatics Technology Initiative) format using Chris Rorden’s \texttt{dcmniix} tool.
+SWI magnitude data files were then used to create subject-specific brain masks that would not erode the SSS during QSM processing, an issue faced by our group in the past [@liFirstStepNeuroimaging2016].
+A step-by-step summary of the pipeline used is shown in @fig-sample.
+
+{{< embed notebooks/Figures.ipynb#fig-graph >}}
+
+First, the fifth echo SWI magnitude file was processed using FSL’s (v.
+6.0.7.3) [@woolrichBayesianAnalysisNeuroimaging2009] \texttt{fslroi}, \texttt{fslmaths}, and \texttt{bet} [@smithFastRobustAutomated2002] to create a preliminary brain mask, similar to our previous efforts, which does not contain the SSS.
+\texttt{Fslroi} was used to isolate the fifth echo of the magnitude data, which was then squared using Fslmaths and the option \texttt{-sqr}.
+Squaring the magnitude image was found to dramatically improve subsequent brain extraction.
+The resulting image was then used to create the preliminary brain mask using bet with the options \texttt{-m} and \texttt{-R}.
+The former flag generated a binary brain mask, while the latter performed a more robust brain centre estimation.
+The brain mask was then dilated by 7 voxels using \texttt{Fslmaths} and the options \texttt{-kernel boxv} and \texttt{-dilM} in order for the dilated mask to contain the SSS (along with unwanted tissue as well).
+This mask was then used, along with the phase images, in a MATLAB script for QSM calculation from Christian Kames [@kamesRapidTwostepDipole2018] to produce a preliminary QSM image that contained the SSS, albeit with fairly low signal-to-noise ratio and other unwanted tissue.
+Given the high contrast in voxel intensity between the SSS and surrounding tissue, the select by intensity tool in \texttt{FSLeyes} [@mccarthyFSLeyes2023] was then used to segment the SSS from the QSM image and create a 3D mask of the selected region.
+Using \texttt{fslmaths} and the options \texttt{-add} and \texttt{-bin}, the SSS mask was then combined with the original brain mask of the fifth echo.
+This resulted in a brain mask that contained only brain and SSS signal.
+Finally, this mask was used in a final QSM post-processing step to create a QSM image that includes the SSS while maintaining a high signal-to-noise ratio, making it suitable to obtain accurate susceptibility values.
+
+STI Suite (v.
+3.0) [@liIntegratedLaplacianBased2014], was used to process the final QSM images as it produced the cleanest images without eroding the SSS.
+The finalized brain mask and the last three echoes of the magnitude and phase images were used in STI Suite along with the following parameters: 0.9766 x 0.9766 x 1 mm$^{3}$ voxel size, 5 ms TE1, 5.3 ms $\Delta$TE, and 77.4 ms sum TE, B0 strength = 3, and B0 direction = [0, 0, 1].
+The 3D GRE data option was selected for the phase processing stage, and STAR-QSM was selected for the QSM stage.
+Finally, the `select by intensity' tool in \texttt{FSLeyes} was then used to semi-automatically make vascular masks of the SSS and CCV from each subject’s QSM image.
+The vascular masks were used to calculate the mean susceptibility of each subject’s SSS and CCV from their QSM image with \texttt{fslstats}.
+
+To isolate the paramagnetic component of subjects’ QSM data, the χ-separation toolbox28 from the Laboratory for Imaging Science and Technology was used.
+Each subject’s magnitude and phase SWI data were used along with the following parameters: 0.9766 x 0.9766 x 1 mm$^{3}$ voxel size; TE (s) = [0.005, 0.0102, 0.0155, 0.0207, 0.026]; delta TE (s) = 0.0052; B0 strength = 3; B0 direction = [0, 0, 1].
+The mean susceptibility of each subject’s SSS and CCV in their paramagnetic maps was calculated with the same vascular masks used for the QSM images.
+Sample images showing the magnitude, final QSM, and final paramagnetic component images are shown in @fig-sample.
 
@@
 
-See @fig-graph
+Once the mean susceptibility values of the SSS and CCV were obtained from the subjects’ QSM images and paramagnetic maps, venous oxygen saturation (SvO2) was calculated with the following equation [@bergInvestigatingEffectFlow2021]:
 
-{{< embed notebooks/Figures.ipynb#fig-graph >}}
+$$
+SvO_{2} = 1 - \frac{\Delta \chi _{blood} - (\Delta \chi _{oxy} * Hct)}{\Delta \chi _{do} * Hct}
+$$ {#eq-svo}
+
+where $\Delta \chi _{blood}$ is the vessel’s measured susceptibility, $\Delta \chi _{oxy}$ is the constant representing the susceptibility changes of oxygenated blood relation to water, $\Delta \chi _{do}$ is the difference in susceptibility between oxygenated and deoxygenated blood, and Hct is the subject’s hematocrit.
+$\Delta \chi _{oxy}$ was -0.21 * 4$\pi$ ppm as per @portnoyHumanUmbilicalCord2018 and [@sedlacikObtainingBloodOxygenation2007], while $\Delta \chi _{do}$ was -0.03 * 4$\pi$ ppm as per [@weisskoffMRISusceptometryImagebased1992].
+Subjects’ Hct for the day of the scan was calculated using a four-parameter Weibull function with previously measured values while still in the NICU.
+
+## Statistical analysis
+
+Statistical analysis of the acquired data was performed using R and RStudio (v.
+2023.09.1 Build 494) [@rcoreteamLanguageEnvironmentStatistical2022; @rstudioteamRStudioIntegratedDevelopment].
+A Wilcoxon signed-rank test was used to determine if a significant difference (p <0.05) existed between the vessel-specific SvO2 values determined through QSM and those determined from the paramagnetic maps processed with $\chi$-separation.
+This was also performed for the raw $\chi$ values obtained through both methods.
 
@@
 
-[@ahmedDiamagneticComponentMap2023a]
+A total sample size of 19 infants were scanned, with a mean gestational age of 28.8 weeks (SD, 1.7) and a mean birth weight of 1276.1 grams (SD, 294.9).
+A comprehensive summary of neonatal characteristics, including additional demographic and clinical data, is provided in Table \ref{tbl-chistats} for reference.
+
+```{=latex}
+\begin{table*}[h]
+  \caption{Technical parameters for MR imaging pulse sequences}\label{tbl-chistats}%
+    \centering
+    \begin{tabular*}{\textwidth}{@{}llllll@{}}
+    \toprule
+        Region & Measure & QSM & \makecell[tl]{Paramagnetic \\ map} & p-value & 95\% CI \\ \midrule
+        \multirow{ 2}{*}{SSS} & Chi (ppm) & `r round(mean(newdata$Gavin_SSS_Chi),2)` \pm `r round(sd(newdata$Gavin_SSS_Chi),2)` & `r round(mean(newdata$Chisep_SSS_Chi),2)` \pm `r round(sd(newdata$Chisep_SSS_Chi),2)` & `r signif(SSSChi.t$p.value,3)` & `r round(SSSChi.t$conf.low,2)`, `r round(SSSChi.t$conf.high,2)`\\
+        & SvO$_{2}$(\%) & `r round(mean(newdata$Gavin_SSSVein_CSvO2*100),2)` \pm `r round(sd(newdata$Gavin_SSSVein_CSvO2*100),2)` & `r round(mean(newdata$Chisep_SSSVein_CSvO2*100),2)` \pm `r round(sd(newdata$Chisep_SSSVein_CSvO2*100),2)` & `r signif(SSSSvO2.t$p.value,3)` & `r round(SSSSvO2.t$conf.low*100,2)`, `r round(SSSSvO2.t$conf.high*100,2)`\\
+        \multirow{ 2}{*}{CCV} & Chi (ppm) & `r round(mean(newdata$Gavin_IntVein_Chi),2)` \pm `r round(sd(newdata$Gavin_IntVein_Chi),2)` & `r round(mean(newdata$Chisep_IntVein_Chi),2)` \pm `r round(sd(newdata$Chisep_IntVein_Chi),2)` & `r signif(CCVChi.t$p.value,3)` & `r round(CCVChi.t$conf.low,2)`, `r round(CCVChi.t$conf.high,2)`\\
+        & SvO$_{2}$(\%) & `r round(mean(newdata$Gavin_IntVein_CSvO2*100),2)` \pm `r round(sd(newdata$Gavin_IntVein_CSvO2*100),2)` & `r round(mean(newdata$Chisep_IntVein_CSvO2*100),2)` \pm `r round(sd(newdata$Chisep_IntVein_CSvO2*100),2)` & `r signif(CCVSvO2.t$p.value,3)` & `r round(CCVSvO2.t$conf.low*100,2)`, `r round(CCVSvO2.t$conf.high*100,2)`\\
+        \botrule
+    \end{tabular*}
+    \footnotetext{}
+\end{table*}
+```
+
+<!-- ::: {#tbl-1} -->
+
+<!-- ```{=latex} -->
+<!-- % Or whatever actual you want, includegraphics, whatever.  -->
+<!-- % (_not_ \begin{table}, just the content of it) -->
+<!-- \begin{tabular}{lll} -->
+<!-- \end{tabular} -->
+<!-- ``` -->
+
+<!-- This is a table caption. -->
+
+<!-- ::: -->
+
+<!-- See @tbl-1 -->
+
+See @fig-regionplot
+
+{{< embed notebooks/gavin_thesis_markdown.qmd#fig-regionplot >}}
 

### 2024-05-31 notebooks/gavin_thesis_markdown.Rmd
May31st, working version

---
title: "gavin_thesis_R"
author: "Gavin"
date: "2024-04-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(ggstatsplot)
library(rstatix)
library(patchwork)
```

Load main dataframe (change directory as needed)
*note: columns titled with with "Gavin_x_x" contain QSM data
```{r}
newdata <- read.csv("newdata.csv")
```

Dataframes for figures
```{r, echo=FALSE}

#QSM chi data - SSS vs. CCV
comp_legit_starfsl <- data.frame(
  Chi = c(newdata$Gavin_SSS_Chi, newdata$Gavin_IntVein_Chi),
  Vein = c(rep("Superior Sagittal Sinus", 19), rep("Central Cerebral Veins", 19))
)

#QSM CSvO2 data - SSS vs. CCV
comp_SSS_IntVein_CSvO2 <- data.frame(
  CSvO2 = c(newdata$Gavin_IntVein_CSvO2, newdata$Gavin_SSSVein_CSvO2),
  Vein = c(rep("Central Cerebral Veins", 19), rep("Superior Sagittal Sinus", 19))
)

#Chisep chi data - SSS vs. CCV
comp_chisep_chi <- data.frame(
  Chi = c(newdata$Chisep_IntVein_Chi, newdata$Chisep_SSS_Chi),
  Vein = c(rep("Central Cerebral Veins", 19), rep("Superior Sagittal Sinus", 19))
)

#Chisep CSvO2 data - SSS vs. CCV
comp_chisep_CSvO2 <- data.frame(
  CSvO2 = c(newdata$Chisep_IntVein_CSvO2, newdata$Chisep_SSSVein_CSvO2),
  Vein = c(rep("Central Cerebral Veins", 19), rep("Superior Sagittal Sinus", 19))
)

#QSM SSS chi vs. chisep SSS chi
comp_SSS_chi <-  data.frame(
  "Measured Susceptibility" = 
    c(newdata$Gavin_SSS_Chi, newdata$Chisep_SSS_Chi),
  Method = c(rep("QSM", 19), rep("Paramagnetic Component", 19))
)

#QSM CCV chi vs. chisep CCV chi
comp_CCV_chi <- data.frame(
  Chi = 
    c(newdata$Gavin_IntVein_Chi, newdata$Chisep_IntVein_Chi),
  Method = c(rep("QSM", 19), rep("Paramagnetic Component", 19))
)

#QSM SSS SvO2 vs. chisep SSS SvO2
comp_SSS_SvO2 <- data.frame(
  "Oxygen Saturation" = 
    c(newdata$Gavin_SSSVein_CSvO2, newdata$Chisep_SSSVein_CSvO2),
  Method = c(rep("QSM", 19), rep("Paramagnetic Component", 19))
)

#QSM CCV SvO2 vs. chisep CCV SvO2
comp_CCV_SvO2 <- data.frame(
  "Oxygen Saturation" = 
    c(newdata$Gavin_IntVein_CSvO2, newdata$Chisep_IntVein_CSvO2),
  Method = c(rep("QSM", 19), rep("Paramagnetic Component", 19))
)
```

Wilcox test for p-values and 95% CI - table 3.2
```{r}
#SSS Chi
wilcox.test(newdata$Gavin_SSS_Chi, newdata$Chisep_SSS_Chi,
            paired = TRUE, conf.int = TRUE)

#SSS SvO2
wilcox.test(newdata$Gavin_SSSVein_CSvO2, newdata$Chisep_SSSVein_CSvO2,
            paired = TRUE, conf.int = TRUE)

#CCV Chi
wilcox.test(newdata$Gavin_IntVein_Chi, newdata$Chisep_IntVein_Chi, paired = TRUE,
            conf.int = TRUE)

#CCV SvO2
wilcox.test(newdata$Gavin_IntVein_CSvO2, newdata$Chisep_IntVein_CSvO2,
            paired = TRUE, conf.int = TRUE)
```

```{r}
#SSS Chi
SSSChi.t <- tidy(t.test(newdata$Gavin_SSS_Chi, newdata$Chisep_SSS_Chi,
            paired = TRUE, conf.int = TRUE))

#SSS SvO2
SSSSvO2.t <- tidy(t.test(newdata$Gavin_SSSVein_CSvO2, newdata$Chisep_SSSVein_CSvO2,
            paired = TRUE, conf.int = TRUE))

#CCV Chi
CCVChi.t <- tidy(t.test(newdata$Gavin_IntVein_Chi, newdata$Chisep_IntVein_Chi, paired = TRUE,
            conf.int = TRUE))

#CCV SvO2
CCVSvO2.t <- tidy(t.test(newdata$Gavin_IntVein_CSvO2, newdata$Chisep_IntVein_CSvO2,
            paired = TRUE, conf.int = TRUE))
```

Figure 3.1
*note: "****" in panels were added outside of R
```{r, echo=FALSE}
m1 <- ggwithinstats(
  data = comp_CCV_chi,
  x = Method,
  y = Chi,
  results.subtitle = FALSE,
  title = "(A)",
  subtitle = "Central Cerebral Veins",
  type = "nonparametric",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid")
)

m2 <- ggwithinstats( 
  data = comp_SSS_chi,
  x = Method,
  y = Measured.Susceptibility, 
  ylab = "Chi",
  results.subtitle = FALSE,
  title = "",
  subtitle = "Superior Sagittal Sinus",
  type = "nonparametric", 
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid")
)

m3 <- ggwithinstats(
  data = comp_CCV_SvO2,
  x = Method,
  y = Oxygen.Saturation,
  ylab = "SvO2",
  results.subtitle = FALSE,
  title = "(B)",
  subtitle = "Central Cerebral Veins",
  type = "nonparametric",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid")
)

m4 <- ggwithinstats(
  data = comp_SSS_SvO2,
  x = Method,
  y = Oxygen.Saturation,
  pairwise.display = "all",
  ylab = "SvO2",
  results.subtitle = FALSE,
  title = "",
  subtitle = "Superior Sagittal Sinus",
  type = "nonparametric",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid")
)

combine_plots(
  plotlist = list(m1, m2, m3, m4),
  plotgrid.args = list(nrow = 2)
)
```

Figure 3.2
*note: "****" in panels (a) and (b) were added outside of R
```{r, echo=FALSE}
p1 <- ggwithinstats(
  data = comp_legit_starfsl,
  x = Vein,
  y = Chi,
  pairwise.display = "all",
  results.subtitle = FALSE,
  xlab = "Region",
  ylab = "Chi",
  title = "(A)",
  type = "nonparametric",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#D95F02") 
)

p2 <- ggwithinstats(
  data = comp_SSS_IntVein_CSvO2,
  x = Vein,
  y = CSvO2,
  pairwise.display = "all",
  results.subtitle = FALSE,
  xlab = "Region",
  ylab = "SvO2",
  title = "(B)",
  type = "nonparametric",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#D95F02")
)

p3 <- ggwithinstats(                     #boxplot of chi comparison between regions
  data = comp_chisep_chi,
  x = Vein,
  y = Chi, 
  pairwise.display = "all",
  results.subtitle = FALSE,
  xlab = "Region",
  ylab = "Chi",
  title = "(C)",
  subtitle = "                                                n.s.",
  type = "nonparametric", 
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#1B9E77")
)

p4 <- ggwithinstats(                    #boxplot of svo2 comparison between regions
  data = comp_chisep_CSvO2,
  x = Vein,
  y = CSvO2, 
  pairwise.display = "all",
  results.subtitle = FALSE,
  xlab = "Region",
  ylab = "SvO2",
  title = "(D)",
  subtitle = "                                                n.s.",
  type = "nonparametric",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#1B9E77")
)

combine_plots(
  plotlist = list(p1, p2, p3, p4),
  plotgrid.args = list(nrow = 2)
)
```

```{r}
save.image('results.RData')
```

```{r}
 p1 <- ggwithinstats(
  data = comp_legit_starfsl,
  x = Vein,
  y = Chi,
  results.subtitle = FALSE,
  type = "p",
  pairwise.display = "all",
  bf.message = FALSE,
  xlab = "Region",
  ylab = "Chi",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#D95F02")) + 
  ggplot2::annotate("text", x=1.5, y=(max(comp_legit_starfsl$Chi)*1.05), size=4, label = "***")

p2 <- ggwithinstats(
  data = comp_SSS_IntVein_CSvO2,
  x = Vein,
  y = CSvO2,
  results.subtitle = FALSE,
  type = "p",
  pairwise.display = "all",
  bf.message = FALSE,
  xlab = "Region",
  ylab = "SvO2",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#D95F02")
) + 
  ggplot2::annotate("text", x=1.5, y=(max(comp_SSS_IntVein_CSvO2$CSvO2)*1.05), size=4, label = "***")

p3 <- ggwithinstats(                     #boxplot of chi comparison between regions
  data = comp_chisep_chi,
  x = Vein,
  y = Chi, 
  results.subtitle = FALSE,
  pairwise.display = "all",
  bf.message = FALSE,
  xlab = "Region",
  ylab = "Chi",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#1B9E77")
) +
  ggplot2::annotate("text", x=1.5, y=(max(comp_chisep_chi$Chi))*1.05, size=4, label = "n.s.")

p4 <- ggwithinstats(                    #boxplot of svo2 comparison between regions
  data = comp_chisep_CSvO2,
  x = Vein,
  y = CSvO2, 
  results.subtitle = FALSE,
  pairwise.display = "all",
  bf.message = FALSE,
  xlab = "Region",
  ylab = "SvO2",
  type = "nonparametric",
  centrality.plotting = FALSE,
  point.path.args = list(linewidth = 0.25, linetype = "solid"),
  point.args = list(color = "#1B9E77")
) +
  ggplot2::annotate("text", x=1.5, y=(max(comp_chisep_CSvO2$CSvO2)*1.05), size=4, label = "n.s.")
```

```{r regionplot, fig.width=8, fig.height=5, fig.cap="rmarkdown test figure"}
(patch1 <- (p1 + p2)/(p3 + p4))
ggsave('figure.png', patch1)
```

### 2024-05-31 notebooks/gavin_thesis_markdown.qmd
May31st, working version

--- 2024-05-31 notebooks/gavin_thesis_markdown.Rmd
@@
 
-```{r setup, include=FALSE}
+```{r}
+#| label = "setup",
+#| include = FALSE
 knitr::opts_chunk$set(echo = TRUE)
@@
 Dataframes for figures
-```{r, echo=FALSE}
+```{r}
+#| echo = FALSE
 
@@
 *note: "****" in panels were added outside of R
-```{r, echo=FALSE}
+```{r}
+#| echo = FALSE
 m1 <- ggwithinstats(
@@
 *note: "****" in panels (a) and (b) were added outside of R
-```{r, echo=FALSE}
+```{r}
+#| echo = FALSE
 p1 <- ggwithinstats(
@@
   ggplot2::annotate("text", x=1.5, y=(max(comp_chisep_CSvO2$CSvO2)*1.05), size=4, label = "n.s.")
+
+(patch1 <- (p1 + p2)/(p3 + p4))
+# patch1
+# ggsave('figure.png', patch1)
 ```
 
-```{r regionplot, fig.width=8, fig.height=5, fig.cap="rmarkdown test figure"}
-(patch1 <- (p1 + p2)/(p3 + p4))
-ggsave('figure.png', patch1)
+```{r}
+#| label: fig-regionplot
+#| fig-width: 8
+#| fig-height: 5
+#| fig-cap: "rmarkdown test figure"
+patch1
 ```

### 2024-06-11 index.qmd
June 10th

--- 2024-05-24 index.qmd
@@
 ---
-title: "ParaChiCSVO2"
+title: "The Application of Magnetic Susceptibility Separation for Measuring Cerebral Oxygenation in Preterm Neonates"
 titlerunning: "CSVO2"
@@
 The acquired imaging data were processed as QSM images to obtain the susceptibility values of the superior sagittal sinus (SSS) and central cerebral veins (CCV).
-These values were used to calculate the oxygen saturation (SvO2) of the SSS and CCV.
+These values were used to calculate the oxygen saturation (SvO~2~) of the SSS and CCV.
 QSM images were subsequently processed to isolate their paramagnetic components.
-SvO2 values of the SSS and CCV were calculated again from the paramagnetic components.
+SvO~2~ values of the SSS and CCV were calculated again from the paramagnetic components.
 
-  **Results**: The mean SvO2 values of the SSS and CCV calculated from QSM images were found to be 72.4% (SD, 3.4%) and 68.7% (SD, 3.5%), respectively.
-The mean SvO2 values calculated from paramagnetic components were found to be 58.1% (SD, 7.3%) for the SSS and 57.7% (SD, 7.0%) for the CCV.
+  **Results**: The mean SvO~2~ values of the SSS and CCV calculated from QSM images were found to be 72.4% (SD, 3.4%) and 68.7% (SD, 3.5%), respectively.
+The mean SvO~2~ values calculated from paramagnetic components were found to be 58.1% (SD, 7.3%) for the SSS and 57.7% (SD, 7.0%) for the CCV.
 
-  **Conclusion**: SSS SvO2 values derived from paramagnetic components agreed well with the existing literature and were closer than the values derived from QSM, however, they displayed greater variability.
-Although the CCV SvO2 data from QSM aligns more closely with existing literature, it is important to note that the current literature on this topic remains relatively limited in the CCV.
+  **Conclusion**: SSS SvO~2~ values derived from paramagnetic components agreed well with the existing literature and were closer than the values derived from QSM, however, they displayed greater variability.
+Although the CCV SvO~2~ data from QSM aligns more closely with existing literature, it is important to note that the current literature on this topic remains relatively limited in the CCV.
 Thus, decomposing QSM images into paramagnetic components shows great promise as a method for more accurately measuring cerebral oxygenation in neonates but may require more research to improve precision.
@@
 
+```{r}
+load("notebooks/results.RData")
+```
+
+```{r}
+rndmean <- function(clm) {
+  return(round(mean(clm),2))
+}
+rndsd <- function(clm) {
+  return(round(sd(clm),2))
+}
+```
+
 # Introduction {#sec-intro}
@@
 <!-- should be around 5-6 paragraphs. Aim for 460 words -->
+
+Preterm birth
+
+Abnormal brain development is a significant concern for parents with children born preterm, as 43% of infants that survive will have neurodevelopmental delays later in life [@dixMonitoringCerebralOxygenation2017].
+Irregularities in early cerebral oxygen levels have been identified as a potential source of such delays, where too little oxygen provided during NICU care can result in white matter injury, while too much oxygen can result in reduced cortical connectivity [@rantakariEarlyOxygenLevels2021].
+As such, being able to precisely, accurately, and non-invasively measure cerebral oxygenation is necessary for understanding and improving neurodevelopmental outcomes in preterm neonates.
+
+In the present study, we set out to determine whether a QSM image alone, or the paramagnetic component of the QSM image, is more accurate in measuring the oxygen present in the major cerebral veins of a cohort of preterm neonates.
+As previous QSM studies have not included the SSS, we also had a secondary aim of preserving this vessel in our QSM images and using this data to determine whether a difference in oxygenation existed between the SSS and the central cerebral veins (CCV).
 
@@
 
-|                          | T1w          | T2w          | pcASL                                                               | SWI                                               |
-|------------|------------|------------|----------------------|---------------|
-| Sequence                 | 3D FSPGR     | 3D CUBE      | Multi-shot 3D fast spin-echo                                        | 3D spoiled GRE flow-compensated                   |
-| Phase-encoding direction | Coronal      | Sagittal     | Axial                                                               | Axial                                             |
-| TR (ms)                  | 7.74         | 2,300        | 4,680                                                               | 30.9                                              |
-| TE (ms)                  | 2.97         | 66.29        | 10.55                                                               | 5.24                                              |
-| Flip angle               | 12$^{\circ}$ | 90$^{\circ}$ | 111$^{\circ}$                                                       | 20$^{\circ}$                                      |
-| FOV (cm)                 | 20           | 20           | 24                                                                  | 25                                                |
-| Acquisition Matrix       | 512 x 512    | 256 x 256    | 128 x 128                                                           | 256 x 256                                         |
-| In-plane resolution (mm) | 0.39 x 0.39  | 0.78 x 0.78  | 1.875 x 1.875                                                       | 0.977 x 0.977                                     |
-| Slice thickness (mm)     | 1            | 1            | 4                                                                   | 2, reconstructed to 1 mm with zero filling (ZIP2) |
-| Number of slices         | 126          | 106          | 50                                                                  | 92                                                |
-| Additional parameters    | n/a          | n/a          | 1,450 ms label period; 2,025 ms pulse lable; 24 control-label pairs | n/a                                               |
-| Scan duration            | 4 min 39 s   | 5 min 1 s    | 5 min 26 s                                                          | 5 min 29 s                                        |
+::: {#tbl-mri}
+```{=latex}
+    \begin{tabular*}{\textwidth}{@{}lllll@{}}
+    \toprule
+        ~ & T1w & T2w & pcASL & SWI \\ \midrule
+        Sequence & 3D FSPGR & 3D CUBE & \makecell[tl]{Multi-shot 3D fast \\ spin-echo} & \makecell[tl]{3D spoiled GRE \\ flow-compensated} \\
+        \makecell[tl]{Phase-encoding \\ direction} & Coronal & Sagittal & Axial & Axial \\
+        TR (ms) & 7.74 & 2,300 & 4,680 & 30.9 \\
+        TE (ms) & 2.97 & 66.29 & 10.55 & 5.24 \\
+        Flip angle & 12$^{\circ}$ & 90$^{\circ}$ & 111$^{\circ}$ & 20$^{\circ}$ \\
+        FOV (cm) & 20 & 20 & 24 & 25 \\
+        Acquisition matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256 \\
+        \makecell[tl]{In-plane resolution \\ (mm)} & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 & 0.977 x 0.977 \\
+        \makecell[tl]{Slice thickness \\ (mm)} & 1 & 1 & 4 & \makecell[tl]{2, reconstructed \\ to 1 with zero \\ filling (ZIP2)} \\
+        Number of slices & 126 & 106 & 50 & 92 \\
+        Additional parameters & n/a & n/a & \makecell[tl]{1,450 ms label period; \\ 2,025 ms pulse label; \\ 24 control-label pairs} & n/a \\
+        Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s \\ \botrule
+    \end{tabular*}
+    \footnotetext{T1w = T1-weighted; T2w = T2-weighted; pcASL = pseudo-continuous arterial spin labelling; SWI = susceptibility weighted imaging; FSPG = fast spoiled gradient echo; CUBE = General Electric name of sequence, not an acronym; GRE = gradient echo; ZIP2 = through-plane zero filling interpolation}
+```
+Technical parameters for MR imaging pulse sequences
+:::
 
-: Technical parameters for MR imaging protocol. {#tbl-mri}
+The MRI scan protocol comprised of the following sequences: a T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin labeling (ASL) scan, a multi-echo susceptibility-weighted imaging scan, and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging (EPI) sequence.
+The DWI sequence was not used for the present study.
 
-The MRI scan protocol comprised several sequences, including a T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin labeling (ASL) scan, a multi-echo susceptibility-weighted imaging scan, and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging (EPI) sequence.
+## Image analysis
+
+The raw DICOM files acquired from the scanning procedure were converted to NIfTI (Neuroimaging Informatics Technology Initiative) format using Chris Rorden’s `dcmniix` tool.
+SWI magnitude data files were then used to create subject-specific brain masks that would not erode the SSS during QSM processing, an issue faced by our group in the past [@liFirstStepNeuroimaging2016].
+A step-by-step summary of the pipeline used is shown in @fig-sample.
+
+{{< embed notebooks/Figures.ipynb#fig-graph >}}
+
+First, the fifth echo SWI magnitude file was processed using FSL’s (v.
+6.0.7.3) [@woolrichBayesianAnalysisNeuroimaging2009] `fslroi`, `fslmaths`, and `bet` [@smithFastRobustAutomated2002] to create a preliminary brain mask, similar to our previous efforts, which does not contain the SSS.
+`Fslroi` was used to isolate the fifth echo of the magnitude data, which was then squared using Fslmaths and the option `-sqr`.
+Squaring the magnitude image was found to dramatically improve subsequent brain extraction.
+The resulting image was then used to create the preliminary brain mask using bet with the options `-m` and `-R`.
+The former flag generated a binary brain mask, while the latter performed a more robust brain centre estimation.
+The brain mask was then dilated by 7 voxels using `Fslmaths` and the options `-kernel boxv` and `-dilM` in order for the dilated mask to contain the SSS (along with unwanted tissue as well).
+This mask was then used, along with the phase images, in a MATLAB script for QSM calculation from Christian Kames [@kamesRapidTwostepDipole2018] to produce a preliminary QSM image that contained the SSS, albeit with fairly low signal-to-noise ratio and other unwanted tissue.
+Given the high contrast in voxel intensity between the SSS and surrounding tissue, the select by intensity tool in `FSLeyes` [@mccarthyFSLeyes2023] was then used to segment the SSS from the QSM image and create a 3D mask of the selected region.
+Using `fslmaths` and the options `-add` and `-bin`, the SSS mask was then combined with the original brain mask of the fifth echo.
+This resulted in a brain mask that contained only brain and SSS signal.
+Finally, this mask was used in a final QSM post-processing step to create a QSM image that includes the SSS while maintaining a high signal-to-noise ratio, making it suitable to obtain accurate susceptibility values.
+
+STI Suite (v.
+3.0) [@liIntegratedLaplacianBased2014], was used to process the final QSM images as it produced the cleanest images without eroding the SSS.
+The finalized brain mask and the last three echoes of the magnitude and phase images were used in STI Suite along with the following parameters: 0.9766 x 0.9766 x 1 mm^3^ voxel size, 5 ms TE1, 5.3 ms $\Delta$TE, and 77.4 ms sum TE, B0 strength = 3, and B0 direction = \[0, 0, 1\].
+The 3D GRE data option was selected for the phase processing stage, and STAR-QSM was selected for the QSM stage.
+Finally, the \`select by intensity' tool in `FSLeyes` was then used to semi-automatically make vascular masks of the SSS and CCV from each subject’s QSM image.
+The vascular masks were used to calculate the mean susceptibility of each subject’s SSS and CCV from their QSM image with `fslstats`.
+
+To isolate the paramagnetic component of subjects’ QSM data, the $\chi$-separation toolbox [@shinHseparationMagneticSusceptibility2021] from the Laboratory for Imaging Science and Technology was used.
+Each subject’s magnitude and phase SWI data were used along with the following parameters: 0.9766 x 0.9766 x 1 mm^3^ voxel size; TE (s) = \[0.005, 0.0102, 0.0155, 0.0207, 0.026\]; delta TE (s) = 0.0052; B0 strength = 3; B0 direction = \[0, 0, 1\].
+The mean susceptibility of each subject’s SSS and CCV in their paramagnetic maps was calculated with the same vascular masks used for the QSM images.
+Sample images showing the magnitude, final QSM, and final paramagnetic component images are shown in @fig-sample.
 
@@
 
-See @fig-graph
+Once the mean susceptibility values of the SSS and CCV were obtained from the subjects’ QSM images and paramagnetic maps, venous oxygen saturation (SvO~2~) was calculated with the following equation [@bergInvestigatingEffectFlow2021]:
 
-{{< embed notebooks/Figures.ipynb#fig-graph >}}
+$$
+SvO_{2} = 1 - \frac{\Delta \chi _{blood} - (\Delta \chi _{oxy} * Hct)}{\Delta \chi _{do} * Hct}
+$$ {#eq-svo}
+
+where $\Delta \chi _{blood}$ is the vessel’s measured susceptibility, $\Delta \chi _{oxy}$ is the constant representing the susceptibility changes of oxygenated blood relation to water, $\Delta \chi _{do}$ is the difference in susceptibility between oxygenated and deoxygenated blood, and Hct is the subject’s hematocrit.
+$\Delta \chi _{oxy}$ was -0.21 \* 4$\pi$ ppm as per @portnoyHumanUmbilicalCord2018 and [@sedlacikObtainingBloodOxygenation2007], while $\Delta \chi _{do}$ was -0.03 \* 4$\pi$ ppm as per [@weisskoffMRISusceptometryImagebased1992].
+Subjects’ Hct for the day of the scan was calculated using a four-parameter Weibull function with previously measured values while still in the NICU.
+
+## Statistical analysis
+
+Statistical analysis of the acquired data was performed using R and RStudio (v.
+2023.09.1 Build 494) [@rcoreteamLanguageEnvironmentStatistical2022; @rstudioteamRStudioIntegratedDevelopment].
+Mean and standard deviation values are reported for most statistics, unless specified otherwise.
+A paired Student's t-test was used to determine statistical significance (p \<0.05) between two parameters (e.g.
+$\chi$ values between venous structures).
+
+<!-- the vessel-specific SvO2 values determined through QSM and those determined from the paramagnetic maps processed with $\chi$-separation.
+This was also performed for the raw $\chi$ values obtained through both methods.
+-->
 
@@
 
-[@ahmedDiamagneticComponentMap2023a]
+A total sample size of `r length(newdata$Subject)` infants were scanned, with a mean ($\pm$ standard deviation) gestational age of `r round(mean(newdata$GA),2)` $\pm$ `r round(sd(newdata$GA),2)` weeks and a mean birth weight of `r round(mean(newdata$BW),2)` $\pm$ `r round(sd(newdata$BW),2)` grams.
+A comprehensive summary of neonatal characteristics, including additional demographic and clinical data, is provided in @tbl-dem for reference.
+
+::: {#tbl-dem}
+```{=latex}
+\begin{tabular}{@{}ll@{}}
+    \toprule
+Variable & Subject Data (n=19) \\
+  \midrule
+Gestational age, weeks (mean $\pm$ SD)                       & `r round(mean(newdata$GA),2)` $\pm$ `r round(sd(newdata$GA),2)`                                                     \\
+Corrected gestational age on scan day, weeks (mean $\pm$ SD) & `r round(mean(newdata$CGAonScanDay),2)` $\pm$ `r round(sd(newdata$CGAonScanDay),2)`                                 \\
+Number of male neonates (%)                                  & `r sum(newdata$Sex == "M")` (`r round(sum(newdata$Sex == "M")/length(newdata$Sex)*100,2)`)                          \\
+Birth weight, g (mean $\pm$ SD)                              & `r round(mean(newdata$BW),2)` $\pm$ `r round(sd(newdata$BW),2)`                                                     \\
+Weight on scan day, g (mean $\pm$ SD)                        & `r round(mean(newdata$Weight.on.Scan.Day..kg.*1000),2)` $\pm$ `r round(sd(newdata$Weight.on.Scan.Day..kg.*1000),2)` \\
+Days spent in NICU (median, IQR)                             & `r median(newdata$Total_Days_NICU)`, `r IQR(newdata$Total_Days_NICU)`                                               \\
+Days on ventilation (median, IQR)                            & `r median(newdata$Total_Days_Ventilation)`, `r IQR(newdata$Total_Days_Ventilation)`                                 \\
+\botrule
+\end{tabular}
+\footnotetext{SD = standard deviation; IQR = inter quartile range}
+```
+Demographic and clinical characteristic of the study sample.
+:::
+
+The mean SvO~2~ values for the SSS and the CCV were found to be `r rndmean(newdata$Gavin_SSSVein_CSvO2)` $\pm$ `r rndsd(newdata$Gavin_SSSVein_CSvO2)` and `r rndmean(newdata$Gavin_IntVein_CSvO2)` $\pm$  `r rndsd(newdata$Gavin_IntVein_CSvO2)` ppm, respectively, when determined from the QSM data.
+When determined from the paramagnetic map, the mean SvO~2~ values for the SSS and the CCV were found to be `r rndmean(newdata$Chisep_SSSVein_CSvO2)` $\pm$ `r rndsd(newdata$Chisep_IntVein_CSvO2)` %, respectively.
+A summary of the measured physiological parameters, including the chi values used to calculate SvO~2~, can found in @tbl-chistats.
+
+::: {#tbl-chistats}
+```{=latex}
+    \begin{tabular}{@{}llllll@{}}
+    \toprule
+        Region & Measure & QSM & \makecell[tl]{Paramagnetic \\ map} & p-value & 95\% CI \\ \midrule
+        \multirow{ 2}{*}{SSS} & Chi (ppm)     & `r round(mean(newdata$Gavin_SSS_Chi),2)` \pm `r round(sd(newdata$Gavin_SSS_Chi),2)`                     & `r round(mean(newdata$Chisep_SSS_Chi),2)` \pm `r round(sd(newdata$Chisep_SSS_Chi),2)`                     & `r signif(SSSChi.t$p.value,3)`                        & `r round(SSSChi.t$conf.low,2)`, `r round(SSSChi.t$conf.high,2)`           \\
+                              & SvO$_{2}$(\%) & `r round(mean(newdata$Gavin_SSSVein_CSvO2*100),2)` \pm `r round(sd(newdata$Gavin_SSSVein_CSvO2*100),2)` & `r round(mean(newdata$Chisep_SSSVein_CSvO2*100),2)` \pm `r round(sd(newdata$Chisep_SSSVein_CSvO2*100),2)` & `r signif(SSSSvO2.t$p.value,3)` & `r round(SSSSvO2.t$conf.low*100,2)`, `r round(SSSSvO2.t$conf.high*100,2)` \\
+        \multirow{ 2}{*}{CCV} & Chi (ppm)     & `r round(mean(newdata$Gavin_IntVein_Chi),2)` \pm `r round(sd(newdata$Gavin_IntVein_Chi),2)`             & `r round(mean(newdata$Chisep_IntVein_Chi),2)` \pm `r round(sd(newdata$Chisep_IntVein_Chi),2)`             & `r signif(CCVChi.t$p.value,3)`                        & `r round(CCVChi.t$conf.low,2)`, `r round(CCVChi.t$conf.high,2)`           \\
+                              & SvO$_{2}$(\%) & `r round(mean(newdata$Gavin_IntVein_CSvO2*100),2)` \pm `r round(sd(newdata$Gavin_IntVein_CSvO2*100),2)` & `r round(mean(newdata$Chisep_IntVein_CSvO2*100),2)` \pm `r round(sd(newdata$Chisep_IntVein_CSvO2*100),2)` & `r signif(CCVSvO2.t$p.value,3)` & `r round(CCVSvO2.t$conf.low*100,2)`, `r round(CCVSvO2.t$conf.high*100,2)` \\
+        \botrule
+    \end{tabular}
+    \footnotetext{QSM = quantitative susceptibility mapping; CI = confidence interval; SSS = superior sagitall sinus; CCV = central cerebral vein}
+```
+Summary of acquired physiological parameters.
+Mean ± SD is shown for chi and SvO~2~ values.
+The P-value and 95% confidence interval (CI) were obtained through the comparison of values between QSM and paramagnetic maps; (n=19).
+:::
+
+Region-specific $\chi$ and SvO~2~ values acquired from QSM were compared to values acquired from paramagnetic maps.
+In both the SSS and CCV, it was found that a significant difference existed between values acquired ($\chi$ and SvO~2~) from QSM and paramagnetic maps (p < 0.05).
+A boxplot showing the comparisons made is shown in @fig-methodplot.
+
+{{< embed notebooks/gavin_thesis_markdown.qmd#fig-methodplot >}}
+
+The acquired $\chi$ and SvO~2~ values were additionally compared between veins.
+In data created from QSM, a significant difference was found between the CCV and SSS in mean $\chi$ (p < 0.05; 95% CI [`r round(RegionQSMChi.t$conf.low,3)`, `r round(RegionQSMChi.t$conf.high,3)`]) and mean SvO~2~ (p < 0.05; 95% CI [`r round(RegionQSMSvO2.t$conf.low,3)`, `r round(RegionQSMSvO2.t$conf.high,3)`]).
+In data acquired from paramagnetic maps, no significant difference was observed between the CCV and the SSS in either mean $\chi$ (p = `r round(RegionChisepChi.t$p.value,3)`; 95% CI [`r round(RegionChisepChi.t$conf.low,3)`, `r round(RegionChisepChi.t$conf.high,3)`]) or mean SvO~2~ (p = `r round(RegionChisepSvO2.t$p.value,3)`; 95% CI [`r round(RegionChisepSvO2.t$conf.low,3)`, `r round(RegionChisepChi.t$conf.high,3)`]).
+A summary of these comparisons is represented in @fig-regionplot.
+
+{{< embed notebooks/gavin_thesis_markdown.qmd#fig-regionplot >}}
+
+# Discussion {#sec-discussion}
+
+The primary objective of the present study was to assess whether the application of magnetic susceptibility separation to neonatal QSM data could provide more accurate cerebral venous vessel oxygenation measurements.
+To the best of our knowledge, we are the first to test this in a neonatal cohort, as susceptibility separation has been typically evaluated as a method of imaging myelin and brain iron in adult subjects [@shinHseparationMagneticSusceptibility2021; @ahmedDiamagneticComponentMap2023a].
+Our results showed that the SvO~2~ values of the SSS and CCV obtained from susceptibility separation are significantly lower than the respective SvO~2~ values obtained from QSM alone.
+When our results were compared to the literature, we found that our SSS SvO~2~ data from susceptibility separation agreed well with the findings of other studies measuring SvO~2~ of the SSS in similar subject populations.
+Conversely, the paramagnetic CCV SvO~2~ data saw less agreement with the existing literature than the corresponding data from QSM.
+However, there is reason to believe our paramagnetic CCV values may be accurate given their similarity to the paramagnetic SSS values and the limitations of the two studies that observed CCV SvO~2~.
+Additionally, it is important to note that our SvO~2~ measurements from susceptibility separation had greater variance than our measurements from QSM, indicating a limitation that should be addressed in future research.
+Overall, the present work demonstrates the promise of susceptibility separation as an MRI post-processing technique that can measure the oxygenation of the cerebral veins of infant subjects, a useful marker of regional oxygen consumption in the brain.
 

### 2024-06-11 _tex/index.tex
Built site for gh-pages

@@
 \setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{5}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
 % Make \paragraph and \subparagraph free-standing
@@
 \hypersetup{
-  pdftitle={ParaChiCSVO2},
+  pdftitle={The Application of Magnetic Susceptibility Separation for Measuring Cerebral Oxygenation in Preterm Neonates},
   pdfauthor={Thomas Gavin Carmichael; Alexander Rauscher; Ruth E Grunau; Alexander Mark Weber},
@@
 
-\title[ParaChiCSVO2]{ParaChiCSVO2}
+\title[The Application of Magnetic Susceptibility Separation for
+Measuring Cerebral Oxygenation in Preterm Neonates]{The Application of
+Magnetic Susceptibility Separation for Measuring Cerebral Oxygenation in
+Preterm Neonates}
 
@@
 superior sagittal sinus (SSS) and central cerebral veins (CCV). These
-values were used to calculate the oxygen saturation (SvO2) of the SSS
-and CCV. QSM images were subsequently processed to isolate their
-paramagnetic components. SvO2 values of the SSS and CCV were calculated
-again from the paramagnetic components.
+values were used to calculate the oxygen saturation
+(SvO\textsubscript{2}) of the SSS and CCV. QSM images were subsequently
+processed to isolate their paramagnetic components. SvO\textsubscript{2}
+values of the SSS and CCV were calculated again from the paramagnetic
+components.
 
-\textbf{Results}: The mean SvO2 values of the SSS and CCV calculated
-from QSM images were found to be 72.4\% (SD, 3.4\%) and 68.7\% (SD,
-3.5\%), respectively. The mean SvO2 values calculated from paramagnetic
-components were found to be 58.1\% (SD, 7.3\%) for the SSS and 57.7\%
-(SD, 7.0\%) for the CCV.
+\textbf{Results}: The mean SvO\textsubscript{2} values of the SSS and
+CCV calculated from QSM images were found to be 72.4\% (SD, 3.4\%) and
+68.7\% (SD, 3.5\%), respectively. The mean SvO\textsubscript{2} values
+calculated from paramagnetic components were found to be 58.1\% (SD,
+7.3\%) for the SSS and 57.7\% (SD, 7.0\%) for the CCV.
 
-\textbf{Conclusion}: SSS SvO2 values derived from paramagnetic
-components agreed well with the existing literature and were closer than
-the values derived from QSM, however, they displayed greater
-variability. Although the CCV SvO2 data from QSM aligns more closely
-with existing literature, it is important to note that the current
-literature on this topic remains relatively limited in the CCV. Thus,
-decomposing QSM images into paramagnetic components shows great promise
-as a method for more accurately measuring cerebral oxygenation in
-neonates but may require more research to improve precision. Notably, no
-significant difference in oxygenation was observed between the CCV and
-the SSS, contrasting with previous studies.}
+\textbf{Conclusion}: SSS SvO\textsubscript{2} values derived from
+paramagnetic components agreed well with the existing literature and
+were closer than the values derived from QSM, however, they displayed
+greater variability. Although the CCV SvO\textsubscript{2} data from QSM
+aligns more closely with existing literature, it is important to note
+that the current literature on this topic remains relatively limited in
+the CCV. Thus, decomposing QSM images into paramagnetic components shows
+great promise as a method for more accurately measuring cerebral
+oxygenation in neonates but may require more research to improve
+precision. Notably, no significant difference in oxygenation was
+observed between the CCV and the SSS, contrasting with previous
+studies.}
 
@@
 
-\textsubscript{Source:
-\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
-Notebook}}
+\section{Introduction}\label{sec-intro}
 
-\section{Introduction}\label{sec-intro}
+Preterm birth
+
+Abnormal brain development is a significant concern for parents with
+children born preterm, as 43\% of infants that survive will have
+neurodevelopmental delays later in life (Dix, van Bel, and Lemmers
+2017). Irregularities in early cerebral oxygen levels have been
+identified as a potential source of such delays, where too little oxygen
+provided during NICU care can result in white matter injury, while too
+much oxygen can result in reduced cortical connectivity (Rantakari et
+al. 2021). As such, being able to precisely, accurately, and
+non-invasively measure cerebral oxygenation is necessary for
+understanding and improving neurodevelopmental outcomes in preterm
+neonates.
+
+In the present study, we set out to determine whether a QSM image alone,
+or the paramagnetic component of the QSM image, is more accurate in
+measuring the oxygen present in the major cerebral veins of a cohort of
+preterm neonates. As previous QSM studies have not included the SSS, we
+also had a secondary aim of preserving this vessel in our QSM images and
+using this data to determine whether a difference in oxygenation existed
+between the SSS and the central cerebral veins (CCV).
 
@@
 
-Participant data comes from a previous study (Zhu et al. UNDER REVIEW
-May 2024). Participants consisted of preterm neonates born between 25-
-and 31-weeks gestational age (GA) who were admitted to the level III
-NICU at BC Women's Hospital. Recruitment took place over a span of one
-year, from February 2021 to January 2022, facilitated by a dedicated
-research nurse. Parents of eligible infants were approached by the
-research nurse prior to discharge from the NICU to explain the study
-objectives and seek their consent for participation. Infants meeting the
-criteria for inclusion were scanned for the study if they had already
-been discharged from the NICU, were in stable condition, and had reached
-a term equivalent age of 37 to 44 weeks GA. However, certain exclusion
-criteria were applied to ensure the homogeneity and integrity of the
-study sample. Infants were excluded if there was clinical evidence of a
+Participant data comes from a previous study (Zhu et al. 2024).
+Participants consisted of preterm neonates born between 25- and 31-weeks
+gestational age (GA) who were admitted to the level III NICU at BC
+Women's Hospital. Recruitment took place over a span of one year, from
+February 2021 to January 2022, facilitated by a dedicated research
+nurse. Parents of eligible infants were approached by the research nurse
+prior to discharge from the NICU to explain the study objectives and
+seek their consent for participation. Infants meeting the criteria for
+inclusion were scanned for the study if they had already been discharged
+from the NICU, were in stable condition, and had reached a term
+equivalent age of 37 to 44 weeks GA. However, certain exclusion criteria
+were applied to ensure the homogeneity and integrity of the study
+sample. Infants were excluded if there was clinical evidence of a
 congenital malformation or syndrome, a TORCH infection, or ultrasound
@@
 
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1644}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1644}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1644}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3014}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2055}}@{}}
-\caption{Technical parameters for MR imaging
-protocol.}\label{tbl-mri}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-T1w
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-T2w
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-pcASL
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-SWI
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-T1w
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-T2w
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-pcASL
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-SWI
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D fast spin-echo & 3D
-spoiled GRE flow-compensated \\
-Phase-encoding direction & Coronal & Sagittal & Axial & Axial \\
+\begin{table}
+
+\caption{\label{tbl-mri}Technical parameters for MR imaging pulse
+sequences}
+
+\centering{
+
+    \begin{tabular*}{\textwidth}{@{}lllll@{}}
+    \toprule
+        ~ & T1w & T2w & pcASL & SWI \\ \midrule
+        Sequence & 3D FSPGR & 3D CUBE & \makecell[tl]{Multi-shot 3D fast \\ spin-echo} & \makecell[tl]{3D spoiled GRE \\ flow-compensated} \\
+        \makecell[tl]{Phase-encoding \\ direction} & Coronal & Sagittal & Axial & Axial \\
         TR (ms) & 7.74 & 2,300 & 4,680 & 30.9 \\
         TE (ms) & 2.97 & 66.29 & 10.55 & 5.24 \\
-Flip angle & 12\(^{\circ}\) & 90\(^{\circ}\) & 111\(^{\circ}\) &
-20\(^{\circ}\) \\
+        Flip angle & 12$^{\circ}$ & 90$^{\circ}$ & 111$^{\circ}$ & 20$^{\circ}$ \\
         FOV (cm) & 20 & 20 & 24 & 25 \\
-Acquisition Matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256 \\
-In-plane resolution (mm) & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 &
-0.977 x 0.977 \\
-Slice thickness (mm) & 1 & 1 & 4 & 2, reconstructed to 1 mm with zero
-filling (ZIP2) \\
+        Acquisition matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256 \\
+        \makecell[tl]{In-plane resolution \\ (mm)} & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 & 0.977 x 0.977 \\
+        \makecell[tl]{Slice thickness \\ (mm)} & 1 & 1 & 4 & \makecell[tl]{2, reconstructed \\ to 1 with zero \\ filling (ZIP2)} \\
         Number of slices & 126 & 106 & 50 & 92 \\
-Additional parameters & n/a & n/a & 1,450 ms label period; 2,025 ms
-pulse lable; 24 control-label pairs & n/a \\
-Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s \\
-\end{longtable}
+        Additional parameters & n/a & n/a & \makecell[tl]{1,450 ms label period; \\ 2,025 ms pulse label; \\ 24 control-label pairs} & n/a \\
+        Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s \\ \botrule
+    \end{tabular*}
+    \footnotetext{T1w = T1-weighted; T2w = T2-weighted; pcASL = pseudo-continuous arterial spin labelling; SWI = susceptibility weighted imaging; FSPG = fast spoiled gradient echo; CUBE = General Electric name of sequence, not an acronym; GRE = gradient echo; ZIP2 = through-plane zero filling interpolation}
 
-The MRI scan protocol comprised several sequences, including a
+}
+
+\end{table}%
+
+The MRI scan protocol comprised of the following sequences: a
 T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin
@@
 and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging
-(EPI) sequence.
+(EPI) sequence. The DWI sequence was not used for the present study.
+
+\subsection{Image analysis}\label{image-analysis}
+
+The raw DICOM files acquired from the scanning procedure were converted
+to NIfTI (Neuroimaging Informatics Technology Initiative) format using
+Chris Rorden's \texttt{dcmniix} tool. SWI magnitude data files were then
+used to create subject-specific brain masks that would not erode the SSS
+during QSM processing, an issue faced by our group in the past (X. Li et
+al. 2016). A step-by-step summary of the pipeline used is shown in
+Figure~\ref{fig-sample}.
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics{index_files/figure-latex/notebooks-Figures-fig-graph-output-1.png}
+
+}
+
+\caption{\label{fig-graph}Pipeline for generating subject-specific brain
+masks that include the superior sagittal sinus (SSS). Initial steps
+involved (1) creating brain a mask from the magnitude of the fifth echo
+of the susceptibility weighted scan. Subsequently, the brain mask is
+dilated and then (2) utilized in conjunction with a quantitative
+susceptibility mapping (QSM) script to generate a preliminary QSM image.
+Further refinement involved (3) segmenting the SSS from the QSM image
+manually to create a tissue mask of the SSS region. Finally, (4) the
+vascular mask of the SSS is integrated with the initial brain mask,
+forming the comprehensive brain mask essential for obtaining
+susceptibility data that includes the SSS.}
+
+\end{figure}%
+
+First, the fifth echo SWI magnitude file was processed using FSL's (v.
+6.0.7.3) (Woolrich et al. 2009) \texttt{fslroi}, \texttt{fslmaths}, and
+\texttt{bet} (Smith 2002) to create a preliminary brain mask, similar to
+our previous efforts, which does not contain the SSS. \texttt{Fslroi}
+was used to isolate the fifth echo of the magnitude data, which was then
+squared using Fslmaths and the option \texttt{-sqr}. Squaring the
+magnitude image was found to dramatically improve subsequent brain
+extraction. The resulting image was then used to create the preliminary
+brain mask using bet with the options \texttt{-m} and \texttt{-R}. The
+former flag generated a binary brain mask, while the latter performed a
+more robust brain centre estimation. The brain mask was then dilated by
+7 voxels using \texttt{Fslmaths} and the options \texttt{-kernel\ boxv}
+and \texttt{-dilM} in order for the dilated mask to contain the SSS
+(along with unwanted tissue as well). This mask was then used, along
+with the phase images, in a MATLAB script for QSM calculation from
+Christian Kames (Kames, Wiggermann, and Rauscher 2018) to produce a
+preliminary QSM image that contained the SSS, albeit with fairly low
+signal-to-noise ratio and other unwanted tissue. Given the high contrast
+in voxel intensity between the SSS and surrounding tissue, the select by
+intensity tool in \texttt{FSLeyes} (McCarthy 2023) was then used to
+segment the SSS from the QSM image and create a 3D mask of the selected
+region. Using \texttt{fslmaths} and the options \texttt{-add} and
+\texttt{-bin}, the SSS mask was then combined with the original brain
+mask of the fifth echo. This resulted in a brain mask that contained
+only brain and SSS signal. Finally, this mask was used in a final QSM
+post-processing step to create a QSM image that includes the SSS while
+maintaining a high signal-to-noise ratio, making it suitable to obtain
+accurate susceptibility values.
+
+STI Suite (v. 3.0) (W. Li et al. 2014), was used to process the final
+QSM images as it produced the cleanest images without eroding the SSS.
+The finalized brain mask and the last three echoes of the magnitude and
+phase images were used in STI Suite along with the following parameters:
+0.9766 x 0.9766 x 1 mm\textsuperscript{3} voxel size, 5 ms TE1, 5.3 ms
+\(\Delta\)TE, and 77.4 ms sum TE, B0 strength = 3, and B0 direction =
+{[}0, 0, 1{]}. The 3D GRE data option was selected for the phase
+processing stage, and STAR-QSM was selected for the QSM stage. Finally,
+the `select by intensity' tool in \texttt{FSLeyes} was then used to
+semi-automatically make vascular masks of the SSS and CCV from each
+subject's QSM image. The vascular masks were used to calculate the mean
+susceptibility of each subject's SSS and CCV from their QSM image with
+\texttt{fslstats}.
+
+To isolate the paramagnetic component of subjects' QSM data, the
+\(\chi\)-separation toolbox (Shin et al. 2021) from the Laboratory for
+Imaging Science and Technology was used. Each subject's magnitude and
+phase SWI data were used along with the following parameters: 0.9766 x
+0.9766 x 1 mm\textsuperscript{3} voxel size; TE (s) = {[}0.005, 0.0102,
+0.0155, 0.0207, 0.026{]}; delta TE (s) = 0.0052; B0 strength = 3; B0
+direction = {[}0, 0, 1{]}. The mean susceptibility of each subject's SSS
+and CCV in their paramagnetic maps was calculated with the same vascular
+masks used for the QSM images. Sample images showing the magnitude,
+final QSM, and final paramagnetic component images are shown in
+Figure~\ref{fig-sample}.
 
@@
 
-\caption{\label{fig-sample}Sample images}
+\caption{\label{fig-sample}An example of subject imaging data. A sample
+sagittal, coronal, and axial slice is displayed for each image. (a) The
+1st echo of the magnitude susceptibility weighted imaging sequence; (b)
+the final quantitative susceptibility mapping image; and (c) the
+paramagnetic component isolated from the quantitative susceptibility
+map. The bars in (b) and (c) indicates the range of susceptibility
+\(\chi\) values.}
 
@@
 
-\textsubscript{Source:
-\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures.ipynb.html\#cell-fig-sample}{Python
-Figures}}
+Once the mean susceptibility values of the SSS and CCV were obtained
+from the subjects' QSM images and paramagnetic maps, venous oxygen
+saturation (SvO\textsubscript{2}) was calculated with the following
+equation (Berg et al. 2021):
 
-See Figure~\ref{fig-graph}
+\begin{equation}\phantomsection\label{eq-svo}{
+SvO_{2} = 1 - \frac{\Delta \chi _{blood} - (\Delta \chi _{oxy} * Hct)}{\Delta \chi _{do} * Hct}
+}\end{equation}
+
+where \(\Delta \chi _{blood}\) is the vessel's measured susceptibility,
+\(\Delta \chi _{oxy}\) is the constant representing the susceptibility
+changes of oxygenated blood relation to water, \(\Delta \chi _{do}\) is
+the difference in susceptibility between oxygenated and deoxygenated
+blood, and Hct is the subject's hematocrit. \(\Delta \chi _{oxy}\) was
+-0.21 * 4\(\pi\) ppm as per Portnoy et al. (2018) and (Sedlacik,
+Rauscher, and Reichenbach 2007), while \(\Delta \chi _{do}\) was -0.03 *
+4\(\pi\) ppm as per (Weisskoff and Kiihne 1992). Subjects' Hct for the
+day of the scan was calculated using a four-parameter Weibull function
+with previously measured values while still in the NICU.
+
+\subsection{Statistical analysis}\label{statistical-analysis}
+
+Statistical analysis of the acquired data was performed using R and
+RStudio (v. 2023.09.1 Build 494) (R Core Team 2022; RStudio Team, n.d.).
+Mean and standard deviation values are reported for most statistics,
+unless specified otherwise. A paired Student's t-test was used to
+determine statistical significance (p \textless0.05) between two
+parameters (e.g.~\(\chi\) values between venous structures).
+
+\section{Results}\label{sec-results}
+
+A total sample size of 19 infants were scanned, with a mean (\(\pm\)
+standard deviation) gestational age of 28.8 \(\pm\) 1.68 weeks and a
+mean birth weight of 1276.05 \(\pm\) 294.87 grams. A comprehensive
+summary of neonatal characteristics, including additional demographic
+and clinical data, is provided in Table~\ref{tbl-dem} for reference.
+
+\begin{table}
+
+\caption{\label{tbl-dem}Demographic and clinical characteristic of the
+study sample.}
+
+\centering{
+
+\begin{tabular}{@{}ll@{}}
+    \toprule
+Variable & Subject Data (n=19) \\
+  \midrule
+Gestational age, weeks (mean $\pm$ SD)                       & 28.8 $\pm$ 1.68                                                     \\
+Corrected gestational age on scan day, weeks (mean $\pm$ SD) & 40.36 $\pm$ 1.4                                 \\
+Number of male neonates (%)                                  & 10 (52.63)                          \\
+Birth weight, g (mean $\pm$ SD)                              & 1276.05 $\pm$ 294.87                                                     \\
+Weight on scan day, g (mean $\pm$ SD)                        & 3396.58 $\pm$ 597.72 \\
+Days spent in NICU (median, IQR)                             & 53, 23                                               \\
+Days on ventilation (median, IQR)                            & 31, 28.5                                 \\
+\botrule
+\end{tabular}
+\footnotetext{SD = standard deviation; IQR = inter quartile range}
+
+}
+
+\end{table}%
+
+The mean SvO\textsubscript{2} values for the SSS and the CCV were found
+to be 0.72 \(\pm\) 0.03 and 0.69 \(\pm\) 0.03 ppm, respectively, when
+determined from the QSM data. When determined from the paramagnetic map,
+the mean SvO\textsubscript{2} values for the SSS and the CCV were found
+to be 0.58 \(\pm\) 0.07 \%, respectively. A summary of the measured
+physiological parameters, including the chi values used to calculate
+SvO\textsubscript{2}, can found in Table~\ref{tbl-chistats}.
+
+\begin{table}
+
+\caption{\label{tbl-chistats}Summary of acquired physiological
+parameters. Mean ± SD is shown for chi and SvO\textsubscript{2} values.
+The P-value and 95\% confidence interval (CI) were obtained through the
+comparison of values between QSM and paramagnetic maps; (n=19).}
+
+\centering{
+
+    \begin{tabular}{@{}llllll@{}}
+    \toprule
+        Region & Measure & QSM & \makecell[tl]{Paramagnetic \\ map} & p-value & 95\% CI \\ \midrule
+        \multirow{ 2}{*}{SSS} & Chi (ppm)     & 0.1 \pm 0.02                     & 0.21 \pm 0.05                     & \ensuremath{2.84\times 10^{-11}}                        & -0.13, -0.09           \\
+                              & SvO$_{2}$(\%) & 72.46 \pm 3.43 & 58.14 \pm 7.3 & \ensuremath{6.12\times 10^{-10}} & 11.78, 16.85 \\
+        \multirow{ 2}{*}{CCV} & Chi (ppm)     & 0.13 \pm 0.02             & 0.22 \pm 0.05             & \ensuremath{6.25\times 10^{-9}}                        & -0.1, -0.07           \\
+                              & SvO$_{2}$(\%) & 68.71 \pm 3.46 & 57.69 \pm 6.97 & \ensuremath{2.16\times 10^{-9}} & 8.9, 13.13 \\
+        \botrule
+    \end{tabular}
+    \footnotetext{QSM = quantitative susceptibility mapping; CI = confidence interval; SSS = superior sagitall sinus; CCV = central cerebral vein}
+
+}
+
+\end{table}%
+
+Region-specific \(\chi\) and SvO\textsubscript{2} values acquired from
+QSM were compared to values acquired from paramagnetic maps. In both the
+SSS and CCV, it was found that a significant difference existed between
+values acquired (\(\chi\) and SvO\textsubscript{2}) from QSM and
+paramagnetic maps (p \textless{} 0.05). A boxplot showing the
+comparisons made is shown in Figure~\ref{fig-methodplot}.
 
@@
 
-\includegraphics{index_files/figure-latex/notebooks-Figures-fig-graph-output-1.png}
+\includegraphics{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-methodplot-output-1.png}
 
@@
 
-\caption{\label{fig-graph}A graph of the methods to create a mask}
+\caption{\label{fig-methodplot}Vein-specific susceptibility and oxygen
+saturation values by method. (A, B) contains violin plots comparing
+subject χ (ppm) acquired from QSM (A) and its paramagnetic component
+(B); (C, D) contains violin plots comparing subject SvO2 (\%) acquired
+from QSM (C) and its paramagnetic component (D). Raw data points from
+paramagnetic maps are shown as filled green circles and raw data points
+from QSM are shown as filled orange circles. Each line connects the raw
+data points of a single subject. (***) indicates P\textless0.05.}
 
@@
 
-\textsubscript{Source:
-\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures.ipynb.html\#cell-fig-graph}{Python
-Figures}}
+The acquired \(\chi\) and SvO\textsubscript{2} values were additionally
+compared between veins. In data created from QSM, a significant
+difference was found between the CCV and SSS in mean \(\chi\) (p
+\textless{} 0.05; 95\% CI {[}0.017, 0.04{]}) and mean
+SvO\textsubscript{2} (p \textless{} 0.05; 95\% CI {[}-0.052, -0.023{]}).
+In data acquired from paramagnetic maps, no significant difference was
+observed between the CCV and the SSS in either mean \(\chi\) (p = 0.711;
+95\% CI {[}-0.02, 0.029{]}) or mean SvO\textsubscript{2} (p = 0.752;
+95\% CI {[}-0.034, 0.029{]}). A summary of these comparisons is
+represented in Figure~\ref{fig-regionplot}.
 
-\section{Results}\label{sec-results}
+\begin{figure}[H]
 
-(Ahmed et al. 2023)
+\centering{
+
+\includegraphics{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-regionplot-output-1.png}
+
+}
+
+\caption{\label{fig-regionplot}Inter-venous comparisons of
+susceptibility and oxygen saturation. Violin plots comparing (A, C) χ
+(ppm) and (B, D) SvO2 (\%) between the CCV and the SSS. Panels (A) and
+(B) used data acquired from QSM, and its raw data points are shown as
+filled orange circles. Panels (C) and (D) used data acquired from
+paramagnetic maps, and its raw data points are shown as filled green
+circles. Each line connects the raw data points of a single subject.
+(***) indicates p\textless0.05; (n.s.) indicates no significant
+difference.}
+
+\end{figure}%
+
+\section{Discussion}\label{sec-discussion}
+
+The primary objective of the present study was to assess whether the
+application of magnetic susceptibility separation to neonatal QSM data
+could provide more accurate cerebral venous vessel oxygenation
+measurements. To the best of our knowledge, we are the first to test
+this in a neonatal cohort, as susceptibility separation has been
+typically evaluated as a method of imaging myelin and brain iron in
+adult subjects (Shin et al. 2021; Ahmed et al. 2023). Our results showed
+that the SvO\textsubscript{2} values of the SSS and CCV obtained from
+susceptibility separation are significantly lower than the respective
+SvO\textsubscript{2} values obtained from QSM alone. When our results
+were compared to the literature, we found that our SSS
+SvO\textsubscript{2} data from susceptibility separation agreed well
+with the findings of other studies measuring SvO\textsubscript{2} of the
+SSS in similar subject populations. Conversely, the paramagnetic CCV
+SvO\textsubscript{2} data saw less agreement with the existing
+literature than the corresponding data from QSM. However, there is
+reason to believe our paramagnetic CCV values may be accurate given
+their similarity to the paramagnetic SSS values and the limitations of
+the two studies that observed CCV SvO\textsubscript{2}. Additionally, it
+is important to note that our SvO\textsubscript{2} measurements from
+susceptibility separation had greater variance than our measurements
+from QSM, indicating a limitation that should be addressed in future
+research. Overall, the present work demonstrates the promise of
+susceptibility separation as an MRI post-processing technique that can
+measure the oxygenation of the cerebral veins of infant subjects, a
+useful marker of regional oxygen consumption in the brain.
 
@@
 
+\bibitem[\citeproctext]{ref-bergInvestigatingEffectFlow2021}
+Berg, Ronja C., Christine Preibisch, David L. Thomas, Karin Shmueli, and
+Emma Biondetti. 2021. {``Investigating the {Effect} of {Flow
+Compensation} and {Quantitative Susceptibility Mapping Method} on the
+{Accuracy} of {Venous Susceptibility Measurement}.''} bioRxiv.
+\url{https://doi.org/10.1101/2021.04.14.439812}.
+
+\bibitem[\citeproctext]{ref-dixMonitoringCerebralOxygenation2017}
+Dix, Laura Marie Louise, Frank van Bel, and Petra Maria Anna Lemmers.
+2017. {``Monitoring {Cerebral Oxygenation} in {Neonates}: {An
+Update}.''} \emph{Frontiers in Pediatrics} 5.
+
+\bibitem[\citeproctext]{ref-kamesRapidTwostepDipole2018}
+Kames, Christian, Vanessa Wiggermann, and Alexander Rauscher. 2018.
+{``Rapid Two-Step Dipole Inversion for Susceptibility Mapping with
+Sparsity Priors.''} \emph{NeuroImage} 167 (February): 276--83.
+\url{https://doi.org/10.1016/j.neuroimage.2017.11.018}.
+
+\bibitem[\citeproctext]{ref-liIntegratedLaplacianBased2014}
+Li, Wei, Alexandru V. Avram, Bing Wu, Xue Xiao, and Chunlei Liu. 2014.
+{``Integrated {Laplacian}-Based Phase Unwrapping and Background Phase
+Removal for Quantitative Susceptibility Mapping.''} \emph{NMR in
+Biomedicine} 27 (2): 219--27. \url{https://doi.org/10.1002/nbm.3056}.
+
+\bibitem[\citeproctext]{ref-liFirstStepNeuroimaging2016}
+Li, Xiangrui, Paul S. Morgan, John Ashburner, Jolinda Smith, and
+Christopher Rorden. 2016. {``The First Step for Neuroimaging Data
+Analysis: {DICOM} to {NIfTI} Conversion.''} \emph{Journal of
+Neuroscience Methods} 264 (March): 47--56.
+\url{https://doi.org/10.1016/j.jneumeth.2016.03.001}.
+
+\bibitem[\citeproctext]{ref-mccarthyFSLeyes2023}
+McCarthy, Paul. 2023. {``{FSLeyes}.''} Zenodo.
+\url{https://doi.org/10.5281/zenodo.8376979}.
+
+\bibitem[\citeproctext]{ref-portnoyHumanUmbilicalCord2018}
+Portnoy, Sharon, Natasha Milligan, Mike Seed, John G. Sled, and
+Christopher K. Macgowan. 2018. {``Human Umbilical Cord Blood Relaxation
+Times and Susceptibility at 3 {T}: {Human Umbilical Cord Blood
+Relaxation Times} and {Susceptibility} at 3 {T}.''} \emph{Magnetic
+Resonance in Medicine} 79 (6): 3194--3206.
+\url{https://doi.org/10.1002/mrm.26978}.
+
+\bibitem[\citeproctext]{ref-rcoreteamLanguageEnvironmentStatistical2022}
+R Core Team. 2022. {``R: {A Language} and {Environment} for {Statistical
+Computing}.''} Vienna, Austria: R Foundation for Statistical Computing.
+
+\bibitem[\citeproctext]{ref-rantakariEarlyOxygenLevels2021}
+Rantakari, Krista, Olli-Pekka Rinta-Koski, Marjo Metsäranta, Jaakko
+Hollmén, Simo Särkkä, Petri Rahkonen, Aulikki Lano, et al. 2021.
+{``Early Oxygen Levels Contribute to Brain Injury in Extremely Preterm
+Infants.''} \emph{Pediatric Research} 90 (1): 131--39.
+\url{https://doi.org/10.1038/s41390-021-01460-3}.
+
+\bibitem[\citeproctext]{ref-rstudioteamRStudioIntegratedDevelopment}
+RStudio Team. n.d. {``{RStudio}: {Integrated Development Environment}
+for {R}.''} Boston, MA: RStudio, PBC.
+
+\bibitem[\citeproctext]{ref-sedlacikObtainingBloodOxygenation2007}
+Sedlacik, Jan, Alexander Rauscher, and Jürgen R. Reichenbach. 2007.
+{``Obtaining Blood Oxygenation Levels from {MR} Signal Behavior in the
+Presence of Single Venous Vessels.''} \emph{Magnetic Resonance in
+Medicine} 58 (5): 1035--44. \url{https://doi.org/10.1002/mrm.21283}.
+
+\bibitem[\citeproctext]{ref-shinHseparationMagneticSusceptibility2021}
+Shin, Hyeong-Geol, Jingu Lee, Young Hyun Yun, Seong Ho Yoo, Jinhee Jang,
+Se-Hong Oh, Yoonho Nam, et al. 2021. {``{\(\chi\)}-Separation:
+{Magnetic} Susceptibility Source Separation Toward Iron and Myelin
+Mapping in the Brain.''} \emph{NeuroImage} 240 (October): 118371.
+\url{https://doi.org/10.1016/j.neuroimage.2021.118371}.
+
+\bibitem[\citeproctext]{ref-smithFastRobustAutomated2002}
+Smith, Stephen M. 2002. {``Fast Robust Automated Brain Extraction.''}
+\emph{Human Brain Mapping} 17 (3): 143--55.
+\url{https://doi.org/10.1002/hbm.10062}.
+
+\bibitem[\citeproctext]{ref-weisskoffMRISusceptometryImagebased1992}
+Weisskoff, Robert M., and Suzanne Kiihne. 1992. {``{MRI} Susceptometry:
+{Image-based} Measurement of Absolute Susceptibility of {MR} Contrast
+Agents and Human Blood: {COMMUNICATIONS}.''} \emph{Magnetic Resonance in
+Medicine} 24 (2): 375--83. \url{https://doi.org/10.1002/mrm.1910240219}.
+
+\bibitem[\citeproctext]{ref-woolrichBayesianAnalysisNeuroimaging2009}
+Woolrich, Mark W., Saad Jbabdi, Brian Patenaude, Michael Chappell,
+Salima Makni, Timothy Behrens, Christian Beckmann, Mark Jenkinson, and
+Stephen M. Smith. 2009. {``Bayesian Analysis of Neuroimaging Data in
+{FSL}.''} \emph{NeuroImage} 45 (1): S173--86.
+\url{https://doi.org/10.1016/j.neuroimage.2008.10.055}.
+
 \bibitem[\citeproctext]{ref-zhu-etal-cmro2}
 Zhu, A., C. Chau, N. Chan, A. Chacko, L. Holsti, R. E. Grunau, and A. M.
-Weber. UNDER REVIEW May 2024. {``Regional Cerebral Metabolic Rate of
-Oxygen and Levels of Respiratory Support in Preterm Neonates.''}
-\emph{Pediatric Research}, UNDER REVIEW May 2024.
+Weber. 2024. {``Regional Cerebral Metabolic Rate of Oxygen and Levels of
+Respiratory Support in Preterm Neonates.''} \emph{Pediatric Research},
+May.
 

### 2024-06-11 index.qmd
Changed tables to be more html friendly

@@
 ```{=latex}
-    \begin{tabular*}{\textwidth}{@{}lllll@{}}
+    \begin{tabular*}{\textwidth}{@{}p{2.5cm}llp{2.5cm}p{2.5cm}@{}}
     \toprule
         ~ & T1w & T2w & pcASL & SWI \\ \midrule
-        Sequence & 3D FSPGR & 3D CUBE & \makecell[tl]{Multi-shot 3D fast \\ spin-echo} & \makecell[tl]{3D spoiled GRE \\ flow-compensated} \\
-        \makecell[tl]{Phase-encoding \\ direction} & Coronal & Sagittal & Axial & Axial \\
+        Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D fast spin-echo & 3D spoiled GRE flow-compensated \\
+        Phase-encoding direction & Coronal & Sagittal & Axial & Axial \\
         TR (ms) & 7.74 & 2,300 & 4,680 & 30.9 \\
@@
         Acquisition matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256 \\
-        \makecell[tl]{In-plane resolution \\ (mm)} & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 & 0.977 x 0.977 \\
-        \makecell[tl]{Slice thickness \\ (mm)} & 1 & 1 & 4 & \makecell[tl]{2, reconstructed \\ to 1 with zero \\ filling (ZIP2)} \\
+        In-plane resolution (mm) & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 & 0.977 x 0.977 \\
+        Slice thickness (mm) & 1 & 1 & 4 & 2, reconstructed to 1 with zero filling (ZIP2) \\
         Number of slices & 126 & 106 & 50 & 92 \\
-        Additional parameters & n/a & n/a & \makecell[tl]{1,450 ms label period; \\ 2,025 ms pulse label; \\ 24 control-label pairs} & n/a \\
+        Additional parameters & n/a & n/a & 1,450 ms label period; 2,025 ms pulse label; 24 control-label pairs & n/a \\
         Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s \\ \botrule
@@
 ```{=latex}
-    \begin{tabular}{@{}llllll@{}}
+    \begin{tabular}{@{}lllp{2cm}ll@{}}
     \toprule
-        Region & Measure & QSM & \makecell[tl]{Paramagnetic \\ map} & p-value & 95\% CI \\ \midrule
+        Region & Measure & QSM & Paramagnetic map & p-value & 95\% CI \\ \midrule
         \multirow{ 2}{*}{SSS} & Chi (ppm)     & `r round(mean(newdata$Gavin_SSS_Chi),2)` \pm `r round(sd(newdata$Gavin_SSS_Chi),2)`                     & `r round(mean(newdata$Chisep_SSS_Chi),2)` \pm `r round(sd(newdata$Chisep_SSS_Chi),2)`                     & `r signif(SSSChi.t$p.value,3)`                        & `r round(SSSChi.t$conf.low,2)`, `r round(SSSChi.t$conf.high,2)`           \\

### 2024-06-11 index.qmd
changed all tables and rendered

@@
 library(Rmpfr)
+library(gt)
 # library(reticulate) # incorporate Python
@@
 ```{r}
+# function to make rounding means and sd easier
 rndmean <- function(clm) {
@@
 
-::: {#tbl-mri}
-```{=latex}
-    \begin{tabular*}{\textwidth}{@{}p{2.5cm}llp{2.5cm}p{2.5cm}@{}}
-    \toprule
-        ~ & T1w & T2w & pcASL & SWI \\ \midrule
-        Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D fast spin-echo & 3D spoiled GRE flow-compensated \\
-        Phase-encoding direction & Coronal & Sagittal & Axial & Axial \\
-        TR (ms) & 7.74 & 2,300 & 4,680 & 30.9 \\
-        TE (ms) & 2.97 & 66.29 & 10.55 & 5.24 \\
-        Flip angle & 12$^{\circ}$ & 90$^{\circ}$ & 111$^{\circ}$ & 20$^{\circ}$ \\
-        FOV (cm) & 20 & 20 & 24 & 25 \\
-        Acquisition matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256 \\
-        In-plane resolution (mm) & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 & 0.977 x 0.977 \\
-        Slice thickness (mm) & 1 & 1 & 4 & 2, reconstructed to 1 with zero filling (ZIP2) \\
-        Number of slices & 126 & 106 & 50 & 92 \\
-        Additional parameters & n/a & n/a & 1,450 ms label period; 2,025 ms pulse label; 24 control-label pairs & n/a \\
-        Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s \\ \botrule
-    \end{tabular*}
-    \footnotetext{T1w = T1-weighted; T2w = T2-weighted; pcASL = pseudo-continuous arterial spin labelling; SWI = susceptibility weighted imaging; FSPG = fast spoiled gradient echo; CUBE = General Electric name of sequence, not an acronym; GRE = gradient echo; ZIP2 = through-plane zero filling interpolation}
+```{r}
+#| label: tbl-mri
+#| tbl-cap: "Technical parameters for MR imaging pulse sequences"
+df <- data.frame(Scan = character(), T1w = character(), T2w = character(), pcASL = character(), SWI = character(), stringsAsFactors = FALSE)
+df[1,] <- c("Sequence", "3D FSPGR", "3D CUBE", "Multi-shot 3D fast spin-echo", "3D spoiled GRE flow-compensated")
+df[2,] <- c("Phase-encoding direction", "Coronal", "Sagittal", "Axial", "Axial")
+df[3,] <- c("TR (ms)", "7.74", "2,300", "4,680", "30.9")
+df[4,] <- c("TE (ms)", "2.97", "66.29", "10.55", "5.24")
+df[5,] <- c("Flip angle", "12\U00B0", "90\U00B0", "111\U00B0", "20\U00B0")
+df[6,] <- c("FOV (cm)", "20", "20", "24", "25")
+df[7,] <- c("Acquisition matrix", "512 x 512", "256 x 256", "128 x 128", "256 x 256")
+df[8,] <- c("In-plane resolution (mm)", "0.39 x 0.39", "0.78 x 0.78", "1.875 x 1.875", "0.977 x 0.977")
+df[9,] <- c("Slice thickness (mm)", "1", "1", "4", "2, reconstructed to 1 with zero filling (ZIP2)")
+df[10,] <- c("Number of slices", "126", "106", "50", "92")
+df[11,] <- c("Additional parameters", "n/a", "n/a", "1,450 ms label period;\n 2,025 ms pulse label;\n 24 control-label pairs", "n/a")
+df[12,] <- c("Scan duration", "4 min 39 s", "5 min 1 s", "5 min 26 s", "5 min 29 s")
+
+footnotetext="T1w = T1-weighted; T2w = T2-weighted; pcASL = pseudo-continuous arterial spin labelling; SWI = susceptibility weighted imaging; FSPG = fast spoiled gradient echo; CUBE = General Electric name of sequence, not an acronym; GRE = gradient echo; ZIP2 = through-plane zero filling interpolation"
+
+if (knitr::is_latex_output()) {
+colnames(df)[1] <- ""
+df[11,] <- linebreak(df[11,])
+df |>
+  kbl(format = "latex",
+    booktabs = TRUE,
+    longtable = TRUE,
+    linesep = "",
+    align = "l",
+    escape = FALSE) |>
+    kable_styling(font_size = 8, position = "center", latex_options = c("hold_position", "scale_down"), full_width = FALSE) |>
+    footnote(general_title = "", 
+           footnote_as_chunk = TRUE,
+           threeparttable = TRUE,
+           general = footnotetext) |>
+    column_spec(1, width="8em") |>
+    column_spec(4, width="9em") |>
+    column_spec(5, width="9em")
+} else {
+df |> 
+    mutate(across(everything(), ~ str_replace_all(., "\n", "<br>"))) |> 
+    gt(rowname_col = "Scan") |> tab_footnote(footnotetext) |>
+    fmt_markdown(columns = TRUE) |>
+    tab_options(quarto.disable_processing = TRUE)
+}
 ```
-Technical parameters for MR imaging pulse sequences
-:::
 
@@
 
-::: {#tbl-dem}
-```{=latex}
-\begin{tabular}{@{}ll@{}}
-    \toprule
-Variable & Subject Data (n=19) \\
-  \midrule
-Gestational age, weeks (mean $\pm$ SD)                       & `r round(mean(newdata$GA),2)` $\pm$ `r round(sd(newdata$GA),2)`                                                     \\
-Corrected gestational age on scan day, weeks (mean $\pm$ SD) & `r round(mean(newdata$CGAonScanDay),2)` $\pm$ `r round(sd(newdata$CGAonScanDay),2)`                                 \\
-Number of male neonates (%)                                  & `r sum(newdata$Sex == "M")` (`r round(sum(newdata$Sex == "M")/length(newdata$Sex)*100,2)`)                          \\
-Birth weight, g (mean $\pm$ SD)                              & `r round(mean(newdata$BW),2)` $\pm$ `r round(sd(newdata$BW),2)`                                                     \\
-Weight on scan day, g (mean $\pm$ SD)                        & `r round(mean(newdata$Weight.on.Scan.Day..kg.*1000),2)` $\pm$ `r round(sd(newdata$Weight.on.Scan.Day..kg.*1000),2)` \\
-Days spent in NICU (median, IQR)                             & `r median(newdata$Total_Days_NICU)`, `r IQR(newdata$Total_Days_NICU)`                                               \\
-Days on ventilation (median, IQR)                            & `r median(newdata$Total_Days_Ventilation)`, `r IQR(newdata$Total_Days_Ventilation)`                                 \\
-\botrule
-\end{tabular}
-\footnotetext{SD = standard deviation; IQR = inter quartile range}
+```{r}
+#| label: tbl-dem
+#| tbl-cap: Demographic and clinical characteristic of the study sample.
+
+df <- data.frame(Variable = character(), "Subject" = character(), stringsAsFactors = FALSE)
+df[1,] <- c("Gestational age, weeks (mean \U00B1 SD)", paste0(rndmean(newdata$GA), " \U00B1 ", rndsd(newdata$GA)))
+df[2,] <- c("Corrected gestational age on scan day, weeks (mean \U00B1 SD) ", paste0(rndmean(newdata$CGAonScanDay), " \U00B1 ", rndsd(newdata$CGAonScanDay)))
+df[3,] <- c("Number of male neonates (\\%)", paste0(sum(newdata$Sex == "M"), " (", (round(sum(newdata$Sex == "M")/length(newdata$Sex)*100,2)), ")"))
+df[4,] <- c("Birth weight, g (mean \U00B1 SD)", paste0(rndmean(newdata$BW), " \U00B1 ", rndsd(newdata$BW)))
+df[5,] <- c("Weight on scan day, g (mean \U00B1 SD)", paste0(rndmean(newdata$Weight.on.Scan.Day..kg.*1000), " \U00B1 ", rndsd(newdata$Weight.on.Scan.Day..kg.*1000)))
+df[6,] <- c("Days spent in NICU (median, IQR)", paste0(median(newdata$Total_Days_NICU), ", ", IQR(newdata$Total_Days_NICU)))
+df[7,] <- c("Days on ventilation (median, IQR)", paste0(median(newdata$Total_Days_Ventilation), ", ", IQR(newdata$Total_Days_Ventilation)))
+df <- df |> rename("Subject data (n = 19)" = Subject)
+footnotetext="SD = standard deviation; IQR = inter quartile range"
+
+if (knitr::is_latex_output()) {
+df |>
+  kbl(format = "latex",
+    booktabs = TRUE,
+    longtable = TRUE,
+    linesep = "",
+    align = "lc",
+    escape = FALSE) |>
+    kable_styling(font_size = 9, position = "center", latex_options = c("hold_position", "scale_down"), full_width = FALSE) |>
+    footnote(general_title = "", 
+           footnote_as_chunk = TRUE,
+           threeparttable = TRUE,
+           general = footnotetext)
+} else {
+df |> 
+    mutate(across(everything(), ~ str_replace_all(., "\n", "<br>"))) |> 
+    gt() |> 
+    cols_align(align = "center", columns = c("Subject data (n = 19)")) |>
+    tab_footnote(footnotetext) |>
+    fmt_markdown(columns = TRUE) |>
+    tab_options(quarto.disable_processing = TRUE)
+}
 ```
-Demographic and clinical characteristic of the study sample.
-:::
 
@@
 
-::: {#tbl-chistats}
-```{=latex}
-    \begin{tabular}{@{}lllp{2cm}ll@{}}
-    \toprule
-        Region & Measure & QSM & Paramagnetic map & p-value & 95\% CI \\ \midrule
-        \multirow{ 2}{*}{SSS} & Chi (ppm)     & `r round(mean(newdata$Gavin_SSS_Chi),2)` \pm `r round(sd(newdata$Gavin_SSS_Chi),2)`                     & `r round(mean(newdata$Chisep_SSS_Chi),2)` \pm `r round(sd(newdata$Chisep_SSS_Chi),2)`                     & `r signif(SSSChi.t$p.value,3)`                        & `r round(SSSChi.t$conf.low,2)`, `r round(SSSChi.t$conf.high,2)`           \\
-                              & SvO$_{2}$(\%) & `r round(mean(newdata$Gavin_SSSVein_CSvO2*100),2)` \pm `r round(sd(newdata$Gavin_SSSVein_CSvO2*100),2)` & `r round(mean(newdata$Chisep_SSSVein_CSvO2*100),2)` \pm `r round(sd(newdata$Chisep_SSSVein_CSvO2*100),2)` & `r signif(SSSSvO2.t$p.value,3)` & `r round(SSSSvO2.t$conf.low*100,2)`, `r round(SSSSvO2.t$conf.high*100,2)` \\
-        \multirow{ 2}{*}{CCV} & Chi (ppm)     & `r round(mean(newdata$Gavin_IntVein_Chi),2)` \pm `r round(sd(newdata$Gavin_IntVein_Chi),2)`             & `r round(mean(newdata$Chisep_IntVein_Chi),2)` \pm `r round(sd(newdata$Chisep_IntVein_Chi),2)`             & `r signif(CCVChi.t$p.value,3)`                        & `r round(CCVChi.t$conf.low,2)`, `r round(CCVChi.t$conf.high,2)`           \\
-                              & SvO$_{2}$(\%) & `r round(mean(newdata$Gavin_IntVein_CSvO2*100),2)` \pm `r round(sd(newdata$Gavin_IntVein_CSvO2*100),2)` & `r round(mean(newdata$Chisep_IntVein_CSvO2*100),2)` \pm `r round(sd(newdata$Chisep_IntVein_CSvO2*100),2)` & `r signif(CCVSvO2.t$p.value,3)` & `r round(CCVSvO2.t$conf.low*100,2)`, `r round(CCVSvO2.t$conf.high*100,2)` \\
-        \botrule
-    \end{tabular}
-    \footnotetext{QSM = quantitative susceptibility mapping; CI = confidence interval; SSS = superior sagitall sinus; CCV = central cerebral vein}
+```{r}
+#| label: tbl-chistats
+#| tbl-cap: Summary of acquired physiological parameters.
+Mean $\pm$ SD is shown for chi and SvO~2~ values.
+The P-value and 95% confidence interval (CI) were obtained through the comparison of values between QSM and paramagnetic maps; (n=19).
+
+df <- data.frame(Region = character(), Measure = character(), QSM = character(), pmap = character(), pvalue = character(), CI = character(), stringsAsFactors = FALSE)
+
+df[1,] <- c("SSS", "Chi (ppm)", paste0(rndmean(newdata$Gavin_SSS_Chi), " \U00B1 ", rndsd(newdata$Gavin_SSS_Chi)), paste0(rndmean(newdata$Chisep_SSS_Chi), " \U00B1 ", rndsd(newdata$Chisep_SSS_Chi)), signif(SSSChi.t$p.value,3), paste0(round(SSSChi.t$conf.low,2), ", ", round(SSSChi.t$conf.high,2)))
+df[2,] <- c("SSS", "SvO\U2082 (\\%)", paste0(rndmean(newdata$Gavin_SSSVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Gavin_SSSVein_CSvO2*100)), paste0(rndmean(newdata$Chisep_SSSVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Chisep_SSSVein_CSvO2*100)), signif(SSSSvO2.t$p.value,3), paste0(round(SSSSvO2.t$conf.low,2), ", ", round(SSSSvO2.t$conf.high,2)))
+df[3,] <- c("CCV", "Chi (ppm)", paste0(rndmean(newdata$Gavin_IntVein_Chi), " \U00B1 ", rndsd(newdata$Gavin_IntVein_Chi)), paste0(rndmean(newdata$Chisep_IntVein_Chi), " \U00B1 ", rndsd(newdata$Chisep_IntVein_Chi)), signif(CCVChi.t$p.value,3), paste0(round(CCVChi.t$conf.low,2), ", ", round(CCVChi.t$conf.high,2)))
+df[4,] <- c("CCV", "SvO\U2082 (\\%)", paste0(rndmean(newdata$Gavin_IntVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Gavin_IntVein_CSvO2*100)), paste0(rndmean(newdata$Chisep_IntVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Chisep_IntVein_CSvO2*100)), signif(CCVSvO2.t$p.value,3), paste0(round(CCVSvO2.t$conf.low,2), ", ", round(CCVSvO2.t$conf.high,2)))
+
+df <- df |> rename("Paramagnetic map" = pmap, "p-value" = pvalue, "95\U0025 CI" = CI)
+
+footnotetext="QSM = quantitative susceptibility mapping; CI = confidence interval; SSS = superior sagitall sinus; CCV = central cerebral vein"
+
+if (knitr::is_latex_output()) {
+df |>
+  kbl(format = "latex",
+    booktabs = TRUE,
+    longtable = TRUE,
+    linesep = "",
+    align = "llcccc",
+    escape = FALSE) |>
+    kable_styling(font_size = 9, position = "center", latex_options = c("hold_position", "scale_down"), full_width = FALSE) |>
+    footnote(general_title = "", 
+           footnote_as_chunk = TRUE,
+           threeparttable = TRUE,
+           general = footnotetext)
+} else {
+df |> 
+    mutate(across(everything(), ~ str_replace_all(., "\n", "<br>"))) |> 
+    gt() |> 
+    cols_align(align = "center", columns = c("QSM", "Paramagnetic map", "p-value", "95\U0025 CI")) |>
+    tab_footnote(footnotetext) |>
+    fmt_markdown(columns = TRUE) |>
+    tab_options(quarto.disable_processing = TRUE)
+}
 ```
-Summary of acquired physiological parameters.
-Mean ± SD is shown for chi and SvO~2~ values.
-The P-value and 95% confidence interval (CI) were obtained through the comparison of values between QSM and paramagnetic maps; (n=19).
-:::
 

### 2024-06-11 _tex/index.tex
Built site for gh-pages

@@
 
-    \begin{tabular*}{\textwidth}{@{}lllll@{}}
+    \begin{tabular*}{\textwidth}{@{}p{2.5cm}llp{2.5cm}p{2.5cm}@{}}
     \toprule
         ~ & T1w & T2w & pcASL & SWI \\ \midrule
-        Sequence & 3D FSPGR & 3D CUBE & \makecell[tl]{Multi-shot 3D fast \\ spin-echo} & \makecell[tl]{3D spoiled GRE \\ flow-compensated} \\
-        \makecell[tl]{Phase-encoding \\ direction} & Coronal & Sagittal & Axial & Axial \\
+        Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D fast spin-echo & 3D spoiled GRE flow-compensated \\
+        Phase-encoding direction & Coronal & Sagittal & Axial & Axial \\
         TR (ms) & 7.74 & 2,300 & 4,680 & 30.9 \\
@@
 
-    \begin{tabular}{@{}llllll@{}}
+    \begin{tabular}{@{}lllp{2cm}ll@{}}
     \toprule
-        Region & Measure & QSM & \makecell[tl]{Paramagnetic \\ map} & p-value & 95\% CI \\ \midrule
+        Region & Measure & QSM & Paramagnetic map & p-value & 95\% CI \\ \midrule
         \multirow{ 2}{*}{SSS} & Chi (ppm)     & 0.1 \pm 0.02                     & 0.21 \pm 0.05                     & \ensuremath{2.84\times 10^{-11}}                        & -0.13, -0.09           \\

### 2024-06-25 index.qmd
working version

@@
 df <- data.frame(Region = character(), Measure = character(), QSM = character(), pmap = character(), pvalue = character(), CI = character(), stringsAsFactors = FALSE)
-
 df[1,] <- c("SSS", "Chi (ppm)", paste0(rndmean(newdata$Gavin_SSS_Chi), " \U00B1 ", rndsd(newdata$Gavin_SSS_Chi)), paste0(rndmean(newdata$Chisep_SSS_Chi), " \U00B1 ", rndsd(newdata$Chisep_SSS_Chi)), signif(SSSChi.t$p.value,3), paste0(round(SSSChi.t$conf.low,2), ", ", round(SSSChi.t$conf.high,2)))
@@
 df[4,] <- c("CCV", "SvO\U2082 (\\%)", paste0(rndmean(newdata$Gavin_IntVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Gavin_IntVein_CSvO2*100)), paste0(rndmean(newdata$Chisep_IntVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Chisep_IntVein_CSvO2*100)), signif(CCVSvO2.t$p.value,3), paste0(round(CCVSvO2.t$conf.low,2), ", ", round(CCVSvO2.t$conf.high,2)))
-
-df <- df |> rename("Paramagnetic map" = pmap, "p-value" = pvalue, "95\U0025 CI" = CI)
-
 footnotetext="QSM = quantitative susceptibility mapping; CI = confidence interval; SSS = superior sagitall sinus; CCV = central cerebral vein"
@@
 if (knitr::is_latex_output()) {
+df <- df |> rename("Paramagnetic map" = pmap, "p-value" = pvalue, "95\\\U0025 CI" = CI)
 df |>
@@
 } else {
+df <- df |> rename("Paramagnetic map" = pmap, "p-value" = pvalue, "95\U0025 CI" = CI)
 df |> 

### 2024-06-25 _tex/index.tex
Built site for gh-pages

@@
 \makeatother
+
 \ifLuaTeX
@@
 
-\begin{table}
+\begingroup\fontsize{8}{10}\selectfont
+
+\begin{ThreePartTable}
+\begin{TableNotes}[para]
+\item T1w = T1-weighted; T2w = T2-weighted; pcASL = pseudo-continuous arterial spin labelling; SWI = susceptibility weighted imaging; FSPG = fast spoiled gradient echo; CUBE = General Electric name of sequence, not an acronym; GRE = gradient echo; ZIP2 = through-plane zero filling interpolation
+\end{TableNotes}
+
+\begin{longtable}[t]{>{\raggedright\arraybackslash}p{8em}ll>{\raggedright\arraybackslash}p{9em}>{\raggedright\arraybackslash}p{9em}}
 
@@
 
-\centering{
+\tabularnewline
 
-    \begin{tabular*}{\textwidth}{@{}p{2.5cm}llp{2.5cm}p{2.5cm}@{}}
 \toprule
-        ~ & T1w & T2w & pcASL & SWI \\ \midrule
-        Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D fast spin-echo & 3D spoiled GRE flow-compensated \\
-        Phase-encoding direction & Coronal & Sagittal & Axial & Axial \\
-        TR (ms) & 7.74 & 2,300 & 4,680 & 30.9 \\
-        TE (ms) & 2.97 & 66.29 & 10.55 & 5.24 \\
-        Flip angle & 12$^{\circ}$ & 90$^{\circ}$ & 111$^{\circ}$ & 20$^{\circ}$ \\
-        FOV (cm) & 20 & 20 & 24 & 25 \\
-        Acquisition matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256 \\
-        In-plane resolution (mm) & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 & 0.977 x 0.977 \\
-        Slice thickness (mm) & 1 & 1 & 4 & 2, reconstructed to 1 with zero filling (ZIP2) \\
-        Number of slices & 126 & 106 & 50 & 92 \\
-        Additional parameters & n/a & n/a & 1,450 ms label period; 2,025 ms pulse label; 24 control-label pairs & n/a \\
-        Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s \\ \botrule
-    \end{tabular*}
-    \footnotetext{T1w = T1-weighted; T2w = T2-weighted; pcASL = pseudo-continuous arterial spin labelling; SWI = susceptibility weighted imaging; FSPG = fast spoiled gradient echo; CUBE = General Electric name of sequence, not an acronym; GRE = gradient echo; ZIP2 = through-plane zero filling interpolation}
+ & T1w & T2w & pcASL & SWI\\
+\midrule
+Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D fast spin-echo & 3D spoiled GRE flow-compensated\\
+Phase-encoding direction & Coronal & Sagittal & Axial & Axial\\
+TR (ms) & 7.74 & 2,300 & 4,680 & 30.9\\
+TE (ms) & 2.97 & 66.29 & 10.55 & 5.24\\
+Flip angle & 12° & 90° & 111° & 20°\\
+FOV (cm) & 20 & 20 & 24 & 25\\
+Acquisition matrix & 512 x 512 & 256 x 256 & 128 x 128 & 256 x 256\\
+In-plane resolution (mm) & 0.39 x 0.39 & 0.78 x 0.78 & 1.875 x 1.875 & 0.977 x 0.977\\
+Slice thickness (mm) & 1 & 1 & 4 & 2, reconstructed to 1 with zero filling (ZIP2)\\
+Number of slices & 126 & 106 & 50 & 92\\
+Additional parameters & n/a & n/a & 1,450 ms label period;\newline  2,025 ms pulse label;\newline  24 control-label pairs & n/a\\
+Scan duration & 4 min 39 s & 5 min 1 s & 5 min 26 s & 5 min 29 s\\
+\bottomrule
+\insertTableNotes
 
-}
+\end{longtable}
 
-\end{table}%
+\end{ThreePartTable}
+\endgroup{}
 
@@
 
-\begin{table}
+\begingroup\fontsize{9}{11}\selectfont
+
+\begin{ThreePartTable}
+\begin{TableNotes}[para]
+\item SD = standard deviation; IQR = inter quartile range
+\end{TableNotes}
+
+\begin{longtable}[t]{lc}
 
@@
 
-\centering{
+\tabularnewline
 
-\begin{tabular}{@{}ll@{}}
 \toprule
-Variable & Subject Data (n=19) \\
+Variable & Subject data (n = 19)\\
 \midrule
-Gestational age, weeks (mean $\pm$ SD)                       & 28.8 $\pm$ 1.68                                                     \\
-Corrected gestational age on scan day, weeks (mean $\pm$ SD) & 40.36 $\pm$ 1.4                                 \\
-Number of male neonates (%)                                  & 10 (52.63)                          \\
-Birth weight, g (mean $\pm$ SD)                              & 1276.05 $\pm$ 294.87                                                     \\
-Weight on scan day, g (mean $\pm$ SD)                        & 3396.58 $\pm$ 597.72 \\
-Days spent in NICU (median, IQR)                             & 53, 23                                               \\
-Days on ventilation (median, IQR)                            & 31, 28.5                                 \\
-\botrule
-\end{tabular}
-\footnotetext{SD = standard deviation; IQR = inter quartile range}
+Gestational age, weeks (mean ± SD) & 28.8 ± 1.68\\
+Corrected gestational age on scan day, weeks (mean ± SD) & 40.36 ± 1.4\\
+Number of male neonates (\%) & 10 (52.63)\\
+Birth weight, g (mean ± SD) & 1276.05 ± 294.87\\
+Weight on scan day, g (mean ± SD) & 3396.58 ± 597.72\\
+Days spent in NICU (median, IQR) & 53, 23\\
+Days on ventilation (median, IQR) & 31, 28.5\\
+\bottomrule
+\insertTableNotes
 
-}
+\end{longtable}
 
-\end{table}%
+\end{ThreePartTable}
+\endgroup{}
 
@@
 
-\begin{table}
+\begingroup\fontsize{9}{11}\selectfont
+
+\begin{ThreePartTable}
+\begin{TableNotes}[para]
+\item QSM = quantitative susceptibility mapping; CI = confidence interval; SSS = superior sagitall sinus; CCV = central cerebral vein
+\end{TableNotes}
+
+\begin{longtable}[t]{llcccc}
 
 \caption{\label{tbl-chistats}Summary of acquired physiological
-parameters. Mean ± SD is shown for chi and SvO\textsubscript{2} values.
-The P-value and 95\% confidence interval (CI) were obtained through the
-comparison of values between QSM and paramagnetic maps; (n=19).}
+parameters. Mean \(\pm\) SD is shown for chi and SvO\textsubscript{2}
+values. The P-value and 95\% confidence interval (CI) were obtained
+through the comparison of values between QSM and paramagnetic maps;
+(n=19).}
 
-\centering{
+\tabularnewline
 
-    \begin{tabular}{@{}lllp{2cm}ll@{}}
 \toprule
-        Region & Measure & QSM & Paramagnetic map & p-value & 95\% CI \\ \midrule
-        \multirow{ 2}{*}{SSS} & Chi (ppm)     & 0.1 \pm 0.02                     & 0.21 \pm 0.05                     & \ensuremath{2.84\times 10^{-11}}                        & -0.13, -0.09           \\
-                              & SvO$_{2}$(\%) & 72.46 \pm 3.43 & 58.14 \pm 7.3 & \ensuremath{6.12\times 10^{-10}} & 11.78, 16.85 \\
-        \multirow{ 2}{*}{CCV} & Chi (ppm)     & 0.13 \pm 0.02             & 0.22 \pm 0.05             & \ensuremath{6.25\times 10^{-9}}                        & -0.1, -0.07           \\
-                              & SvO$_{2}$(\%) & 68.71 \pm 3.46 & 57.69 \pm 6.97 & \ensuremath{2.16\times 10^{-9}} & 8.9, 13.13 \\
-        \botrule
-    \end{tabular}
-    \footnotetext{QSM = quantitative susceptibility mapping; CI = confidence interval; SSS = superior sagitall sinus; CCV = central cerebral vein}
+Region & Measure & QSM & Paramagnetic map & p-value & 95\% CI\\
+\midrule
+SSS & Chi (ppm) & 0.1 ± 0.02 & 0.21 ± 0.05 & 2.84e-11 & -0.13, -0.09\\
+SSS & SvO₂ (\%) & 72.46 ± 3.43 & 58.14 ± 7.3 & 6.12e-10 & 0.12, 0.17\\
+CCV & Chi (ppm) & 0.13 ± 0.02 & 0.22 ± 0.05 & 6.25e-09 & -0.1, -0.07\\
+CCV & SvO₂ (\%) & 68.71 ± 3.46 & 57.69 ± 6.97 & 2.16e-09 & 0.09, 0.13\\
+\bottomrule
+\insertTableNotes
 
-}
+\end{longtable}
 
-\end{table}%
+\end{ThreePartTable}
+\endgroup{}
 

### 2024-06-25 index.qmd
added notebook links

@@
 number-sections: false
-notebook-links: false
+notebook-links: true
 ---
@@
 # library(reticulate) # incorporate Python
-# use_virtualenv('./pythonvenv', required = TRUE) # load pythong venv from path
+# use_virtualenv('./pyvenv_csvo', required = TRUE) # load pythong venv from path
+
 # library(neurobase)
@@
 SWI magnitude data files were then used to create subject-specific brain masks that would not erode the SSS during QSM processing, an issue faced by our group in the past [@liFirstStepNeuroimaging2016].
-A step-by-step summary of the pipeline used is shown in @fig-sample.
+A step-by-step summary of the pipeline used is shown in @fig-graph.
 

### 2024-06-25 _tex/index.tex
Built site for gh-pages

@@
 
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
+Notebook}}
+
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
+Notebook}}
+
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
+Notebook}}
+
 \section{Introduction}\label{sec-intro}
@@
 
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
+Notebook}}
+
 The MRI scan protocol comprised of the following sequences: a
@@
 al. 2016). A step-by-step summary of the pipeline used is shown in
-Figure~\ref{fig-sample}.
+Figure~\ref{fig-graph}.
 
@@
 \end{figure}%
+
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures-preview.html\#cell-fig-graph}{Python
+Figures}}
 
@@
 
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures-preview.html\#cell-fig-sample}{Python
+Figures}}
+
 Once the mean susceptibility values of the SSS and CCV were obtained
@@
 
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
+Notebook}}
+
 The mean SvO\textsubscript{2} values for the SSS and the CCV were found
@@
 
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
+Notebook}}
+
 Region-specific \(\chi\) and SvO\textsubscript{2} values acquired from
@@
 \end{figure}%
+
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/gavin_thesis_markdown-preview.html\#cell-fig-methodplot}{gavin\_thesis\_R}}
 
@@
 \end{figure}%
+
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/gavin_thesis_markdown-preview.html\#cell-fig-regionplot}{gavin\_thesis\_R}}
 

### 2024-06-25 notebooks/gavin_thesis_markdown.qmd
Built site for gh-pages

@@
 newdata <- read.csv("newdata.csv")
+
+signif="***"
+notsig="n.s."
+
+sigfunc <- function(pval) {
+if (pval < 0.05){
+  label = signif
+} else {
+    label = notsig
+}
+  return (label)
+}
 ```
@@
 
-Wilcox test for p-values and 95% CI - table 3.2
-```{r}
-#SSS Chi
-wilcox.test(newdata$Gavin_SSS_Chi, newdata$Chisep_SSS_Chi,
-            paired = TRUE, conf.int = TRUE)
-
-#SSS SvO2
-wilcox.test(newdata$Gavin_SSSVein_CSvO2, newdata$Chisep_SSSVein_CSvO2,
-            paired = TRUE, conf.int = TRUE)
-
-#CCV Chi
-wilcox.test(newdata$Gavin_IntVein_Chi, newdata$Chisep_IntVein_Chi, paired = TRUE,
-            conf.int = TRUE)
-
-#CCV SvO2
-wilcox.test(newdata$Gavin_IntVein_CSvO2, newdata$Chisep_IntVein_CSvO2,
-            paired = TRUE, conf.int = TRUE)
-```
-
 ```{r}
@@
 Figure 3.1
-*note: "****" in panels were added outside of R
 ```{r}
 #| echo = FALSE
+label = sigfunc(SSSChi.t$p.value)
 m1 <- ggwithinstats(
@@
   y = Chi,
+  ylab = "Chi (ppm)",
   results.subtitle = FALSE,
-  title = "(A)",
-  subtitle = "Central Cerebral Veins",
-  type = "nonparametric",
+  type = "p",
+  pairwise.display = "all",
+  bf.message = FALSE,
   centrality.plotting = FALSE,
-  point.path.args = list(linewidth = 0.25, linetype = "solid")
-)
+  point.path.args = list(linewidth = 0.25, linetype = "solid")) +
+  ggplot2::annotate("text", x=1.5, y=(max(comp_CCV_chi$Chi)*1.05), size=4, label = label)
 
+label = sigfunc(SSSSvO2.t$p.value)
 m2 <- ggwithinstats( 
@@
   y = Measured.Susceptibility, 
-  ylab = "Chi",
+  ylab = "Chi (ppm)",
   results.subtitle = FALSE,
-  title = "",
-  subtitle = "Superior Sagittal Sinus",
-  type = "nonparametric", 
+  type = "p",
+  pairwise.display = "all",
+  bf.message = FALSE,
   centrality.plotting = FALSE,
-  point.path.args = list(linewidth = 0.25, linetype = "solid")
-)
+  point.path.args = list(linewidth = 0.25, linetype = "solid")) +
+  ggplot2::annotate("text", x=1.5, y=(max(comp_SSS_chi$Measured.Susceptibility)*1.05), size=4, label = label )
 
+label = sigfunc(CCVChi.t$p.value)
 m3 <- ggwithinstats(
@@
   y = Oxygen.Saturation,
-  ylab = "SvO2",
+  ylab = "SvO2 (%)",
   results.subtitle = FALSE,
-  title = "(B)",
-  subtitle = "Central Cerebral Veins",
-  type = "nonparametric",
+  type = "p",
+  pairwise.display = "all",
+  bf.message = FALSE,
   centrality.plotting = FALSE,
-  point.path.args = list(linewidth = 0.25, linetype = "solid")
-)
+  point.path.args = list(linewidth = 0.25, linetype = "solid") ) +
+  ggplot2::annotate("text", x=1.5, y=(max(comp_CCV_SvO2$Oxygen.Saturation)*1.05), size=4, label = label )
 
+label = sigfunc(CCVSvO2.t$p.value)
 m4 <- ggwithinstats(
@@
   y = Oxygen.Saturation,
+  ylab = "SvO2 (%)",
+  results.subtitle = FALSE,
+  type = "p",
   pairwise.display = "all",
-  ylab = "SvO2",
-  results.subtitle = FALSE,
-  title = "",
-  subtitle = "Superior Sagittal Sinus",
-  type = "nonparametric",
+  bf.message = FALSE,
   centrality.plotting = FALSE,
-  point.path.args = list(linewidth = 0.25, linetype = "solid")
-)
-
-combine_plots(
-  plotlist = list(m1, m2, m3, m4),
-  plotgrid.args = list(nrow = 2)
-)
-```
-
-Figure 3.2
-*note: "****" in panels (a) and (b) were added outside of R
-```{r}
-#| echo = FALSE
-p1 <- ggwithinstats(
-  data = comp_legit_starfsl,
-  x = Vein,
-  y = Chi,
-  pairwise.display = "all",
-  results.subtitle = FALSE,
-  xlab = "Region",
-  ylab = "Chi",
-  title = "(A)",
-  type = "nonparametric",
-  centrality.plotting = FALSE,
-  point.path.args = list(linewidth = 0.25, linetype = "solid"),
-  point.args = list(color = "#D95F02") 
-)
-
-p2 <- ggwithinstats(
-  data = comp_SSS_IntVein_CSvO2,
-  x = Vein,
-  y = CSvO2,
-  pairwise.display = "all",
-  results.subtitle = FALSE,
-  xlab = "Region",
-  ylab = "SvO2",
-  title = "(B)",
-  type = "nonparametric",
-  centrality.plotting = FALSE,
-  point.path.args = list(linewidth = 0.25, linetype = "solid"),
-  point.args = list(color = "#D95F02")
-)
-
-p3 <- ggwithinstats(                     #boxplot of chi comparison between regions
-  data = comp_chisep_chi,
-  x = Vein,
-  y = Chi, 
-  pairwise.display = "all",
-  results.subtitle = FALSE,
-  xlab = "Region",
-  ylab = "Chi",
-  title = "(C)",
-  subtitle = "                                                n.s.",
-  type = "nonparametric", 
-  centrality.plotting = FALSE,
-  point.path.args = list(linewidth = 0.25, linetype = "solid"),
-  point.args = list(color = "#1B9E77")
-)
-
-p4 <- ggwithinstats(                    #boxplot of svo2 comparison between regions
-  data = comp_chisep_CSvO2,
-  x = Vein,
-  y = CSvO2, 
-  pairwise.display = "all",
-  results.subtitle = FALSE,
-  xlab = "Region",
-  ylab = "SvO2",
-  title = "(D)",
-  subtitle = "                                                n.s.",
-  type = "nonparametric",
-  centrality.plotting = FALSE,
-  point.path.args = list(linewidth = 0.25, linetype = "solid"),
-  point.args = list(color = "#1B9E77")
-)
-
-combine_plots(
-  plotlist = list(p1, p2, p3, p4),
-  plotgrid.args = list(nrow = 2)
-)
+  point.path.args = list(linewidth = 0.25, linetype = "solid") ) +
+  ggplot2::annotate("text", x=1.5, y=(max(comp_SSS_SvO2$Oxygen.Saturation)*1.05), size=4, label = label )
+(patch1 <- (m1 + m2)/(m3 + m4) + 
+  plot_annotation(tag_levels = 'A') & 
+  theme(plot.tag = element_text(size = 10)))
 ```
@@
 ```{r}
-save.image('results.RData')
+#| label: fig-methodplot
+#| fig-width: 8
+#| fig-height: 5
+#| fig-cap: "Vein-specific susceptibility and oxygen saturation values by method.
+(A, B) contains violin plots comparing subject χ (ppm) acquired from QSM (A) and its paramagnetic component (B); (C, D) contains violin plots comparing subject SvO2 (%) acquired from QSM (C) and its paramagnetic component (D).
+Raw data points from paramagnetic maps are shown as filled green circles and raw data points from QSM are shown as filled orange circles.
+Each line connects the raw data points of a single subject.
+(***) indicates P<0.05."
+patch1
+```
+
+# Region analysis: Chi and CSvO2
+
+```{r}
+#SSS Chi
+RegionQSMChi.t <- tidy(t.test(newdata$Gavin_IntVein_Chi, newdata$Gavin_SSS_Chi, 
+            paired = TRUE, conf.int = TRUE))
+
+#SSS SvO2
+RegionQSMSvO2.t <- tidy(t.test(newdata$Gavin_IntVein_CSvO2, newdata$Gavin_SSSVein_CSvO2,
+            paired = TRUE, conf.int = TRUE))
+
+#CCV Chi
+RegionChisepChi.t <- tidy(t.test(newdata$Chisep_IntVein_Chi, newdata$Chisep_SSS_Chi, paired = TRUE,
+            conf.int = TRUE))
+
+#CCV SvO2
+RegionChisepSvO2.t <- tidy(t.test(newdata$Chisep_IntVein_CSvO2, newdata$Chisep_SSSVein_CSvO2,
+            paired = TRUE, conf.int = TRUE))
 ```
@@
 ```{r}
+
+label = sigfunc(RegionQSMChi.t$p.value)
  p1 <- ggwithinstats(
@@
   xlab = "Region",
-  ylab = "Chi",
+  ylab = "Chi (ppm)",
   centrality.plotting = FALSE,
@@
   point.args = list(color = "#D95F02")) + 
-  ggplot2::annotate("text", x=1.5, y=(max(comp_legit_starfsl$Chi)*1.05), size=4, label = "***")
+  ggplot2::annotate("text", x=1.5, y=(max(comp_legit_starfsl$Chi)*1.05), size=4, label = label)
 
+label = sigfunc(RegionQSMSvO2.t$p.value)
 p2 <- ggwithinstats(
@@
   xlab = "Region",
-  ylab = "SvO2",
+  ylab = "SvO2 (%)",
   centrality.plotting = FALSE,
@@
 ) + 
-  ggplot2::annotate("text", x=1.5, y=(max(comp_SSS_IntVein_CSvO2$CSvO2)*1.05), size=4, label = "***")
+  ggplot2::annotate("text", x=1.5, y=(max(comp_SSS_IntVein_CSvO2$CSvO2)*1.05), size=4, label = label)
 
+label = sigfunc(RegionChisepChi.t$p.value)
 p3 <- ggwithinstats(                     #boxplot of chi comparison between regions
@@
 ) +
-  ggplot2::annotate("text", x=1.5, y=(max(comp_chisep_chi$Chi))*1.05, size=4, label = "n.s.")
+  ggplot2::annotate("text", x=1.5, y=(max(comp_chisep_chi$Chi))*1.05, size=4, label = label)
 
+label = sigfunc(RegionChisepSvO2.t$p.value)
 p4 <- ggwithinstats(                    #boxplot of svo2 comparison between regions
@@
   xlab = "Region",
-  ylab = "SvO2",
+  ylab = "SvO2 (%)",
   type = "nonparametric",
@@
 ) +
-  ggplot2::annotate("text", x=1.5, y=(max(comp_chisep_CSvO2$CSvO2)*1.05), size=4, label = "n.s.")
+  ggplot2::annotate("text", x=1.5, y=(max(comp_chisep_CSvO2$CSvO2)*1.05), size=4, label = label)
 
-(patch1 <- (p1 + p2)/(p3 + p4))
-# patch1
-# ggsave('figure.png', patch1)
+(patch2 <- (p1 + p2)/(p3 + p4) + 
+  plot_annotation(tag_levels = 'A') & 
+  theme(plot.tag = element_text(size = 10)))
 ```
@@
 #| fig-height: 5
-#| fig-cap: "rmarkdown test figure"
-patch1
+#| fig-cap: "Inter-venous comparisons of susceptibility and oxygen saturation.
+Violin plots comparing (A, C) χ (ppm) and (B, D) SvO2 (%) between the CCV and the SSS.
+Panels (A) and (B) used data acquired from QSM, and its raw data points are shown as filled orange circles.
+Panels (C) and (D) used data acquired from paramagnetic maps, and its raw data points are shown as filled green circles.
+Each line connects the raw data points of a single subject.
+(***) indicates p<0.05; (n.s.) indicates no significant difference."
+patch2
 ```
+
+```{r}
+save.image('results.RData')
+```

### 2024-10-03 index.qmd
Final draft

@@
 title: "The Application of Magnetic Susceptibility Separation for Measuring Cerebral Oxygenation in Preterm Neonates"
-titlerunning: "CSVO2"
-author:
+titlerunning: "SVO2"
+authors:
   - name: Thomas Gavin Carmichael
@@
     postal-code: V6T 1Z4
+  - id: 2
+    department: BC Children's Hospital Research Institute
+    name: The University of British Columbia
+    address: 938 West 28th Avenue
+    city: Vancouver
+    state: BC
+    country: Canada
+    postal-code: V5Z 4H4
   - id: 3
@@
     postal-code: V6T 1Z4
-  - id: 2
-    department: BC Children's Hospital Research Institute
-    name: The University of British Columbia
-    address: 938 West 28th Avenue
-    city: Vancouver
-    state: BC
-    country: Canada
-    postal-code: V5Z 4H4
 
@@
 number-sections: false
-notebook-links: true
+notebook-links: false
+# csl: sn-jnl.csl
+filters:
+  - authors-block
+crossref:
+  fig-title: '**Figure**'
+  tbl-title: '**Table**'
+  title-delim: "**.**"
 ---
@@
 
+## Acknowledgments
+
+We wish to acknowledge the work of and thank Victoria Tapics (Research Nurse); Vicki Goh (Research Nurse); Chacko Anil (Neonatologist); and Michael A. Sargent (Radiologist).
+
+## Funding
+
+Authors AMW and REG were co-primary applicants for a BC Children’s Hospital Research Institute - Brain, Behaviour and Development Catalyst Grant ($20,000).
+AMW was supported by an establishment award from BCCHRI.
+Scanning was partly funded through a special award to AMW from BCCHRI .
+{{< pagebreak >}}
+
 # Introduction {#sec-intro}
@@
 
-Preterm birth
-
-Abnormal brain development is a significant concern for parents with children born preterm, as 43% of infants that survive will have neurodevelopmental delays later in life [@dixMonitoringCerebralOxygenation2017].
+With advances in neonatal medical care, more infants born preterm are surviving into childhood [@mckenzieScaffoldingParentingHealth2022].
+These children are at high risk of acquiring adverse neurodevelopmental outcomes when compared to their term-born peers [@twilhaarCognitiveOutcomesChildren2018].
 Irregularities in early cerebral oxygen levels have been identified as a potential source of such delays, where too little oxygen provided during NICU care can result in white matter injury, while too much oxygen can result in reduced cortical connectivity [@rantakariEarlyOxygenLevels2021].
@@
 
-In the present study, we set out to determine whether a QSM image alone, or the paramagnetic component of the QSM image, is more accurate in measuring the oxygen present in the major cerebral veins of a cohort of preterm neonates.
-As previous QSM studies have not included the SSS, we also had a secondary aim of preserving this vessel in our QSM images and using this data to determine whether a difference in oxygenation existed between the SSS and the central cerebral veins (CCV).
+Unfortunately, there exist many challenges in measuring cerebral oxygen metabolism in neonates.
+Cerebral metabolic rate of oxygen (CMRO~2~) can be measured in humans using oxygen-15 positron emission tomography (PET) [@mintunBrainOxygenUtilization1984], has been measured in infants [@altmanCerebralBloodFlow1988], and is considered the gold standard.
+However, this method is invasive, requiring ionizing radiation, which limits its suitability for neonates.
+A less invasive option for evaluating brain hemodynamics is near-infrared spectroscopy (NIRS), which uses the attenuation of near-infrared light (\~650–950 nm) as it passes through biological tissue [@skovEstimationCerebralVenous1993].
+Deoxygenated and oxygenated hemoglobin absorb this light differently, allowing NIRS to estimate changes in deoxyhemoglobin and oxyhemoglobin [@wrayCharacterizationInfraredAbsorption1988] and thus provide an estimate of cerebral venous oxygen saturation (SvO~2~).
+While NIRS offers the advantage of being non-invasive and continuous bedside monitoring, it is limited to regional assessments where the probe is placed and is sensitive only to superficial brain tissue due to the shallow penetration depth of near-infrared light [@boasDiffuseOpticalImaging2004].
+
+For the preceding reasons, non-invasive MRI-based techniques are actively being explored to assess whole-brain oxygenation.
+MRI based methods have been developed for adults [@jainRapidMagneticResonance2011; @luQuantitativeEvaluationOxygenation2008; @xuNoninvasiveQuantificationWholebrain2009] and are only beginning to be explored in neonates [@devisNoninvasiveMRIMeasurements2014; @liuQuantitativeAssessmentGlobal2014; @qiHemodynamicMetabolicAssessment2018; @jainRapidMagneticResonance2011; @jiangVesselspecificQuantificationNeonatal2019].
+One limitation of these methods, however, is the fact that SvO~2~ is often measured using a single slice, averaging values within several voxels in the superior sagitall sinus (SSS).
+An alternative MRI method using quantiative susceptibility mapping (QSM) has been proposed, which measured SvO~2~ across the whole-brain.
+However, this method left room for improvement, as it removed the SSS, and required an arbitraty threshold value of 0.15 ppm in order to acquire realistic results [@weberQuantitativeSusceptibilityMapping2021].
+
+In the present study, we set out to determine whether decomposing the QSM image into its paramagnetic and diamagnetic comoponents would allow for a more accurate assessment of SvO~2~ in the central cerebral veins (CCV) of a cohort of preterm neonates.
+We also had a secondary aim of preserving the SSS vessel in our QSM images and using this data to determine whether a difference in oxygenation existed between the SSS and the CCV.
 
@@
 
-Participant data comes from a previous study [@zhu-etal-cmro2].
+Participant data comes from a previous study *** <!-- [@zhu-etal-cmro2] (submitted) -->.
 Participants consisted of preterm neonates born between 25- and 31-weeks gestational age (GA) who were admitted to the level III NICU at BC Women’s Hospital.
@@
 #| label: tbl-mri
-#| tbl-cap: "Technical parameters for MR imaging pulse sequences"
+#| tbl-cap: "**Technical parameters for MR imaging pulse sequences**"
 df <- data.frame(Scan = character(), T1w = character(), T2w = character(), pcASL = character(), SWI = character(), stringsAsFactors = FALSE)
@@
 
-The raw DICOM files acquired from the scanning procedure were converted to NIfTI (Neuroimaging Informatics Technology Initiative) format using Chris Rorden’s `dcmniix` tool.
-SWI magnitude data files were then used to create subject-specific brain masks that would not erode the SSS during QSM processing, an issue faced by our group in the past [@liFirstStepNeuroimaging2016].
+The raw DICOM files acquired from the scanning procedure were converted to NIfTI (Neuroimaging Informatics Technology Initiative) format using Chris Rorden’s `dcmniix` tool [@liFirstStepNeuroimaging2016].
+SWI magnitude data files were then used to create subject-specific brain masks that would not erode the SSS during QSM processing, an issue faced by our group in the past *** <!-- [@weberQuantitativeSusceptibilityMapping2021] -->.
 A step-by-step summary of the pipeline used is shown in @fig-graph.
@@
 6.0.7.3) [@woolrichBayesianAnalysisNeuroimaging2009] `fslroi`, `fslmaths`, and `bet` [@smithFastRobustAutomated2002] to create a preliminary brain mask, similar to our previous efforts, which does not contain the SSS.
-`Fslroi` was used to isolate the fifth echo of the magnitude data, which was then squared using Fslmaths and the option `-sqr`.
+`Fslroi` was used to isolate the fifth echo of the magnitude data, which was then squared using `fslmaths` and the option `-sqr`.
 Squaring the magnitude image was found to dramatically improve subsequent brain extraction.
@@
 The former flag generated a binary brain mask, while the latter performed a more robust brain centre estimation.
-The brain mask was then dilated by 7 voxels using `Fslmaths` and the options `-kernel boxv` and `-dilM` in order for the dilated mask to contain the SSS (along with unwanted tissue as well).
+The brain mask was then dilated by 7 voxels using `fslmaths` and the options `-kernel boxv` and `-dilM` in order for the dilated mask to contain the SSS (along with unwanted tissue as well).
 This mask was then used, along with the phase images, in a MATLAB script for QSM calculation from Christian Kames [@kamesRapidTwostepDipole2018] to produce a preliminary QSM image that contained the SSS, albeit with fairly low signal-to-noise ratio and other unwanted tissue.
@@
 STI Suite (v.
-3.0) [@liIntegratedLaplacianBased2014], was used to process the final QSM images as it produced the cleanest images without eroding the SSS.
+3.0) [@liIntegratedLaplacianbasedPhase2014], was used to process the final QSM images as it produced the cleanest images without eroding the SSS.
 The finalized brain mask and the last three echoes of the magnitude and phase images were used in STI Suite along with the following parameters: 0.9766 x 0.9766 x 1 mm^3^ voxel size, 5 ms TE1, 5.3 ms $\Delta$TE, and 77.4 ms sum TE, B0 strength = 3, and B0 direction = \[0, 0, 1\].
@@
 #| label: tbl-dem
-#| tbl-cap: Demographic and clinical characteristic of the study sample.
+#| tbl-cap: "**Demographic and clinical characteristic of the study sample.**"
 
@@
     linesep = "",
-    align = "lc",
+    align = "lr",
     escape = FALSE) |>
@@
     gt() |> 
-    cols_align(align = "center", columns = c("Subject data (n = 19)")) |>
+    cols_align(align = "right", columns = c("Subject data (n = 19)")) |>
     tab_footnote(footnotetext) |>
@@
 #| label: tbl-chistats
-#| tbl-cap: Summary of acquired physiological parameters.
-Mean $\pm$ SD is shown for chi and SvO~2~ values.
+#| tbl-cap: \textbf{Summary of acquired physiological parameters.} Mean $\pm$ SD is shown for chi and SvO~2~ values.
 The P-value and 95% confidence interval (CI) were obtained through the comparison of values between QSM and paramagnetic maps; (n=19).
@@
 
-The primary objective of the present study was to assess whether the application of magnetic susceptibility separation to neonatal QSM data could provide more accurate cerebral venous vessel oxygenation measurements.
+The primary objective of the present study was to assess whether the application of magnetic susceptibility separation to neonatal QSM data could provide more accurate SvO~2~ measurements without the need for an arbitrary threshold value.
 To the best of our knowledge, we are the first to test this in a neonatal cohort, as susceptibility separation has been typically evaluated as a method of imaging myelin and brain iron in adult subjects [@shinHseparationMagneticSusceptibility2021; @ahmedDiamagneticComponentMap2023a].
@@
 
+## Comparison with literature values
+
+To evaluate the validity of our results, we compared the mean SvO~2~ values we obtained through QSM and susceptibility separation to the mean SvO~2~ values found by MRI studies investigating the oxygenation of the SSS or the CCV.
+Notably, the number of studies that measure the SvO~2~ of the CCV, or any of its individual veins, in infants is fairly lower than the number of studies investigating the oxygenation of the SSS.
+Our comparison is summarized in @tbl-litvalues.
+
+```{r}
+#| label: tbl-litvalues
+#| tbl-cap: "**Cerebral oxygenation values of neonates and fetuses in the literature.**"
+
+df <- data.frame(Region = character(), Study = character(), Subjects = character(), Method = character(), SvO2 = character(), stringsAsFactors = FALSE)
+df[1,] <- c("Whole-brain", "Skov et al. \n(1993)    ", "Preterm neonates \n(n=9)                   ", "NIRS            ", "53.4 \U00B1 15.4 ")
+df[2,] <- c("Whole-brain", "Skov et al. \n(1993)    ", "Asphyxiated term \nneonates (n=10)           ", "NIRS            ", "67.3 \U00B1 9.4  ")
+df[3,] <- c("Whole-brain", "Altman et al. \n(1993)    ", "Preterm and term \nneonates with HIE \n(n=11)", "PET            ", "21.6 \U00B1 21.0 ")
+df[4,] <- c("SSS        ", "Gou et al. \n(2024)     ", "Healthy neonates \n(n=37)                    ", "MRI: \nTRUST   ", "66.7 \U00B1 4.9  ")
+df[5,] <- c("SSS        ", "Jiang et al. \n(2019)   ", "Healthy neonates \n(n=4)                     ", "MRI: \naTRUPC  ", "64.8 \U00B1 2.9  ")
+df[6,] <- c("SSS        ", "De Vis et al. \n(2014)  ", "PT-TEA \neonates (n=18)                      ", "MRI: \nT2-TRIR ", "52.0 \U00B1 12.0 ")
+df[7,] <- c("SSS", "Yadav et al. \n(2020)", "Late third \ntrimester fetuses \n(n=33)", "MRI: \nSusceptometry", "58.6 \U00B1 4.8")
+df[8,] <- c("SSS", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \nQSM", "72.5 \U00B1 3.4")
+df[9,] <- c("SSS", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \n$\\chi$-separation", "58.1 \U00B1 7.3")
+df[10,] <- c("CCV", "Weber et al. (2021)", "Preterm neonates \nwith HIE (n=8)", "MRI: \nQSM", "72.2 \U00B1 6.0")
+df[11,] <- c("CCV", "Weber et al. (2021)", "Healthy neonates \n(n=8)", "MRI: \nQSM", "73.6 \U00B1 2.8")
+df[12,] <- c("CCV", "Jiang et al. (2019)", "Healthy neonates \n(n=4)", "MRI: \naTRUPC", "70.2 \U00B1 3.3")
+df[13,] <- c("CCV", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \nQSM", "68.7 \U00B1 3.5")
+df[14,] <- c("CCV", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \n$\\chi$-separation", "57.7 \U00B1 7.0")
+footnotetext="PT-TEA = born preterm and scanned at term-equivalent age; late third trimester = >35 weeks gestational age; HIE = hypoxic-ischemic encephalopathy; TRUST = T2-relaxation-under-spin tagging; aTRUPC = accelerated T2-relaxation-under-phase-contrast; T2-TRIR = T2-tissue-relaxation-inversion-recovery; SWI = susceptibility weighted imaging."
+
+if (knitr::is_latex_output()) {
+df <- df |> rename("SvO$_{2}$ (\\\U0025)" = SvO2)
+df[1,] <- linebreak(df[1,])
+df[2,] <- linebreak(df[2,])
+df[3,] <- linebreak(df[3,])
+df[4,] <- linebreak(df[4,])
+df[5,] <- linebreak(df[5,])
+df[6,] <- linebreak(df[6,])
+df[7,] <- linebreak(df[7,])
+df[8,] <- linebreak(df[8,])
+df[9,] <- linebreak(df[9,])
+df[10,] <- linebreak(df[10,])
+df[11,] <- linebreak(df[11,])
+df[12,] <- linebreak(df[12,])
+df[13,] <- linebreak(df[13,])
+df[14,] <- linebreak(df[14,])
+df |>
+  kbl(format = "latex",
+    booktabs = TRUE,
+    longtable = TRUE,
+    linesep = "",
+    align = "ccllr",
+    escape = FALSE) |>
+    # collapse_rows() |>
+    kable_styling(font_size = 9, position = "center", latex_options = c("hold_position", "scale_down"), full_width = FALSE) |>
+    footnote(general_title = "", 
+           footnote_as_chunk = TRUE,
+           threeparttable = TRUE,
+           general = footnotetext)
+} else {
+df <- df |> rename("SvO~2~ (\\\U0025)" = SvO2)
+df |> 
+    mutate(across(everything(), ~ str_replace_all(., "\n", "<br>"))) |> 
+    gt() |> 
+    cols_align(align = "center", columns = c("Region", "Study")) |>
+    tab_footnote(footnotetext) |>
+    fmt_markdown(columns = TRUE) |>
+    tab_options(quarto.disable_processing = TRUE)
+}
+```
+As shown in @tbl-litvalues, the infants observed in MRI studies investigating cerebral vein oxygenation noticeably differ in clinical status, with three studies involving healthy neonates [@weberQuantitativeSusceptibilityMapping2021; @gouAutomaticRejectionBased2024; @jiangVesselspecificQuantificationNeonatal2019], three studies (including the present study) involving preterm neonates [@weberQuantitativeSusceptibilityMapping2021; @devisNoninvasiveMRIMeasurements2014], and one study involving late third trimester fetuses [@portnoyHumanUmbilicalCord2018].
+In the studies involving healthy neonates, the SvO~2~ of the SSS fell within the range of 64.8% – 66.6% [@gouAutomaticRejectionBased2024; @jiangVesselspecificQuantificationNeonatal2019], while the SvO~2~ of the CCV fell within the range of 70.2% - 73.6% [@weberQuantitativeSusceptibilityMapping2021; @jiangVesselspecificQuantificationNeonatal2019].
+Notably, the SvO~2~ value of the SSS we obtained from susceptibility separation (58.1%) was closest to values obtained from the studies involving late third trimester fetuses [@yadavImagingPutativeFoetal2018] or pre-term neonates [@devisNoninvasiveMRIMeasurements2014], each finding an SSS SvO~2~ value of 58.6% and 52.0%, respectively.
+It is important to note the difference in MRI modalities used to obtain these values.
+For their study, @yadavImagingPutativeFoetal2018 used MR susceptometry, which involves measuring the difference in phase between the chosen vessel and its background in imaging data from an SWI scanning sequence [@yadavImagingPutativeFoetal2018].
+In @devisNoninvasiveMRIMeasurements2014, the authors used T2-TRIR, which allowed them to determine the transverse relaxation rate of blood within the vessel, which can be used alongside hematocrit data to estimate SvO~2~.
+Additionally, the GA of infants scanned in our study ranged between 37 and 44 weeks, while the GA of the fetuses scanned in @yadavImagingPutativeFoetal2018 was ≥35 weeks and the GA of infants scanned in @devisNoninvasiveMRIMeasurements2014 ranged between 38 and 40 weeks.
+As such, our SSS SvO~2~ values found through susceptibility separation show promise given their similarity to the SvO~2~ values found by @yadavImagingPutativeFoetal2018 and @devisNoninvasiveMRIMeasurements2014, two studies that involved comparable subject populations and used considerably different methods.
+
+Conversely, the SvO~2~ value of the CCV we obtained through QSM (68.7%) was closest to values from similar studies in the literature.
+In @weberQuantitativeSusceptibilityMapping2021, QSM was used to measure an SvO~2~ of 71.5% in preterm neonates with HIE and an SvO~2~ of 73.6% in healthy neonates.
+In their study, @jiangVesselspecificQuantificationNeonatal2019 also involved healthy neonates and obtained an SvO~2~ of 70.2% through an accelerated TRUPC sequence.
+In contrast, the SvO~2~ of the CCV found through susceptibility separation was 57.7%.
+This disparity from the literature, however, may not undermine the value we obtained, as the study design of @weberQuantitativeSusceptibilityMapping2021 and @jiangVesselspecificQuantificationNeonatal2019 may prevent their values from being representative of the study demographic.
+In @weberQuantitativeSusceptibilityMapping2021, the authors utilized an arbitrary 0.15 ppm threshold and included all $\chi$ values above 0.15 when measuring the mean $\chi$ of the CCV, which potentially led to the introduction of $\chi$ from veins outside the CCV.
+In @jiangVesselspecificQuantificationNeonatal2019, the authors acquired their data from 4 subjects, a notably small sample size.
+Given the limitations of the existing literature and the similarity of the mean paramagnetic CCV SvO~2~ value (57.7%) to the mean paramagnetic SSS SvO~2~ value (58.1%), it is plausible that susceptibility separation provides accurate measurements of oxygenation in both cortical and subcortical veins.
+
+Another notable distinction between our findings and those of the existing literature was that we observed no significant oxygenation difference between the SSS and the CCV when $\chi$ was derived from paramagnetic maps.
+@jiangVesselspecificQuantificationNeonatal2019, the only other study that also measured SvO~2~ in both the SSS and CCV, observed significantly lower oxygenation in the SSS (64.8%) when compared to the CCV (70.2%).
+Given the small sample size utilized by @jiangVesselspecificQuantificationNeonatal2019, it is difficult to ascertain whether this is generalizable to all neonates.
+
+## Limitations and future directions
+
+This study has a few limitations that should be considered for future research.
+Firstly, only 19 infants were recruited for scanning.
+Given the emotional toll placed on parents when their child is born preterm, it is understandable that they may show reluctance in consenting to further testing that is not medically necessary.
+Obtaining a larger sample size in future studies, however, may provide greater insight into the efficacy of susceptibility separation.
+Secondly, this study did not include healthy neonates born at term, resulting in a lack of a control cohort.
+The addition of such a group may provide further validity to any findings and may reveal potential differences in cerebral oxygen consumption between term and preterm neonates.
+Finally, future studies should consider the use of multi-echo T2 imaging data when performing the decomposition of QSM images.
+The toolbox applied by this study for QSM decomposition [@shinHseparationMagneticSusceptibility2021] utilizes R2 data, which can be obtained from multi-echo T2 imaging.
+Our study protocol involved the collection of multi-echo SWI imaging data, and as such, we could only use R2^*^ data to perform the decomposition.
+Furthermore, this may account for the reduced precision of SvO~2~ values obtained through susceptibility separation.
+
 # Conclusion {#sec-conclusion}
+
+This study aimed to evaluate how the use of susceptibility separation on preterm neonatal QSM images in order to  provide greater accuracy in determining the oxygenation of cerebral venous vessels.
+We compared venous specific SvO~2~ values obtained from QSM images and their respective paramagnetic components to SvO~2~ values from neonatal MRI studies.
+We found that susceptibility separation provided SvO~2~ values of the SSS that were comparable to values found in the literature, providing evidence that this processing technique may be a valid tool for measuring regional cerebral oxygen consumption.
+Additionally, we were able to simultaneously measure SvO~2~ in both the SSS and CCV, which permitted us to observe no difference in oxygenation between the two vessels when considering data from isolated paramagnetic components.
+Ultimately, we hope our work inspires future studies that seek to explore and improve the capabilities of magnetic susceptibility separation, culminating in the development of a tool for clinicians and researchers alike.
 

### 2024-10-03 index.qmd
fixed spelling errors

@@
 MRI based methods have been developed for adults [@jainRapidMagneticResonance2011; @luQuantitativeEvaluationOxygenation2008; @xuNoninvasiveQuantificationWholebrain2009] and are only beginning to be explored in neonates [@devisNoninvasiveMRIMeasurements2014; @liuQuantitativeAssessmentGlobal2014; @qiHemodynamicMetabolicAssessment2018; @jainRapidMagneticResonance2011; @jiangVesselspecificQuantificationNeonatal2019].
-One limitation of these methods, however, is the fact that SvO~2~ is often measured using a single slice, averaging values within several voxels in the superior sagitall sinus (SSS).
-An alternative MRI method using quantiative susceptibility mapping (QSM) has been proposed, which measured SvO~2~ across the whole-brain.
-However, this method left room for improvement, as it removed the SSS, and required an arbitraty threshold value of 0.15 ppm in order to acquire realistic results [@weberQuantitativeSusceptibilityMapping2021].
+One limitation of these methods, however, is the fact that SvO~2~ is often measured using a single slice, averaging values within several voxels in the superior sagittal sinus (SSS).
+An alternative MRI method using quantitative susceptibility mapping (QSM) has been proposed, which measured SvO~2~ across the whole-brain.
+However, this method left room for improvement, as it removed the SSS, and required an arbitrary threshold value of 0.15 ppm in order to acquire realistic results [@weberQuantitativeSusceptibilityMapping2021].
 
-In the present study, we set out to determine whether decomposing the QSM image into its paramagnetic and diamagnetic comoponents would allow for a more accurate assessment of SvO~2~ in the central cerebral veins (CCV) of a cohort of preterm neonates.
+In the present study, we set out to determine whether decomposing the QSM image into its paramagnetic and diamagnetic components would allow for a more accurate assessment of SvO~2~ in the central cerebral veins (CCV) of a cohort of preterm neonates.
 We also had a secondary aim of preserving the SSS vessel in our QSM images and using this data to determine whether a difference in oxygenation existed between the SSS and the CCV.

### 2024-10-11 index.qmd
Submission to Pediatric Research

@@
   - id: 1
-    name: The University of British Columbia
-    department: Integrated Sciences
-    address: 2329 West Mall
-    city: Vancouver
-    region: BC
-    country: Canada
-    postal-code: V6T 1Z4
+    name: Integrated Sciences, The University of British Columbia, Vancouver, BC, Canada
   - id: 2
-    department: BC Children's Hospital Research Institute
-    name: The University of British Columbia
-    address: 938 West 28th Avenue
-    city: Vancouver
-    state: BC
-    country: Canada
-    postal-code: V5Z 4H4
+    name: BC Children's Hospital Research Institute, The University of British Columbia, Vancouver, BC, Canada
   - id: 3
-    name: The University of British Columbia
-    department: Pediatrics
-    address: 2329 West Mall
-    city: Vancouver
-    region: BC
-    country: Canada
-    postal-code: V6T 1Z4
+    name: Pediatrics, The University of British Columbia, Vancouver, BC, Canada
 
@@
   - Cerebral Venous Oxygen Saturation
+
 abstract: |
-  **Background**: Quantitative susceptibility mapping (QSM) is a magnetic resonance imaging (MRI) modality proposed to be a viable method of measuring cerebral oxygenation in neonates given its sensitivity to deoxyhemoglobin, a paramagnetic molecule.
-During QSM, however, paramagnetic sources can be obscured by opposing diamagnetic sources such as water and myelin.
-We sought to evaluate whether QSM images alone, or an algorithm that attempts to isolate their paramagnetic components, are more accurate in measuring oxygenation of the major cerebral veins in a cohort of neonates born preterm.
-Additionally, we aimed to determine whether a difference in oxygenation existed between the major cerebral veins.
+  **Background**: Quantitative susceptibility mapping (QSM), an magnetic resonance imaging (MRI) modality sensitive to deoxyhemoglobin, is a promising method for measuring cerebral oxygenation in human neonates.
+Paramagnetic sources, like deoxyhemoglobin, however, can be obscured by diamagnetic sources such as water and myelin.
+This study evaluated whether QSM images, or isolated paramagnetic components, are more accurate for measuring oxygenation of cerebral veins of preterm neonates, and explored oxygenation differences between the major cerebral veins.
 
-  **Methods**: 19 neonates born preterm were scanned on a 3T research MRI at term equivalent age.
-The protocol included a multi-echo susceptibility-weighted imaging sequence.
-The acquired imaging data were processed as QSM images to obtain the susceptibility values of the superior sagittal sinus (SSS) and central cerebral veins (CCV).
-These values were used to calculate the oxygen saturation (SvO~2~) of the SSS and CCV.
-QSM images were subsequently processed to isolate their paramagnetic components.
-SvO~2~ values of the SSS and CCV were calculated again from the paramagnetic components.
+  **Methods**: 19 preterm neonates were scanned on at term equivalent age on a 3T MRI using a multi-echo susceptibility-weighted imaging sequence.
+Susceptibility values were calculated from QSM images to determine oxygen saturation (SvO~2~) in the superior sagittal sinus (SSS) and central cerebral veins (CCV).
+The paramagnetic components of QSM images were isolated, and SvO~2~ values were recalculated.
 
-  **Results**: The mean SvO~2~ values of the SSS and CCV calculated from QSM images were found to be 72.4% (SD, 3.4%) and 68.7% (SD, 3.5%), respectively.
-The mean SvO~2~ values calculated from paramagnetic components were found to be 58.1% (SD, 7.3%) for the SSS and 57.7% (SD, 7.0%) for the CCV.
+  **Results**: The mean SvO~2~ values from QSM were 72.4% (SD, 3.4%) for the SSS and and 68.7% (SD, 3.5%) for the CCV.
+SvO~2~ values for paramagnetic components were 58.1% (SD, 7.3%) for the SSS and 57.7% (SD, 7.0%) for the CCV.
 
-  **Conclusion**: SSS SvO~2~ values derived from paramagnetic components agreed well with the existing literature and were closer than the values derived from QSM, however, they displayed greater variability.
-Although the CCV SvO~2~ data from QSM aligns more closely with existing literature, it is important to note that the current literature on this topic remains relatively limited in the CCV.
-Thus, decomposing QSM images into paramagnetic components shows great promise as a method for more accurately measuring cerebral oxygenation in neonates but may require more research to improve precision.
-Notably, no significant difference in oxygenation was observed between the CCV and the SSS, contrasting with previous studies.
+  **Conclusion**: While paramagnetic component decomposition yielded more accurate SSS values, it increased variability.
+No significant oxygenation differences were found between the SSS and CCV, contrasting with prior studies.
 
-plain-language-summary: |
-
-key-points:
-  - 
-date: last-modified
-bibliography: [ Gavin_Thesis_Ref.bib ]
-citation:
-  container-title: Unpublished
-number-sections: false
-notebook-links: false
-# csl: sn-jnl.csl
 filters:
   - authors-block
+
+bibliography: Gavin_Thesis_Ref.bib
+csl: pediatric-research.csl
+
 crossref:
@@
 ---
+
+**Category of Study:** basic science
+
+## Impact Statement
+
+- This study evaluated the accuracy of using QSM and its paramagnetic components to measure cerebral oxygenation in neonates.
+- By comparing susceptibility-derived oxygen saturation (SvO~2~) in the superior sagittal sinus (SSS) and central cerebral veins (CCV), it adds to the field of neonatal cerebral oxygenation measurement.
+- Decomposing QSM into paramagnetic components shows potential for improving SvO~2~ accuracy, particularly in the SSS, though variability remains a challenge.
+- The results suggest no significant oxygenation difference between the SSS and CCV, contrasting with previous findings, indicating a need for further research on neonatal venous oxygenation.
 
@@
 Scanning was partly funded through a special award to AMW from BCCHRI .
+
+## Author contributions
+
+TGC wrote the original draft, performed the formal analysis, and contributed to methodology, validation, and visualization.
+AR helped with writing, reviewing & editing.
+REG helped with writing, reviewing & editing, and with initial funding acquisition.
+AMW was involved in project administration, supervision, validation, visualization, resources, methodology, formal analysis, funding acquisition, writing, reviewing & editing, conceptualization, data curation, and investigation.
+
+## Competing interests
+
+The authors have no competing interests to declare.
+
+## Consent statement
+
+The study was approved by the Clinical Research Ethics Board at the University of British Columbia and Children's & Women's Hospital (H21-00655) and written informed consent was obtained from the parent/guardian for each infant.
+
 {{< pagebreak >}}
@@
 Unfortunately, there exist many challenges in measuring cerebral oxygen metabolism in neonates.
-Cerebral metabolic rate of oxygen (CMRO~2~) can be measured in humans using oxygen-15 positron emission tomography (PET) [@mintunBrainOxygenUtilization1984], has been measured in infants [@altmanCerebralBloodFlow1988], and is considered the gold standard.
+Cerebral metabolic rate of oxygen (CMRO~2~) using oxygen-15 positron emission tomography (PET) [@mintunBrainOxygenUtilization1984], has been measured in infants [@altmanCerebralBloodFlow1988], and is considered the gold standard.
 However, this method is invasive, requiring ionizing radiation, which limits its suitability for neonates.
@@
 
-For the preceding reasons, non-invasive MRI-based techniques are actively being explored to assess whole-brain oxygenation.
-MRI based methods have been developed for adults [@jainRapidMagneticResonance2011; @luQuantitativeEvaluationOxygenation2008; @xuNoninvasiveQuantificationWholebrain2009] and are only beginning to be explored in neonates [@devisNoninvasiveMRIMeasurements2014; @liuQuantitativeAssessmentGlobal2014; @qiHemodynamicMetabolicAssessment2018; @jainRapidMagneticResonance2011; @jiangVesselspecificQuantificationNeonatal2019].
-One limitation of these methods, however, is the fact that SvO~2~ is often measured using a single slice, averaging values within several voxels in the superior sagittal sinus (SSS).
-An alternative MRI method using quantitative susceptibility mapping (QSM) has been proposed, which measured SvO~2~ across the whole-brain.
-However, this method left room for improvement, as it removed the SSS, and required an arbitrary threshold value of 0.15 ppm in order to acquire realistic results [@weberQuantitativeSusceptibilityMapping2021].
+For the preceding reasons, non-invasive MRI-based techniques are actively being explored to assess regional and whole-brain blood oxygenation.
+MRI based methods have been developed for adults [@jainRapidMagneticResonance2011; @luQuantitativeEvaluationOxygenation2008; @xuNoninvasiveQuantificationWholebrain2009] and are only beginning to be explored in neonates [@devisNoninvasiveMRIMeasurements2014; @liuQuantitativeAssessmentGlobal2014; @qiHemodynamicMetabolicAssessment2018; @jainCerebralOxygenMetabolism2014; @jiangVesselspecificQuantificationNeonatal2019].
+These methods almost all rely on T2 relaxation to estimate CSvO~2~ [@devisNoninvasiveMRIMeasurements2014; @liuQuantitativeAssessmentGlobal2014; @qiHemodynamicMetabolicAssessment2018; @jiangVesselspecificQuantificationNeonatal2019] with the exception of @jainCerebralOxygenMetabolism2014 which used susceptometry [@jainMRIEstimationGlobal2010].
+One limitation of these T2 relaxation methods, however, is the fact that SvO~2~ is often measured using a single imaging slice, averaging values across several voxels, and only in the superior sagittal sinus (SSS).
+In the case of @jainCerebralOxygenMetabolism2014, they obtained regional and whole-brain data, but with thick slices (5mm), and still only estimated CSvO~2~ in the SSS.
+An alternative MRI method using quantitative susceptibility mapping (QSM) has been proposed, which can measure SvO~2~ regionally and across the whole-brain at high resolution (< 1mm^3^ per voxel) [@weberQuantitativeSusceptibilityMapping2021].
+However, this method left room for improvement, as it removed the SSS (averaging CSvO~2~ across the internal veins), and required an arbitrary threshold value of 0.15 ppm in order to acquire realistic results [@weberQuantitativeSusceptibilityMapping2021].
 
@@
 Participant data comes from a previous study *** <!-- [@zhu-etal-cmro2] (submitted) -->.
-Participants consisted of preterm neonates born between 25- and 31-weeks gestational age (GA) who were admitted to the level III NICU at BC Women’s Hospital.
+Participants consisted of preterm neonates born between 25- and 31-weeks gestational age (GA) who were admitted to the level III NICU at *** <!-- BC Women’s Hospital -->.
 Recruitment took place over a span of one year, from February 2021 to January 2022, facilitated by a dedicated research nurse.
@@
 Infants meeting the criteria for inclusion were scanned for the study if they had already been discharged from the NICU, were in stable condition, and had reached a term equivalent age of 37 to 44 weeks GA.
-However, certain exclusion criteria were applied to ensure the homogeneity and integrity of the study sample.
-Infants were excluded if there was clinical evidence of a congenital malformation or syndrome, a TORCH infection, or ultrasound evidence of large parenchymal hemorrhagic infarction (\>2 cm, Grade 4 intraventricular hemorrhage).
+However, certain exclusion criteria were applied to ensure the homogeneity and integrity of the study sample: infants were excluded if there was clinical evidence of a congenital malformation or syndrome, a TORCH infection, or ultrasound evidence of large parenchymal hemorrhagic infarction (\>2 cm, Grade 4 intraventricular hemorrhage).
 
@@
 
-MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750 scanner (scanner software version DV26.0_R03) equipped with a SREE Medical Systems single-channel neonatal head coil (@tbl-mri).
-The scans were conducted at the BC Children’s MRI Research Facility.
+MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750 scanner (scanner software version DV26.0_R03) equipped with a SREE Medical Systems (Cleveland, OH, USA) single-channel neonatal head coil (@tbl-mri).
+The scans were conducted at the *** <!-- BC Children -->’s MRI Research Facility.
 Prior to the scanning procedure, subjects were carefully prepared by a research nurse.
@@
 df[3,] <- c("TR (ms)", "7.74", "2,300", "4,680", "30.9")
-df[4,] <- c("TE (ms)", "2.97", "66.29", "10.55", "5.24")
+df[4,] <- c("TE (ms)", "2.97", "66.29", "10.55", "5 echoes; first echo: 5; echo spacing: 5.24")
 df[5,] <- c("Flip angle", "12\U00B0", "90\U00B0", "111\U00B0", "20\U00B0")
@@
 
-The MRI scan protocol comprised of the following sequences: a T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin labeling (ASL) scan, a multi-echo susceptibility-weighted imaging scan, and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging (EPI) sequence.
+The MRI scan protocol comprised of the following sequences: a T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin labeling (ASL) scan [@alsopRecommendedImplementationArterial2015], a multi-echo susceptibility-weighted imaging scan [@denkSusceptibilityWeightedImaging2010], and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging (EPI) sequence.
 The DWI sequence was not used for the present study.
@@
 
-First, the fifth echo SWI magnitude file was processed using FSL’s (v.
+First, the fifth echo (TE = 25.96 ms) SWI magnitude file was processed using FSL’s (v.
 6.0.7.3) [@woolrichBayesianAnalysisNeuroimaging2009] `fslroi`, `fslmaths`, and `bet` [@smithFastRobustAutomated2002] to create a preliminary brain mask, similar to our previous efforts, which does not contain the SSS.
@@
 
-Once the mean susceptibility values of the SSS and CCV were obtained from the subjects’ QSM images and paramagnetic maps, venous oxygen saturation (SvO~2~) was calculated with the following equation [@bergInvestigatingEffectFlow2021]:
+Once the mean susceptibility values of the SSS and CCV were obtained from the subjects’ QSM images or paramagnetic maps, venous oxygen saturation (SvO~2~) was calculated with the following equation [@bergInvestigatingEffectFlow2021]:
 
@@
 
-Statistical analysis of the acquired data was performed using R and RStudio (v.
+Statistical analysis was performed using R and RStudio (v.
 2023.09.1 Build 494) [@rcoreteamLanguageEnvironmentStatistical2022; @rstudioteamRStudioIntegratedDevelopment].
@@
 
-The mean SvO~2~ values for the SSS and the CCV were found to be `r rndmean(newdata$Gavin_SSSVein_CSvO2)` $\pm$ `r rndsd(newdata$Gavin_SSSVein_CSvO2)` and `r rndmean(newdata$Gavin_IntVein_CSvO2)` $\pm$  `r rndsd(newdata$Gavin_IntVein_CSvO2)` ppm, respectively, when determined from the QSM data.
-When determined from the paramagnetic map, the mean SvO~2~ values for the SSS and the CCV were found to be `r rndmean(newdata$Chisep_SSSVein_CSvO2)` $\pm$ `r rndsd(newdata$Chisep_IntVein_CSvO2)` %, respectively.
+The mean SvO~2~ values for the SSS and the CCV were found to be `r rndmean(newdata$Gavin_SSSVein_CSvO2)` $\pm$ `r rndsd(newdata$Gavin_SSSVein_CSvO2)`% and `r rndmean(newdata$Gavin_IntVein_CSvO2)` $\pm$  `r rndsd(newdata$Gavin_IntVein_CSvO2)`%, respectively, when determined from the QSM data.
+When determined from the paramagnetic map, the mean SvO~2~ values for the SSS and the CCV were found to be `r rndmean(newdata$Chisep_SSSVein_CSvO2)` $\pm$ `r rndsd(newdata$Chisep_SSSVein_CSvO2)`% and `r rndmean(newdata$Chisep_IntVein_CSvO2)` $\pm$ `r rndsd(newdata$Chisep_IntVein_CSvO2)`%, respectively.
 A summary of the measured physiological parameters, including the chi values used to calculate SvO~2~, can found in @tbl-chistats.
@@
 Our results showed that the SvO~2~ values of the SSS and CCV obtained from susceptibility separation are significantly lower than the respective SvO~2~ values obtained from QSM alone.
-When our results were compared to the literature, we found that our SSS SvO~2~ data from susceptibility separation agreed well with the findings of other studies measuring SvO~2~ of the SSS in similar subject populations.
+When our results were compared to the literature (see below), we found that our SSS SvO~2~ data from susceptibility separation agreed well with the findings of other studies measuring SvO~2~ of the SSS in similar subject populations.
 Conversely, the paramagnetic CCV SvO~2~ data saw less agreement with the existing literature than the corresponding data from QSM.
@@
 df[7,] <- c("SSS", "Yadav et al. \n(2020)", "Late third \ntrimester fetuses \n(n=33)", "MRI: \nSusceptometry", "58.6 \U00B1 4.8")
-df[8,] <- c("SSS", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \nQSM", "72.5 \U00B1 3.4")
-df[9,] <- c("SSS", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \n$\\chi$-separation", "58.1 \U00B1 7.3")
+df[8,] <- c("SSS", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \nQSM", paste0(rndmean(newdata$Gavin_SSSVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Gavin_SSSVein_CSvO2*100)))
+df[9,] <- c("SSS", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \n$\\chi$-separation", paste0(rndmean(newdata$Chisep_SSSVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Chisep_SSSVein_CSvO2*100)))
 df[10,] <- c("CCV", "Weber et al. (2021)", "Preterm neonates \nwith HIE (n=8)", "MRI: \nQSM", "72.2 \U00B1 6.0")
@@
 df[12,] <- c("CCV", "Jiang et al. (2019)", "Healthy neonates \n(n=4)", "MRI: \naTRUPC", "70.2 \U00B1 3.3")
-df[13,] <- c("CCV", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \nQSM", "68.7 \U00B1 3.5")
-df[14,] <- c("CCV", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \n$\\chi$-separation", "57.7 \U00B1 7.0")
+df[13,] <- c("CCV", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \nQSM", paste0(rndmean(newdata$Gavin_IntVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Gavin_IntVein_CSvO2)))
+df[14,] <- c("CCV", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \n$\\chi$-separation", paste0(rndmean(newdata$Chisep_IntVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Chisep_IntVein_CSvO2*100)))
 footnotetext="PT-TEA = born preterm and scanned at term-equivalent age; late third trimester = >35 weeks gestational age; HIE = hypoxic-ischemic encephalopathy; TRUST = T2-relaxation-under-spin tagging; aTRUPC = accelerated T2-relaxation-under-phase-contrast; T2-TRIR = T2-tissue-relaxation-inversion-recovery; SWI = susceptibility weighted imaging."
@@
 In the studies involving healthy neonates, the SvO~2~ of the SSS fell within the range of 64.8% – 66.6% [@gouAutomaticRejectionBased2024; @jiangVesselspecificQuantificationNeonatal2019], while the SvO~2~ of the CCV fell within the range of 70.2% - 73.6% [@weberQuantitativeSusceptibilityMapping2021; @jiangVesselspecificQuantificationNeonatal2019].
-Notably, the SvO~2~ value of the SSS we obtained from susceptibility separation (58.1%) was closest to values obtained from the studies involving late third trimester fetuses [@yadavImagingPutativeFoetal2018] or pre-term neonates [@devisNoninvasiveMRIMeasurements2014], each finding an SSS SvO~2~ value of 58.6% and 52.0%, respectively.
+Notably, the SvO~2~ value of the SSS we obtained from susceptibility separation (`r paste0(rndmean(newdata$Chisep_SSSVein_CSvO2*100))`%) was closest to values obtained from the studies involving late third trimester fetuses [@yadavImagingPutativeFoetal2018] or pre-term neonates [@devisNoninvasiveMRIMeasurements2014], each finding an SSS SvO~2~ value of 58.6% and 52.0%, respectively.
 It is important to note the difference in MRI modalities used to obtain these values.
@@
 
-Conversely, the SvO~2~ value of the CCV we obtained through QSM (68.7%) was closest to values from similar studies in the literature.
+Conversely, the SvO~2~ value of the CCV we obtained through QSM (`r paste0(rndmean(newdata$Gavin_IntVein_CSvO2*100))`%) was closest to values from similar studies in the literature.
 In @weberQuantitativeSusceptibilityMapping2021, QSM was used to measure an SvO~2~ of 71.5% in preterm neonates with HIE and an SvO~2~ of 73.6% in healthy neonates.
 In their study, @jiangVesselspecificQuantificationNeonatal2019 also involved healthy neonates and obtained an SvO~2~ of 70.2% through an accelerated TRUPC sequence.
-In contrast, the SvO~2~ of the CCV found through susceptibility separation was 57.7%.
+In contrast, the SvO~2~ of the CCV we obtained through susceptibility separation was `r paste0(rndmean(newdata$Chisep_IntVein_CSvO2*100))`%.
 This disparity from the literature, however, may not undermine the value we obtained, as the study design of @weberQuantitativeSusceptibilityMapping2021 and @jiangVesselspecificQuantificationNeonatal2019 may prevent their values from being representative of the study demographic.
@@
 In @jiangVesselspecificQuantificationNeonatal2019, the authors acquired their data from 4 subjects, a notably small sample size.
-Given the limitations of the existing literature and the similarity of the mean paramagnetic CCV SvO~2~ value (57.7%) to the mean paramagnetic SSS SvO~2~ value (58.1%), it is plausible that susceptibility separation provides accurate measurements of oxygenation in both cortical and subcortical veins.
+Given the limitations of the existing literature and the similarity of the mean paramagnetic CCV SvO~2~ value (`r paste0(rndmean(newdata$Chisep_IntVein_CSvO2*100))`%) to the mean paramagnetic SSS SvO~2~ value (`r paste0(rndmean(newdata$Chisep_SSSVein_CSvO2*100))`%), it is plausible that susceptibility separation provides more accurate measurements of oxygenation in both cortical and subcortical veins.
+One reason for this is due to its ability to mitigate partial volume effects, which are likely to contaminate other methods resulting in inaccurate CSvO~2~ values [@shinHseparationMagneticSusceptibility2021].
 
@@
 Secondly, this study did not include healthy neonates born at term, resulting in a lack of a control cohort.
+This is because recruiting healthy controls when there is no contraindication is very difficult.
 The addition of such a group may provide further validity to any findings and may reveal potential differences in cerebral oxygen consumption between term and preterm neonates.
@@
 The toolbox applied by this study for QSM decomposition [@shinHseparationMagneticSusceptibility2021] utilizes R2 data, which can be obtained from multi-echo T2 imaging.
-Our study protocol involved the collection of multi-echo SWI imaging data, and as such, we could only use R2^*^ data to perform the decomposition.
+Our study protocol involved the collection of multi-echo SWI imaging data, and as such, we could only use R2^\*^ data to perform the decomposition.
 Furthermore, this may account for the reduced precision of SvO~2~ values obtained through susceptibility separation.

### 2024-10-11 notebooks/gavin_thesis_markdown.qmd
Submission to Pediatric Research

@@
   y = Chi,
+  title = "Cerebral Central Veins",
   ylab = "Chi (ppm)",
@@
   y = Measured.Susceptibility, 
+  title = "Superior Sagittal Sinus",
   ylab = "Chi (ppm)",
@@
   y = Oxygen.Saturation,
+  title = "Cerebral Central Veins",
   ylab = "SvO2 (%)",
@@
   y = Oxygen.Saturation,
+  title = "Superior Sagittal Sinus",
   ylab = "SvO2 (%)",
@@
 #| fig-height: 5
-#| fig-cap: "Vein-specific susceptibility and oxygen saturation values by method.
-(A, B) contains violin plots comparing subject χ (ppm) acquired from QSM (A) and its paramagnetic component (B); (C, D) contains violin plots comparing subject SvO2 (%) acquired from QSM (C) and its paramagnetic component (D).
+#| fig-cap: "**Vein-specific susceptibility and oxygen saturation values by method.** (A, B) contains violin plots comparing subject chi (ppm) acquired from the cerebral central veins; (C, D) contains violin plots comparing subject SvO2 (%) acquired from the superior sagittal sinus.
 Raw data points from paramagnetic maps are shown as filled green circles and raw data points from QSM are shown as filled orange circles.
@@
   ylab = "Chi (ppm)",
+  title = "QSM",
   centrality.plotting = FALSE,
@@
   y = CSvO2,
+  title = "QSM",
   results.subtitle = FALSE,
@@
   y = Chi, 
+  title = "Paramagnetic susceptibility",
   results.subtitle = FALSE,
@@
   y = CSvO2, 
+  title = "Paramagnetic susceptibility",
   results.subtitle = FALSE,
@@
 #| fig-height: 5
-#| fig-cap: "Inter-venous comparisons of susceptibility and oxygen saturation.
-Violin plots comparing (A, C) χ (ppm) and (B, D) SvO2 (%) between the CCV and the SSS.
+#| fig-cap: "**Inter-venous comparisons of susceptibility and oxygen saturation**.
+Violin plots comparing (A, C) chi (ppm) and (B, D) SvO2 (%) between the CCV and the SSS.
 Panels (A) and (B) used data acquired from QSM, and its raw data points are shown as filled orange circles.

### 2024-10-11 _tex/index.tex
Built site for gh-pages

--- 2024-06-25 _tex/index.tex
@@
 \documentclass[
-sn-nature
+true
 ]{sn-jnl}
@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
 \makeatother
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
-  \fi
-  % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 \ifdefined\figurename
-  \renewcommand*\figurename{Figure}
+  \renewcommand*\figurename{\textbf{Figure}}
 \else
-  \newcommand\figurename{Figure}
+  \newcommand\figurename{\textbf{Figure}}
 \fi
 \ifdefined\tablename
-  \renewcommand*\tablename{Table}
+  \renewcommand*\tablename{\textbf{Table}}
 \else
-  \newcommand\tablename{Table}
+  \newcommand\tablename{\textbf{Table}}
 \fi
@@
 \newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\captionsetup{labelsep=period}
 \makeatother
@@
 
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
 \usepackage{bookmark}
@@
   pdftitle={The Application of Magnetic Susceptibility Separation for Measuring Cerebral Oxygenation in Preterm Neonates},
-  pdfauthor={Thomas Gavin Carmichael; Alexander Rauscher; Ruth E Grunau; Alexander Mark Weber},
+  pdfauthor={, , , and },
   pdfkeywords={Quantitative Susceptbility
@@
 % affil setup
-\affil[1]{\orgdiv{Integrated Sciences}, \orgname{The University of
-British Columbia}, \orgaddress{\street{2329 West
-Mall}, \city{Vancouver}, \postcode{V6T 1Z4}, \country{Canada}}}
-\affil[3]{\orgdiv{Pediatrics}, \orgname{The University of British
-Columbia}, \orgaddress{\street{2329 West
-Mall}, \city{Vancouver}, \postcode{V6T 1Z4}, \country{Canada}}}
-\affil[2]{\orgdiv{BC Children's Hospital Research
-Institute}, \orgname{The University of British
-Columbia}, \orgaddress{\street{938 West 28th
-Avenue}, \city{Vancouver}, \postcode{V5Z 4H4}, \country{Canada}}}
+\affil[1]{, \orgname{Integrated Sciences, The University of British
+Columbia, Vancouver, BC, Canada}}
+\affil[2]{, \orgname{BC Children's Hospital Research Institute, The
+University of British Columbia, Vancouver, BC, Canada}}
+\affil[3]{, \orgname{Pediatrics, The University of British Columbia,
+Vancouver, BC, Canada}}
 
@@
 
-\abstract{\textbf{Background}: Quantitative susceptibility mapping (QSM)
-is a magnetic resonance imaging (MRI) modality proposed to be a viable
-method of measuring cerebral oxygenation in neonates given its
-sensitivity to deoxyhemoglobin, a paramagnetic molecule. During QSM,
-however, paramagnetic sources can be obscured by opposing diamagnetic
-sources such as water and myelin. We sought to evaluate whether QSM
-images alone, or an algorithm that attempts to isolate their
-paramagnetic components, are more accurate in measuring oxygenation of
-the major cerebral veins in a cohort of neonates born preterm.
-Additionally, we aimed to determine whether a difference in oxygenation
-existed between the major cerebral veins.
+\abstract{\textbf{Background}: Quantitative susceptibility mapping
+(QSM), an magnetic resonance imaging (MRI) modality sensitive to
+deoxyhemoglobin, is a promising method for measuring cerebral
+oxygenation in human neonates. Paramagnetic sources, like
+deoxyhemoglobin, however, can be obscured by diamagnetic sources such as
+water and myelin. This study evaluated whether QSM images, or isolated
+paramagnetic components, are more accurate for measuring oxygenation of
+cerebral veins of preterm neonates, and explored oxygenation differences
+between the major cerebral veins.
 
-\textbf{Methods}: 19 neonates born preterm were scanned on a 3T research
-MRI at term equivalent age. The protocol included a multi-echo
-susceptibility-weighted imaging sequence. The acquired imaging data were
-processed as QSM images to obtain the susceptibility values of the
-superior sagittal sinus (SSS) and central cerebral veins (CCV). These
-values were used to calculate the oxygen saturation
-(SvO\textsubscript{2}) of the SSS and CCV. QSM images were subsequently
-processed to isolate their paramagnetic components. SvO\textsubscript{2}
-values of the SSS and CCV were calculated again from the paramagnetic
-components.
+\textbf{Methods}: 19 preterm neonates were scanned on at term equivalent
+age on a 3T MRI using a multi-echo susceptibility-weighted imaging
+sequence. Susceptibility values were calculated from QSM images to
+determine oxygen saturation (SvO\textsubscript{2}) in the superior
+sagittal sinus (SSS) and central cerebral veins (CCV). The paramagnetic
+components of QSM images were isolated, and SvO\textsubscript{2} values
+were recalculated.
 
-\textbf{Results}: The mean SvO\textsubscript{2} values of the SSS and
-CCV calculated from QSM images were found to be 72.4\% (SD, 3.4\%) and
-68.7\% (SD, 3.5\%), respectively. The mean SvO\textsubscript{2} values
-calculated from paramagnetic components were found to be 58.1\% (SD,
+\textbf{Results}: The mean SvO\textsubscript{2} values from QSM were
+72.4\% (SD, 3.4\%) for the SSS and and 68.7\% (SD, 3.5\%) for the CCV.
+SvO\textsubscript{2} values for paramagnetic components were 58.1\% (SD,
 7.3\%) for the SSS and 57.7\% (SD, 7.0\%) for the CCV.
 
-\textbf{Conclusion}: SSS SvO\textsubscript{2} values derived from
-paramagnetic components agreed well with the existing literature and
-were closer than the values derived from QSM, however, they displayed
-greater variability. Although the CCV SvO\textsubscript{2} data from QSM
-aligns more closely with existing literature, it is important to note
-that the current literature on this topic remains relatively limited in
-the CCV. Thus, decomposing QSM images into paramagnetic components shows
-great promise as a method for more accurately measuring cerebral
-oxygenation in neonates but may require more research to improve
-precision. Notably, no significant difference in oxygenation was
-observed between the CCV and the SSS, contrasting with previous
-studies.}
+\textbf{Conclusion}: While paramagnetic component decomposition yielded
+more accurate SSS values, it increased variability. No significant
+oxygenation differences were found between the SSS and CCV, contrasting
+with prior studies.}
 
@@
 
+\textsuperscript{1} Integrated Sciences, The University of British
+Columbia, Vancouver, BC, Canada\\
+\textsuperscript{2} BC Children's Hospital Research Institute, The
+University of British Columbia, Vancouver, BC, Canada\\
+\textsuperscript{3} Pediatrics, The University of British Columbia,
+Vancouver, BC, Canada
+
+\textsuperscript{*} Correspondence:
+\href{mailto:aweber@bcchr.ca}{Alexander Mark Weber
+\textless{}aweber@bcchr.ca\textgreater{}}
+
+\textbf{Category of Study:} basic science
+
+\subsection{Impact Statement}\label{impact-statement}
+
+\begin{itemize}
+\tightlist
+\item
+  This study evaluated the accuracy of using QSM and its paramagnetic
+  components to measure cerebral oxygenation in neonates.
+\item
+  By comparing susceptibility-derived oxygen saturation
+  (SvO\textsubscript{2}) in the superior sagittal sinus (SSS) and
+  central cerebral veins (CCV), it adds to the field of neonatal
+  cerebral oxygenation measurement.
+\item
+  Decomposing QSM into paramagnetic components shows potential for
+  improving SvO\textsubscript{2} accuracy, particularly in the SSS,
+  though variability remains a challenge.
+\item
+  The results suggest no significant oxygenation difference between the
+  SSS and CCV, contrasting with previous findings, indicating a need for
+  further research on neonatal venous oxygenation.
+\end{itemize}
+
+\subsection{Acknowledgments}\label{acknowledgments}
+
+We wish to acknowledge the work of and thank Victoria Tapics (Research
+Nurse); Vicki Goh (Research Nurse); Chacko Anil (Neonatologist); and
+Michael A. Sargent (Radiologist).
+
+\subsection{Funding}\label{funding}
+
+Authors AMW and REG were co-primary applicants for a BC Children's
+Hospital Research Institute - Brain, Behaviour and Development Catalyst
+Grant (\$20,000). AMW was supported by an establishment award from
+BCCHRI. Scanning was partly funded through a special award to AMW from
+BCCHRI .
+
+\subsection{Author contributions}\label{author-contributions}
+
+TGC wrote the original draft, performed the formal analysis, and
+contributed to methodology, validation, and visualization. AR helped
+with writing, reviewing \& editing. REG helped with writing, reviewing
+\& editing, and with initial funding acquisition. AMW was involved in
+project administration, supervision, validation, visualization,
+resources, methodology, formal analysis, funding acquisition, writing,
+reviewing \& editing, conceptualization, data curation, and
+investigation.
+
+\subsection{Competing interests}\label{competing-interests}
+
+The authors have no competing interests to declare.
+
+\subsection{Consent statement}\label{consent-statement}
+
+The study was approved by the Clinical Research Ethics Board at the
+University of British Columbia and Children's \& Women's Hospital
+(H21-00655) and written informed consent was obtained from the
+parent/guardian for each infant.
+
+\newpage{}
+
 \section{Introduction}\label{sec-intro}
 
-Preterm birth
+With advances in neonatal medical care, more infants born preterm are
+surviving into childhood \citep{mckenzieScaffoldingParentingHealth2022}.
+These children are at high risk of acquiring adverse neurodevelopmental
+outcomes when compared to their term-born peers
+\citep{twilhaarCognitiveOutcomesChildren2018}. Irregularities in early
+cerebral oxygen levels have been identified as a potential source of
+such delays, where too little oxygen provided during NICU care can
+result in white matter injury, while too much oxygen can result in
+reduced cortical connectivity \citep{rantakariEarlyOxygenLevels2021}. As
+such, being able to precisely, accurately, and non-invasively measure
+cerebral oxygenation is necessary for understanding and improving
+neurodevelopmental outcomes in preterm neonates.
 
-Abnormal brain development is a significant concern for parents with
-children born preterm, as 43\% of infants that survive will have
-neurodevelopmental delays later in life (Dix, van Bel, and Lemmers
-2017). Irregularities in early cerebral oxygen levels have been
-identified as a potential source of such delays, where too little oxygen
-provided during NICU care can result in white matter injury, while too
-much oxygen can result in reduced cortical connectivity (Rantakari et
-al. 2021). As such, being able to precisely, accurately, and
-non-invasively measure cerebral oxygenation is necessary for
-understanding and improving neurodevelopmental outcomes in preterm
-neonates.
+Unfortunately, there exist many challenges in measuring cerebral oxygen
+metabolism in neonates. Cerebral metabolic rate of oxygen
+(CMRO\textsubscript{2}) using oxygen-15 positron emission tomography
+(PET) \citep{mintunBrainOxygenUtilization1984}, has been measured in
+infants \citep{altmanCerebralBloodFlow1988}, and is considered the gold
+standard. However, this method is invasive, requiring ionizing
+radiation, which limits its suitability for neonates. A less invasive
+option for evaluating brain hemodynamics is near-infrared spectroscopy
+(NIRS), which uses the attenuation of near-infrared light
+(\textasciitilde650--950 nm) as it passes through biological tissue
+\citep{skovEstimationCerebralVenous1993}. Deoxygenated and oxygenated
+hemoglobin absorb this light differently, allowing NIRS to estimate
+changes in deoxyhemoglobin and oxyhemoglobin
+\citep{wrayCharacterizationInfraredAbsorption1988} and thus provide an
+estimate of cerebral venous oxygen saturation (SvO\textsubscript{2}).
+While NIRS offers the advantage of being non-invasive and continuous
+bedside monitoring, it is limited to regional assessments where the
+probe is placed and is sensitive only to superficial brain tissue due to
+the shallow penetration depth of near-infrared light
+\citep{boasDiffuseOpticalImaging2004}.
 
-In the present study, we set out to determine whether a QSM image alone,
-or the paramagnetic component of the QSM image, is more accurate in
-measuring the oxygen present in the major cerebral veins of a cohort of
-preterm neonates. As previous QSM studies have not included the SSS, we
-also had a secondary aim of preserving this vessel in our QSM images and
-using this data to determine whether a difference in oxygenation existed
-between the SSS and the central cerebral veins (CCV).
+For the preceding reasons, non-invasive MRI-based techniques are
+actively being explored to assess regional and whole-brain blood
+oxygenation. MRI based methods have been developed for adults
+\citep{jainRapidMagneticResonance2011, luQuantitativeEvaluationOxygenation2008, xuNoninvasiveQuantificationWholebrain2009}
+and are only beginning to be explored in neonates
+\citep{devisNoninvasiveMRIMeasurements2014, liuQuantitativeAssessmentGlobal2014, qiHemodynamicMetabolicAssessment2018, jainCerebralOxygenMetabolism2014, jiangVesselspecificQuantificationNeonatal2019}.
+These methods almost all rely on T2 relaxation to estimate
+CSvO\textsubscript{2}
+\citep{devisNoninvasiveMRIMeasurements2014, liuQuantitativeAssessmentGlobal2014, qiHemodynamicMetabolicAssessment2018, jiangVesselspecificQuantificationNeonatal2019}
+with the exception of \citet{jainCerebralOxygenMetabolism2014} which
+used susceptometry \citep{jainMRIEstimationGlobal2010}. One limitation
+of these T2 relaxation methods, however, is the fact that
+SvO\textsubscript{2} is often measured using a single imaging slice,
+averaging values across several voxels, and only in the superior
+sagittal sinus (SSS). In the case of
+\citet{jainCerebralOxygenMetabolism2014}, they obtained regional and
+whole-brain data, but with thick slices (5mm), and still only estimated
+CSvO\textsubscript{2} in the SSS. An alternative MRI method using
+quantitative susceptibility mapping (QSM) has been proposed, which can
+measure SvO\textsubscript{2} regionally and across the whole-brain at
+high resolution (\textless{} 1mm\textsuperscript{3} per voxel)
+\citep{weberQuantitativeSusceptibilityMapping2021}. However, this method
+left room for improvement, as it removed the SSS (averaging
+CSvO\textsubscript{2} across the internal veins), and required an
+arbitrary threshold value of 0.15 ppm in order to acquire realistic
+results \citep{weberQuantitativeSusceptibilityMapping2021}.
+
+In the present study, we set out to determine whether decomposing the
+QSM image into its paramagnetic and diamagnetic components would allow
+for a more accurate assessment of SvO\textsubscript{2} in the central
+cerebral veins (CCV) of a cohort of preterm neonates. We also had a
+secondary aim of preserving the SSS vessel in our QSM images and using
+this data to determine whether a difference in oxygenation existed
+between the SSS and the CCV.
 
@@
 
-Participant data comes from a previous study (Zhu et al. 2024).
-Participants consisted of preterm neonates born between 25- and 31-weeks
-gestational age (GA) who were admitted to the level III NICU at BC
-Women's Hospital. Recruitment took place over a span of one year, from
-February 2021 to January 2022, facilitated by a dedicated research
-nurse. Parents of eligible infants were approached by the research nurse
-prior to discharge from the NICU to explain the study objectives and
-seek their consent for participation. Infants meeting the criteria for
-inclusion were scanned for the study if they had already been discharged
-from the NICU, were in stable condition, and had reached a term
-equivalent age of 37 to 44 weeks GA. However, certain exclusion criteria
-were applied to ensure the homogeneity and integrity of the study
-sample. Infants were excluded if there was clinical evidence of a
-congenital malformation or syndrome, a TORCH infection, or ultrasound
-evidence of large parenchymal hemorrhagic infarction (\textgreater2 cm,
-Grade 4 intraventricular hemorrhage).
+Participant data comes from a previous study *** . Participants
+consisted of preterm neonates born between 25- and 31-weeks gestational
+age (GA) who were admitted to the level III NICU at *** . Recruitment
+took place over a span of one year, from February 2021 to January 2022,
+facilitated by a dedicated research nurse. Parents of eligible infants
+were approached by the research nurse prior to discharge from the NICU
+to explain the study objectives and seek their consent for
+participation. Infants meeting the criteria for inclusion were scanned
+for the study if they had already been discharged from the NICU, were in
+stable condition, and had reached a term equivalent age of 37 to 44
+weeks GA. However, certain exclusion criteria were applied to ensure the
+homogeneity and integrity of the study sample: infants were excluded if
+there was clinical evidence of a congenital malformation or syndrome, a
+TORCH infection, or ultrasound evidence of large parenchymal hemorrhagic
+infarction (\textgreater2 cm, Grade 4 intraventricular hemorrhage).
 
@@
 scanner (scanner software version DV26.0\_R03) equipped with a SREE
-Medical Systems single-channel neonatal head coil (Table~\ref{tbl-mri}).
-The scans were conducted at the BC Children's MRI Research Facility.
-Prior to the scanning procedure, subjects were carefully prepared by a
-research nurse. Swaddling and feeding were used to ensure the comfort
-and cooperation of the subjects during the scan. Importantly, no
-sedatives or invasive markers were utilized throughout the procedure.
-Subjects were placed within a specially designed SREE Medical Systems
-MRI compatible incubator, which facilitated both safety and motion
-minimization. Molded foam was strategically positioned around the head
-and body within the incubator to further restrict subject movement. To
-protect against potential hearing damage, ear plugs were employed during
-the scanning process. Additionally, a pulse oximeter was affixed to the
-subject's foot to monitor arterial oxygen saturation and heart rate
-throughout the scan.
+Medical Systems (Cleveland, OH, USA) single-channel neonatal head coil
+(Table~\ref{tbl-mri}). The scans were conducted at the *** 's MRI
+Research Facility. Prior to the scanning procedure, subjects were
+carefully prepared by a research nurse. Swaddling and feeding were used
+to ensure the comfort and cooperation of the subjects during the scan.
+Importantly, no sedatives or invasive markers were utilized throughout
+the procedure. Subjects were placed within a specially designed SREE
+Medical Systems MRI compatible incubator, which facilitated both safety
+and motion minimization. Molded foam was strategically positioned around
+the head and body within the incubator to further restrict subject
+movement. To protect against potential hearing damage, ear plugs were
+employed during the scanning process. Additionally, a pulse oximeter was
+affixed to the subject's foot to monitor arterial oxygen saturation and
+heart rate throughout the scan.
 
@@
 
-\caption{\label{tbl-mri}Technical parameters for MR imaging pulse
-sequences}
+\caption{\label{tbl-mri}\textbf{Technical parameters for MR imaging
+pulse sequences}}
 
@@
 TR (ms) & 7.74 & 2,300 & 4,680 & 30.9\\
-TE (ms) & 2.97 & 66.29 & 10.55 & 5.24\\
+TE (ms) & 2.97 & 66.29 & 10.55 & 5 echoes; first echo: 5; echo spacing: 5.24\\
 Flip angle & 12° & 90° & 111° & 20°\\
@@
 T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin
-labeling (ASL) scan, a multi-echo susceptibility-weighted imaging scan,
-and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging
-(EPI) sequence. The DWI sequence was not used for the present study.
+labeling (ASL) scan \citep{alsopRecommendedImplementationArterial2015},
+a multi-echo susceptibility-weighted imaging scan
+\citep{denkSusceptibilityWeightedImaging2010}, and a diffusion-weighted
+imaging (DWI) spin-echo echo planar imaging (EPI) sequence. The DWI
+sequence was not used for the present study.
 
@@
 to NIfTI (Neuroimaging Informatics Technology Initiative) format using
-Chris Rorden's \texttt{dcmniix} tool. SWI magnitude data files were then
+Chris Rorden's \texttt{dcmniix} tool
+\citep{liFirstStepNeuroimaging2016}. SWI magnitude data files were then
 used to create subject-specific brain masks that would not erode the SSS
-during QSM processing, an issue faced by our group in the past (X. Li et
-al. 2016). A step-by-step summary of the pipeline used is shown in
-Figure~\ref{fig-sample}.
+during QSM processing, an issue faced by our group in the past *** . A
+step-by-step summary of the pipeline used is shown in
+Figure~\ref{fig-graph}.
 
@@
 
-\includegraphics{index_files/figure-latex/notebooks-Figures-fig-graph-output-1.png}
+\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-graph-output-1.png}}
 
@@
 
-\caption{\label{fig-graph}Pipeline for generating subject-specific brain
-masks that include the superior sagittal sinus (SSS). Initial steps
-involved (1) creating brain a mask from the magnitude of the fifth echo
-of the susceptibility weighted scan. Subsequently, the brain mask is
-dilated and then (2) utilized in conjunction with a quantitative
-susceptibility mapping (QSM) script to generate a preliminary QSM image.
-Further refinement involved (3) segmenting the SSS from the QSM image
-manually to create a tissue mask of the SSS region. Finally, (4) the
-vascular mask of the SSS is integrated with the initial brain mask,
-forming the comprehensive brain mask essential for obtaining
-susceptibility data that includes the SSS.}
+\caption{\label{fig-graph}\textbf{Pipeline for generating
+subject-specific brain masks that include the superior sagittal sinus
+(SSS).} Initial steps involved (1) creating brain a mask from the
+magnitude of the fifth echo of the susceptibility weighted scan.
+Subsequently, the brain mask is dilated and then (2) utilized in
+conjunction with a quantitative susceptibility mapping (QSM) script to
+generate a preliminary QSM image. Further refinement involved (3)
+segmenting the SSS from the QSM image manually to create a tissue mask
+of the SSS region. Finally, (4) the vascular mask of the SSS is
+integrated with the initial brain mask, forming the comprehensive brain
+mask essential for obtaining susceptibility data that includes the SSS.}
 
@@
 
-First, the fifth echo SWI magnitude file was processed using FSL's (v.
-6.0.7.3) (Woolrich et al. 2009) \texttt{fslroi}, \texttt{fslmaths}, and
-\texttt{bet} (Smith 2002) to create a preliminary brain mask, similar to
-our previous efforts, which does not contain the SSS. \texttt{Fslroi}
-was used to isolate the fifth echo of the magnitude data, which was then
-squared using Fslmaths and the option \texttt{-sqr}. Squaring the
-magnitude image was found to dramatically improve subsequent brain
-extraction. The resulting image was then used to create the preliminary
-brain mask using bet with the options \texttt{-m} and \texttt{-R}. The
-former flag generated a binary brain mask, while the latter performed a
-more robust brain centre estimation. The brain mask was then dilated by
-7 voxels using \texttt{Fslmaths} and the options \texttt{-kernel\ boxv}
-and \texttt{-dilM} in order for the dilated mask to contain the SSS
-(along with unwanted tissue as well). This mask was then used, along
-with the phase images, in a MATLAB script for QSM calculation from
-Christian Kames (Kames, Wiggermann, and Rauscher 2018) to produce a
-preliminary QSM image that contained the SSS, albeit with fairly low
-signal-to-noise ratio and other unwanted tissue. Given the high contrast
-in voxel intensity between the SSS and surrounding tissue, the select by
-intensity tool in \texttt{FSLeyes} (McCarthy 2023) was then used to
-segment the SSS from the QSM image and create a 3D mask of the selected
-region. Using \texttt{fslmaths} and the options \texttt{-add} and
-\texttt{-bin}, the SSS mask was then combined with the original brain
-mask of the fifth echo. This resulted in a brain mask that contained
-only brain and SSS signal. Finally, this mask was used in a final QSM
-post-processing step to create a QSM image that includes the SSS while
-maintaining a high signal-to-noise ratio, making it suitable to obtain
-accurate susceptibility values.
+First, the fifth echo (TE = 25.96 ms) SWI magnitude file was processed
+using FSL's (v. 6.0.7.3)
+\citep{woolrichBayesianAnalysisNeuroimaging2009} \texttt{fslroi},
+\texttt{fslmaths}, and \texttt{bet} \citep{smithFastRobustAutomated2002}
+to create a preliminary brain mask, similar to our previous efforts,
+which does not contain the SSS. \texttt{Fslroi} was used to isolate the
+fifth echo of the magnitude data, which was then squared using
+\texttt{fslmaths} and the option \texttt{-sqr}. Squaring the magnitude
+image was found to dramatically improve subsequent brain extraction. The
+resulting image was then used to create the preliminary brain mask using
+bet with the options \texttt{-m} and \texttt{-R}. The former flag
+generated a binary brain mask, while the latter performed a more robust
+brain centre estimation. The brain mask was then dilated by 7 voxels
+using \texttt{fslmaths} and the options \texttt{-kernel\ boxv} and
+\texttt{-dilM} in order for the dilated mask to contain the SSS (along
+with unwanted tissue as well). This mask was then used, along with the
+phase images, in a MATLAB script for QSM calculation from Christian
+Kames \citep{kamesRapidTwostepDipole2018} to produce a preliminary QSM
+image that contained the SSS, albeit with fairly low signal-to-noise
+ratio and other unwanted tissue. Given the high contrast in voxel
+intensity between the SSS and surrounding tissue, the select by
+intensity tool in \texttt{FSLeyes} \citep{mccarthyFSLeyes2023} was then
+used to segment the SSS from the QSM image and create a 3D mask of the
+selected region. Using \texttt{fslmaths} and the options \texttt{-add}
+and \texttt{-bin}, the SSS mask was then combined with the original
+brain mask of the fifth echo. This resulted in a brain mask that
+contained only brain and SSS signal. Finally, this mask was used in a
+final QSM post-processing step to create a QSM image that includes the
+SSS while maintaining a high signal-to-noise ratio, making it suitable
+to obtain accurate susceptibility values.
 
-STI Suite (v. 3.0) (W. Li et al. 2014), was used to process the final
-QSM images as it produced the cleanest images without eroding the SSS.
-The finalized brain mask and the last three echoes of the magnitude and
-phase images were used in STI Suite along with the following parameters:
-0.9766 x 0.9766 x 1 mm\textsuperscript{3} voxel size, 5 ms TE1, 5.3 ms
-\(\Delta\)TE, and 77.4 ms sum TE, B0 strength = 3, and B0 direction =
-{[}0, 0, 1{]}. The 3D GRE data option was selected for the phase
-processing stage, and STAR-QSM was selected for the QSM stage. Finally,
-the `select by intensity' tool in \texttt{FSLeyes} was then used to
-semi-automatically make vascular masks of the SSS and CCV from each
-subject's QSM image. The vascular masks were used to calculate the mean
-susceptibility of each subject's SSS and CCV from their QSM image with
-\texttt{fslstats}.
+STI Suite (v. 3.0) \citep{liIntegratedLaplacianbasedPhase2014}, was used
+to process the final QSM images as it produced the cleanest images
+without eroding the SSS. The finalized brain mask and the last three
+echoes of the magnitude and phase images were used in STI Suite along
+with the following parameters: 0.9766 x 0.9766 x 1 mm\textsuperscript{3}
+voxel size, 5 ms TE1, 5.3 ms \(\Delta\)TE, and 77.4 ms sum TE, B0
+strength = 3, and B0 direction = {[}0, 0, 1{]}. The 3D GRE data option
+was selected for the phase processing stage, and STAR-QSM was selected
+for the QSM stage. Finally, the `select by intensity' tool in
+\texttt{FSLeyes} was then used to semi-automatically make vascular masks
+of the SSS and CCV from each subject's QSM image. The vascular masks
+were used to calculate the mean susceptibility of each subject's SSS and
+CCV from their QSM image with \texttt{fslstats}.
 
 To isolate the paramagnetic component of subjects' QSM data, the
-\(\chi\)-separation toolbox (Shin et al. 2021) from the Laboratory for
-Imaging Science and Technology was used. Each subject's magnitude and
-phase SWI data were used along with the following parameters: 0.9766 x
-0.9766 x 1 mm\textsuperscript{3} voxel size; TE (s) = {[}0.005, 0.0102,
-0.0155, 0.0207, 0.026{]}; delta TE (s) = 0.0052; B0 strength = 3; B0
-direction = {[}0, 0, 1{]}. The mean susceptibility of each subject's SSS
-and CCV in their paramagnetic maps was calculated with the same vascular
-masks used for the QSM images. Sample images showing the magnitude,
-final QSM, and final paramagnetic component images are shown in
-Figure~\ref{fig-sample}.
+\(\chi\)-separation toolbox
+\citep{shinHseparationMagneticSusceptibility2021} from the Laboratory
+for Imaging Science and Technology was used. Each subject's magnitude
+and phase SWI data were used along with the following parameters: 0.9766
+x 0.9766 x 1 mm\textsuperscript{3} voxel size; TE (s) = {[}0.005,
+0.0102, 0.0155, 0.0207, 0.026{]}; delta TE (s) = 0.0052; B0 strength =
+3; B0 direction = {[}0, 0, 1{]}. The mean susceptibility of each
+subject's SSS and CCV in their paramagnetic maps was calculated with the
+same vascular masks used for the QSM images. Sample images showing the
+magnitude, final QSM, and final paramagnetic component images are shown
+in Figure~\ref{fig-sample}.
 
@@
 
-\includegraphics{index_files/figure-latex/notebooks-Figures-fig-sample-output-1.png}
+\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-sample-output-1.png}}
 
@@
 
-\caption{\label{fig-sample}An example of subject imaging data. A sample
-sagittal, coronal, and axial slice is displayed for each image. (a) The
-1st echo of the magnitude susceptibility weighted imaging sequence; (b)
-the final quantitative susceptibility mapping image; and (c) the
-paramagnetic component isolated from the quantitative susceptibility
-map. The bars in (b) and (c) indicates the range of susceptibility
-\(\chi\) values.}
+\caption{\label{fig-sample}\textbf{An example of subject imaging data.}
+A sample sagittal, coronal, and axial slice is displayed for each image.
+(a) The 1st echo of the magnitude susceptibility weighted imaging
+sequence; (b) the final quantitative susceptibility mapping image; and
+(c) the paramagnetic component isolated from the quantitative
+susceptibility map. The bars in (b) and (c) indicates the range of
+susceptibility \(\chi\) values.}
 
@@
 Once the mean susceptibility values of the SSS and CCV were obtained
-from the subjects' QSM images and paramagnetic maps, venous oxygen
+from the subjects' QSM images or paramagnetic maps, venous oxygen
 saturation (SvO\textsubscript{2}) was calculated with the following
-equation (Berg et al. 2021):
+equation \citep{bergInvestigatingEffectFlow2021}:
 
@@
 blood, and Hct is the subject's hematocrit. \(\Delta \chi _{oxy}\) was
--0.21 * 4\(\pi\) ppm as per Portnoy et al. (2018) and (Sedlacik,
-Rauscher, and Reichenbach 2007), while \(\Delta \chi _{do}\) was -0.03 *
-4\(\pi\) ppm as per (Weisskoff and Kiihne 1992). Subjects' Hct for the
+-0.21 * 4\(\pi\) ppm as per \citet{portnoyHumanUmbilicalCord2018} and
+\citep{sedlacikObtainingBloodOxygenation2007}, while
+\(\Delta \chi _{do}\) was -0.03 * 4\(\pi\) ppm as per
+\citep{weisskoffMRISusceptometryImagebased1992}. Subjects' Hct for the
 day of the scan was calculated using a four-parameter Weibull function
@@
 
-Statistical analysis of the acquired data was performed using R and
-RStudio (v. 2023.09.1 Build 494) (R Core Team 2022; RStudio Team, n.d.).
+Statistical analysis was performed using R and RStudio (v. 2023.09.1
+Build 494)
+\citep{rcoreteamLanguageEnvironmentStatistical2022, rstudioteamRStudioIntegratedDevelopment}.
 Mean and standard deviation values are reported for most statistics,
@@
 
-\begin{longtable}[t]{lc}
+\begin{longtable}[t]{lr}
 
-\caption{\label{tbl-dem}Demographic and clinical characteristic of the
-study sample.}
+\caption{\label{tbl-dem}\textbf{Demographic and clinical characteristic
+of the study sample.}}
 
@@
 The mean SvO\textsubscript{2} values for the SSS and the CCV were found
-to be 0.72 \(\pm\) 0.03 and 0.69 \(\pm\) 0.03 ppm, respectively, when
+to be 0.72 \(\pm\) 0.03\% and 0.69 \(\pm\) 0.03\%, respectively, when
 determined from the QSM data. When determined from the paramagnetic map,
 the mean SvO\textsubscript{2} values for the SSS and the CCV were found
-to be 0.58 \(\pm\) 0.07 \%, respectively. A summary of the measured
-physiological parameters, including the chi values used to calculate
-SvO\textsubscript{2}, can found in Table~\ref{tbl-chistats}.
+to be 0.58 \(\pm\) 0.07\% and 0.58 \(\pm\) 0.07\%, respectively. A
+summary of the measured physiological parameters, including the chi
+values used to calculate SvO\textsubscript{2}, can found in
+Table~\ref{tbl-chistats}.
 
@@
 
-\caption{\label{tbl-chistats}Summary of acquired physiological
-parameters. Mean \(\pm\) SD is shown for chi and SvO\textsubscript{2}
-values. The P-value and 95\% confidence interval (CI) were obtained
-through the comparison of values between QSM and paramagnetic maps;
-(n=19).}
+\caption{\label{tbl-chistats}\textbf{Summary of acquired physiological parameters.}
+Mean \(\pm\) SD is shown for chi and SvO\textsubscript{2} values. The
+P-value and 95\% confidence interval (CI) were obtained through the
+comparison of values between QSM and paramagnetic maps; (n=19).}
 
@@
 
-\includegraphics{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-methodplot-output-1.png}
+\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-methodplot-output-1.png}}
 
@@
 
-\caption{\label{fig-methodplot}Vein-specific susceptibility and oxygen
-saturation values by method. (A, B) contains violin plots comparing
-subject χ (ppm) acquired from QSM (A) and its paramagnetic component
-(B); (C, D) contains violin plots comparing subject SvO2 (\%) acquired
-from QSM (C) and its paramagnetic component (D). Raw data points from
-paramagnetic maps are shown as filled green circles and raw data points
-from QSM are shown as filled orange circles. Each line connects the raw
-data points of a single subject. (***) indicates P\textless0.05.}
+\caption{\label{fig-methodplot}\textbf{Vein-specific susceptibility and
+oxygen saturation values by method.} (A, B) contains violin plots
+comparing subject chi (ppm) acquired from the cerebral central veins;
+(C, D) contains violin plots comparing subject SvO2 (\%) acquired from
+the superior sagittal sinus. Raw data points from paramagnetic maps are
+shown as filled green circles and raw data points from QSM are shown as
+filled orange circles. Each line connects the raw data points of a
+single subject. (***) indicates P\textless0.05.}
 
@@
 
-\includegraphics{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-regionplot-output-1.png}
+\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-regionplot-output-1.png}}
 
@@
 
-\caption{\label{fig-regionplot}Inter-venous comparisons of
-susceptibility and oxygen saturation. Violin plots comparing (A, C) χ
+\caption{\label{fig-regionplot}\textbf{Inter-venous comparisons of
+susceptibility and oxygen saturation}. Violin plots comparing (A, C) chi
 (ppm) and (B, D) SvO2 (\%) between the CCV and the SSS. Panels (A) and
@@
 application of magnetic susceptibility separation to neonatal QSM data
-could provide more accurate cerebral venous vessel oxygenation
-measurements. To the best of our knowledge, we are the first to test
-this in a neonatal cohort, as susceptibility separation has been
-typically evaluated as a method of imaging myelin and brain iron in
-adult subjects (Shin et al. 2021; Ahmed et al. 2023). Our results showed
-that the SvO\textsubscript{2} values of the SSS and CCV obtained from
-susceptibility separation are significantly lower than the respective
-SvO\textsubscript{2} values obtained from QSM alone. When our results
-were compared to the literature, we found that our SSS
-SvO\textsubscript{2} data from susceptibility separation agreed well
-with the findings of other studies measuring SvO\textsubscript{2} of the
-SSS in similar subject populations. Conversely, the paramagnetic CCV
-SvO\textsubscript{2} data saw less agreement with the existing
+could provide more accurate SvO\textsubscript{2} measurements without
+the need for an arbitrary threshold value. To the best of our knowledge,
+we are the first to test this in a neonatal cohort, as susceptibility
+separation has been typically evaluated as a method of imaging myelin
+and brain iron in adult subjects
+\citep{shinHseparationMagneticSusceptibility2021, ahmedDiamagneticComponentMap2023a}.
+Our results showed that the SvO\textsubscript{2} values of the SSS and
+CCV obtained from susceptibility separation are significantly lower than
+the respective SvO\textsubscript{2} values obtained from QSM alone. When
+our results were compared to the literature (see below), we found that
+our SSS SvO\textsubscript{2} data from susceptibility separation agreed
+well with the findings of other studies measuring SvO\textsubscript{2}
+of the SSS in similar subject populations. Conversely, the paramagnetic
+CCV SvO\textsubscript{2} data saw less agreement with the existing
 literature than the corresponding data from QSM. However, there is
@@
 
+\subsection{Comparison with literature
+values}\label{comparison-with-literature-values}
+
+To evaluate the validity of our results, we compared the mean
+SvO\textsubscript{2} values we obtained through QSM and susceptibility
+separation to the mean SvO\textsubscript{2} values found by MRI studies
+investigating the oxygenation of the SSS or the CCV. Notably, the number
+of studies that measure the SvO\textsubscript{2} of the CCV, or any of
+its individual veins, in infants is fairly lower than the number of
+studies investigating the oxygenation of the SSS. Our comparison is
+summarized in Table~\ref{tbl-litvalues}.
+
+\begingroup\fontsize{9}{11}\selectfont
+
+\begin{ThreePartTable}
+\begin{TableNotes}[para]
+\item PT-TEA = born preterm and scanned at term-equivalent age; late third trimester = >35 weeks gestational age; HIE = hypoxic-ischemic encephalopathy; TRUST = T2-relaxation-under-spin tagging; aTRUPC = accelerated T2-relaxation-under-phase-contrast; T2-TRIR = T2-tissue-relaxation-inversion-recovery; SWI = susceptibility weighted imaging.
+\end{TableNotes}
+
+\begin{longtable}[t]{ccllr}
+
+\caption{\label{tbl-litvalues}\textbf{Cerebral oxygenation values of
+neonates and fetuses in the literature.}}
+
+\tabularnewline
+
+\toprule
+Region & Study & Subjects & Method & SvO$_{2}$ (\%)\\
+\midrule
+Whole-brain & \makecell[l]{Skov et al. \\(1993)    } & \makecell[l]{Preterm neonates \\(n=9)                   } & NIRS & 53.4 ± 15.4\\
+Whole-brain & \makecell[l]{Skov et al. \\(1993)    } & \makecell[l]{Asphyxiated term \\neonates (n=10)           } & NIRS & 67.3 ± 9.4\\
+Whole-brain & \makecell[l]{Altman et al. \\(1993)    } & \makecell[l]{Preterm and term \\neonates with HIE \\(n=11)} & PET & 21.6 ± 21.0\\
+SSS & \makecell[l]{Gou et al. \\(2024)     } & \makecell[l]{Healthy neonates \\(n=37)                    } & \makecell[l]{MRI: \\TRUST   } & 66.7 ± 4.9\\
+SSS & \makecell[l]{Jiang et al. \\(2019)   } & \makecell[l]{Healthy neonates \\(n=4)                     } & \makecell[l]{MRI: \\aTRUPC  } & 64.8 ± 2.9\\
+SSS & \makecell[l]{De Vis et al. \\(2014)  } & \makecell[l]{PT-TEA \\eonates (n=18)                      } & \makecell[l]{MRI: \\T2-TRIR } & 52.0 ± 12.0\\
+SSS & \makecell[l]{Yadav et al. \\(2020)} & \makecell[l]{Late third \\trimester fetuses \\(n=33)} & \makecell[l]{MRI: \\Susceptometry} & 58.6 ± 4.8\\
+SSS & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\QSM} & 72.46 ± 3.43\\
+SSS & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\$\chi$-separation} & 58.14 ± 7.3\\
+CCV & Weber et al. (2021) & \makecell[l]{Preterm neonates \\with HIE (n=8)} & \makecell[l]{MRI: \\QSM} & 72.2 ± 6.0\\
+CCV & Weber et al. (2021) & \makecell[l]{Healthy neonates \\(n=8)} & \makecell[l]{MRI: \\QSM} & 73.6 ± 2.8\\
+CCV & Jiang et al. (2019) & \makecell[l]{Healthy neonates \\(n=4)} & \makecell[l]{MRI: \\aTRUPC} & 70.2 ± 3.3\\
+CCV & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\QSM} & 68.71 ± 0.03\\
+CCV & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\$\chi$-separation} & 57.69 ± 6.97\\
+\bottomrule
+\insertTableNotes
+
+\end{longtable}
+
+\end{ThreePartTable}
+\endgroup{}
+
+As shown in Table~\ref{tbl-litvalues}, the infants observed in MRI
+studies investigating cerebral vein oxygenation noticeably differ in
+clinical status, with three studies involving healthy neonates
+\citep{weberQuantitativeSusceptibilityMapping2021, gouAutomaticRejectionBased2024, jiangVesselspecificQuantificationNeonatal2019},
+three studies (including the present study) involving preterm neonates
+\citep{weberQuantitativeSusceptibilityMapping2021, devisNoninvasiveMRIMeasurements2014},
+and one study involving late third trimester fetuses
+\citep{portnoyHumanUmbilicalCord2018}. In the studies involving healthy
+neonates, the SvO\textsubscript{2} of the SSS fell within the range of
+64.8\% -- 66.6\%
+\citep{gouAutomaticRejectionBased2024, jiangVesselspecificQuantificationNeonatal2019},
+while the SvO\textsubscript{2} of the CCV fell within the range of
+70.2\% - 73.6\%
+\citep{weberQuantitativeSusceptibilityMapping2021, jiangVesselspecificQuantificationNeonatal2019}.
+Notably, the SvO\textsubscript{2} value of the SSS we obtained from
+susceptibility separation (58.14\%) was closest to values obtained from
+the studies involving late third trimester fetuses
+\citep{yadavImagingPutativeFoetal2018} or pre-term neonates
+\citep{devisNoninvasiveMRIMeasurements2014}, each finding an SSS
+SvO\textsubscript{2} value of 58.6\% and 52.0\%, respectively. It is
+important to note the difference in MRI modalities used to obtain these
+values. For their study, \citet{yadavImagingPutativeFoetal2018} used MR
+susceptometry, which involves measuring the difference in phase between
+the chosen vessel and its background in imaging data from an SWI
+scanning sequence \citep{yadavImagingPutativeFoetal2018}. In
+\citet{devisNoninvasiveMRIMeasurements2014}, the authors used T2-TRIR,
+which allowed them to determine the transverse relaxation rate of blood
+within the vessel, which can be used alongside hematocrit data to
+estimate SvO\textsubscript{2}. Additionally, the GA of infants scanned
+in our study ranged between 37 and 44 weeks, while the GA of the fetuses
+scanned in \citet{yadavImagingPutativeFoetal2018} was ≥35 weeks and the
+GA of infants scanned in \citet{devisNoninvasiveMRIMeasurements2014}
+ranged between 38 and 40 weeks. As such, our SSS SvO\textsubscript{2}
+values found through susceptibility separation show promise given their
+similarity to the SvO\textsubscript{2} values found by
+\citet{yadavImagingPutativeFoetal2018} and
+\citet{devisNoninvasiveMRIMeasurements2014}, two studies that involved
+comparable subject populations and used considerably different methods.
+
+Conversely, the SvO\textsubscript{2} value of the CCV we obtained
+through QSM (68.71\%) was closest to values from similar studies in the
+literature. In \citet{weberQuantitativeSusceptibilityMapping2021}, QSM
+was used to measure an SvO\textsubscript{2} of 71.5\% in preterm
+neonates with HIE and an SvO\textsubscript{2} of 73.6\% in healthy
+neonates. In their study,
+\citet{jiangVesselspecificQuantificationNeonatal2019} also involved
+healthy neonates and obtained an SvO\textsubscript{2} of 70.2\% through
+an accelerated TRUPC sequence. In contrast, the SvO\textsubscript{2} of
+the CCV we obtained through susceptibility separation was 57.69\%. This
+disparity from the literature, however, may not undermine the value we
+obtained, as the study design of
+\citet{weberQuantitativeSusceptibilityMapping2021} and
+\citet{jiangVesselspecificQuantificationNeonatal2019} may prevent their
+values from being representative of the study demographic. In
+\citet{weberQuantitativeSusceptibilityMapping2021}, the authors utilized
+an arbitrary 0.15 ppm threshold and included all \(\chi\) values above
+0.15 when measuring the mean \(\chi\) of the CCV, which potentially led
+to the introduction of \(\chi\) from veins outside the CCV. In
+\citet{jiangVesselspecificQuantificationNeonatal2019}, the authors
+acquired their data from 4 subjects, a notably small sample size. Given
+the limitations of the existing literature and the similarity of the
+mean paramagnetic CCV SvO\textsubscript{2} value (57.69\%) to the mean
+paramagnetic SSS SvO\textsubscript{2} value (58.14\%), it is plausible
+that susceptibility separation provides more accurate measurements of
+oxygenation in both cortical and subcortical veins. One reason for this
+is due to its ability to mitigate partial volume effects, which are
+likely to contaminate other methods resulting in inaccurate
+CSvO\textsubscript{2} values
+\citep{shinHseparationMagneticSusceptibility2021}.
+
+Another notable distinction between our findings and those of the
+existing literature was that we observed no significant oxygenation
+difference between the SSS and the CCV when \(\chi\) was derived from
+paramagnetic maps.
+\citet{jiangVesselspecificQuantificationNeonatal2019}, the only other
+study that also measured SvO\textsubscript{2} in both the SSS and CCV,
+observed significantly lower oxygenation in the SSS (64.8\%) when
+compared to the CCV (70.2\%). Given the small sample size utilized by
+\citet{jiangVesselspecificQuantificationNeonatal2019}, it is difficult
+to ascertain whether this is generalizable to all neonates.
+
+\subsection{Limitations and future
+directions}\label{limitations-and-future-directions}
+
+This study has a few limitations that should be considered for future
+research. Firstly, only 19 infants were recruited for scanning. Given
+the emotional toll placed on parents when their child is born preterm,
+it is understandable that they may show reluctance in consenting to
+further testing that is not medically necessary. Obtaining a larger
+sample size in future studies, however, may provide greater insight into
+the efficacy of susceptibility separation. Secondly, this study did not
+include healthy neonates born at term, resulting in a lack of a control
+cohort. This is because recruiting healthy controls when there is no
+contraindication is very difficult. The addition of such a group may
+provide further validity to any findings and may reveal potential
+differences in cerebral oxygen consumption between term and preterm
+neonates. Finally, future studies should consider the use of multi-echo
+T2 imaging data when performing the decomposition of QSM images. The
+toolbox applied by this study for QSM decomposition
+\citep{shinHseparationMagneticSusceptibility2021} utilizes R2 data,
+which can be obtained from multi-echo T2 imaging. Our study protocol
+involved the collection of multi-echo SWI imaging data, and as such, we
+could only use R2\textsuperscript{*} data to perform the decomposition.
+Furthermore, this may account for the reduced precision of
+SvO\textsubscript{2} values obtained through susceptibility separation.
+
 \section{Conclusion}\label{sec-conclusion}
+
+This study aimed to evaluate how the use of susceptibility separation on
+preterm neonatal QSM images in order to provide greater accuracy in
+determining the oxygenation of cerebral venous vessels. We compared
+venous specific SvO\textsubscript{2} values obtained from QSM images and
+their respective paramagnetic components to SvO\textsubscript{2} values
+from neonatal MRI studies. We found that susceptibility separation
+provided SvO\textsubscript{2} values of the SSS that were comparable to
+values found in the literature, providing evidence that this processing
+technique may be a valid tool for measuring regional cerebral oxygen
+consumption. Additionally, we were able to simultaneously measure
+SvO\textsubscript{2} in both the SSS and CCV, which permitted us to
+observe no difference in oxygenation between the two vessels when
+considering data from isolated paramagnetic components. Ultimately, we
+hope our work inspires future studies that seek to explore and improve
+the capabilities of magnetic susceptibility separation, culminating in
+the development of a tool for clinicians and researchers alike.
 
@@
 
-\phantomsection\label{refs}
-\begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-ahmedDiamagneticComponentMap2023a}
-Ahmed, Maruf, Jingjia Chen, Arvin Arani, Matthew L. Senjem, Petrice M.
-Cogswell, Clifford R. Jack, and Chunlei Liu. 2023. {``The Diamagnetic
-Component Map from Quantitative Susceptibility Mapping ({QSM}) Source
-Separation Reveals Pathological Alteration in {Alzheimer}'s
-Disease-Driven Neurodegeneration.''} \emph{NeuroImage} 280 (October):
-120357. \url{https://doi.org/10.1016/j.neuroimage.2023.120357}.
-
-\bibitem[\citeproctext]{ref-bergInvestigatingEffectFlow2021}
-Berg, Ronja C., Christine Preibisch, David L. Thomas, Karin Shmueli, and
-Emma Biondetti. 2021. {``Investigating the {Effect} of {Flow
-Compensation} and {Quantitative Susceptibility Mapping Method} on the
-{Accuracy} of {Venous Susceptibility Measurement}.''} bioRxiv.
-\url{https://doi.org/10.1101/2021.04.14.439812}.
-
-\bibitem[\citeproctext]{ref-dixMonitoringCerebralOxygenation2017}
-Dix, Laura Marie Louise, Frank van Bel, and Petra Maria Anna Lemmers.
-2017. {``Monitoring {Cerebral Oxygenation} in {Neonates}: {An
-Update}.''} \emph{Frontiers in Pediatrics} 5.
-
-\bibitem[\citeproctext]{ref-kamesRapidTwostepDipole2018}
-Kames, Christian, Vanessa Wiggermann, and Alexander Rauscher. 2018.
-{``Rapid Two-Step Dipole Inversion for Susceptibility Mapping with
-Sparsity Priors.''} \emph{NeuroImage} 167 (February): 276--83.
-\url{https://doi.org/10.1016/j.neuroimage.2017.11.018}.
-
-\bibitem[\citeproctext]{ref-liIntegratedLaplacianBased2014}
-Li, Wei, Alexandru V. Avram, Bing Wu, Xue Xiao, and Chunlei Liu. 2014.
-{``Integrated {Laplacian}-Based Phase Unwrapping and Background Phase
-Removal for Quantitative Susceptibility Mapping.''} \emph{NMR in
-Biomedicine} 27 (2): 219--27. \url{https://doi.org/10.1002/nbm.3056}.
-
-\bibitem[\citeproctext]{ref-liFirstStepNeuroimaging2016}
-Li, Xiangrui, Paul S. Morgan, John Ashburner, Jolinda Smith, and
-Christopher Rorden. 2016. {``The First Step for Neuroimaging Data
-Analysis: {DICOM} to {NIfTI} Conversion.''} \emph{Journal of
-Neuroscience Methods} 264 (March): 47--56.
-\url{https://doi.org/10.1016/j.jneumeth.2016.03.001}.
-
-\bibitem[\citeproctext]{ref-mccarthyFSLeyes2023}
-McCarthy, Paul. 2023. {``{FSLeyes}.''} Zenodo.
-\url{https://doi.org/10.5281/zenodo.8376979}.
-
-\bibitem[\citeproctext]{ref-portnoyHumanUmbilicalCord2018}
-Portnoy, Sharon, Natasha Milligan, Mike Seed, John G. Sled, and
-Christopher K. Macgowan. 2018. {``Human Umbilical Cord Blood Relaxation
-Times and Susceptibility at 3 {T}: {Human Umbilical Cord Blood
-Relaxation Times} and {Susceptibility} at 3 {T}.''} \emph{Magnetic
-Resonance in Medicine} 79 (6): 3194--3206.
-\url{https://doi.org/10.1002/mrm.26978}.
-
-\bibitem[\citeproctext]{ref-rcoreteamLanguageEnvironmentStatistical2022}
-R Core Team. 2022. {``R: {A Language} and {Environment} for {Statistical
-Computing}.''} Vienna, Austria: R Foundation for Statistical Computing.
-
-\bibitem[\citeproctext]{ref-rantakariEarlyOxygenLevels2021}
-Rantakari, Krista, Olli-Pekka Rinta-Koski, Marjo Metsäranta, Jaakko
-Hollmén, Simo Särkkä, Petri Rahkonen, Aulikki Lano, et al. 2021.
-{``Early Oxygen Levels Contribute to Brain Injury in Extremely Preterm
-Infants.''} \emph{Pediatric Research} 90 (1): 131--39.
-\url{https://doi.org/10.1038/s41390-021-01460-3}.
-
-\bibitem[\citeproctext]{ref-rstudioteamRStudioIntegratedDevelopment}
-RStudio Team. n.d. {``{RStudio}: {Integrated Development Environment}
-for {R}.''} Boston, MA: RStudio, PBC.
-
-\bibitem[\citeproctext]{ref-sedlacikObtainingBloodOxygenation2007}
-Sedlacik, Jan, Alexander Rauscher, and Jürgen R. Reichenbach. 2007.
-{``Obtaining Blood Oxygenation Levels from {MR} Signal Behavior in the
-Presence of Single Venous Vessels.''} \emph{Magnetic Resonance in
-Medicine} 58 (5): 1035--44. \url{https://doi.org/10.1002/mrm.21283}.
-
-\bibitem[\citeproctext]{ref-shinHseparationMagneticSusceptibility2021}
-Shin, Hyeong-Geol, Jingu Lee, Young Hyun Yun, Seong Ho Yoo, Jinhee Jang,
-Se-Hong Oh, Yoonho Nam, et al. 2021. {``{\(\chi\)}-Separation:
-{Magnetic} Susceptibility Source Separation Toward Iron and Myelin
-Mapping in the Brain.''} \emph{NeuroImage} 240 (October): 118371.
-\url{https://doi.org/10.1016/j.neuroimage.2021.118371}.
-
-\bibitem[\citeproctext]{ref-smithFastRobustAutomated2002}
-Smith, Stephen M. 2002. {``Fast Robust Automated Brain Extraction.''}
-\emph{Human Brain Mapping} 17 (3): 143--55.
-\url{https://doi.org/10.1002/hbm.10062}.
-
-\bibitem[\citeproctext]{ref-weisskoffMRISusceptometryImagebased1992}
-Weisskoff, Robert M., and Suzanne Kiihne. 1992. {``{MRI} Susceptometry:
-{Image-based} Measurement of Absolute Susceptibility of {MR} Contrast
-Agents and Human Blood: {COMMUNICATIONS}.''} \emph{Magnetic Resonance in
-Medicine} 24 (2): 375--83. \url{https://doi.org/10.1002/mrm.1910240219}.
-
-\bibitem[\citeproctext]{ref-woolrichBayesianAnalysisNeuroimaging2009}
-Woolrich, Mark W., Saad Jbabdi, Brian Patenaude, Michael Chappell,
-Salima Makni, Timothy Behrens, Christian Beckmann, Mark Jenkinson, and
-Stephen M. Smith. 2009. {``Bayesian Analysis of Neuroimaging Data in
-{FSL}.''} \emph{NeuroImage} 45 (1): S173--86.
-\url{https://doi.org/10.1016/j.neuroimage.2008.10.055}.
-
-\bibitem[\citeproctext]{ref-zhu-etal-cmro2}
-Zhu, A., C. Chau, N. Chan, A. Chacko, L. Holsti, R. E. Grunau, and A. M.
-Weber. 2024. {``Regional Cerebral Metabolic Rate of Oxygen and Levels of
-Respiratory Support in Preterm Neonates.''} \emph{Pediatric Research},
-May.
-
-\end{CSLReferences}
+\renewcommand{\bibsection}{}
+\bibliography{Gavin_Thesis_Ref.bib}
 

### 2025-02-15 PediatricResearchSubmission_Oct2024/ReviewerResponse_2024Dec03/ResponsetoReviewers.md
final

---
title: Response to Reviewers
geometry: margin=1in
---

Reviewer's original suggestions are in quote blocks and in **bold**  
Our response follows in normal font, and changes to the document are in quote blocks and normal font.

# Editor Comments to Author

## Section Editor: 1

> **Comments to the Author:**  
> **This manuscript by Weber et al. is focused on oxygenation quantification in the SSS and CCV in 19 premature infants utilizing susceptibility mapping and paramagnetic separation, which is compared directly and thoroughly with the preceding literature using a variety of techniques. This is significant in that it may provide an additional, and hopefully more accurate, way to measure oxygenation in premature infants. Overall this is an impressive and well planned study.**

Thank you for your kind words and for taking the time to review our manuscript.

---

> **The main issue which needs to be addressed revolves around the conclusions, and that the presented data do not support that this has superior accuracy. It is important because some of the findings contradict previous studies. The language here needs to be softened**

We agree. It is important that our conclusions do not overstate the evidence provided. Based on the editor's recommendation, we have now softened our language in the abstract, impact statement, and conclusion:

Abstract:

> While paramagnetic component decomposition yielded SSS values closer to those found in the literature, it increased variability. No significant oxygenation differences were found between the SSS and CCV, contrasting with prior studies.

Impact Statement:

> This study evaluated the use of QSM and its paramagnetic components to measure cerebral oxygenation in neonates.

Conclusion:

> This study aimed to evaluate how the use of susceptibility separation on preterm neonatal QSM images could be used in determining the oxygenation of cerebral venous vessels

---

> **In the methods, there are 3 lines where "\*\*\*" appears, I am not sure what that is supposed to represent (lines 102, 104, 117).**

These asterisks' were included in order to abide by what we believed *Pediatric Research's* submission guidelines were, to anonymize the paper from reviewers. These will be replaced with the identifying information once accepted. Or, if we made a mistake, we are happy to de-anonymize them for the reviewers.

---

> **Requires additional clarification related to imaging analysis and how it leads to their conclusion**

We agree with the editor and with the reviewers below who made similar comments. Please see our responses below for a detailed summary of the changes we made.

# Reviewer(s)' Comments to Author

## Reviewer: 1

> **Comments to the Author**  
> **Summary:**  
> **This study aims to quantify oxygenation in the superior sagittal sinus and in central cerebral veins in a cohort of 19 preterm neonates using quantitative susceptibility mapping and the separation of its paramagnetic components.**  
> **The authors observed no significant differences between the oxygenation of the SSS and CCV, which contradicts previous results reported in the literature. They reported that the accuracy of the paramagnetic source separation was higher than QSM, but at the cost of an increased variability.**  
> **The methods of the manuscript are well detailed, and the results are appropriately compared to the literature. A few aspects require some clarification, see below.**

We thank the reviewer for taking the time to read our manuscript and for making suggestions to help improve the final result.

---

> **One limitation of this work is the lack of sufficient data to support the claim of increased accuracy with susceptibility sources separation compared to QSM. In the absence of a ground truth and with results that contradict literature, this work does not support the increased accuracy claimed. The conclusion of the work should be re-written to better reflect the actual results.**

We agree with the reviewer. Without a ground truth to compare to, it is difficult to say if our results reflect a more accurate measure. We have now included wording saying as much, and rewritten the conclusion as follows:

> This study aimed to evaluate how the use of susceptibility separation on preterm neonatal QSM images could be used in determining the oxygenation of cerebral venous vessels

The impact statement was also changed:

> This study evaluated the use of QSM and its paramagnetic components to measure cerebral oxygenation in neonates.

Along with the abstract:

> While paramagnetic component decomposition yielded SSS values closer to those found in the literature, it increased variability. No significant oxygenation differences were found between the SSS and CCV, contrasting with prior studies.

---

> **Specific Issues:**  
> **Abstract:**  
> **1. Please re-write the first sentence of the conclusion section of the abstract to better reflect the results presented in this work, which do not fully support the current claim of increased accuracy with susceptibility source separation.**

The first sentence of the conclusion of the abstract has been changed to better reflect our findings:

> While paramagnetic component decomposition yielded SSS values closer to those found in the literature, it increased variability. No significant oxygenation differences were found between the SSS and CCV, contrasting with prior studies.

---

> **Introduction:**  
> **2. It is mentioned that methods to assess whole brain oxygenation have been developed for adults and that they are starting to be explored in neonates. Please detail what are the challenges for transferring these methods to neonate population.**

We have adjusted this section to now read:

> For the preceding reasons, non-invasive MRI-based techniques are actively being explored to assess regional and whole-brain blood oxygenation. While MRI-based methods have been developed for adults (9-11), their application in neonates is only beginning to be explored (12-16). This delay is likely due to the unique challenges posed by neonates, including their smaller anatomies, distinct hemodynamic profiles, susceptibility to motion artifacts, and the difficulties associated with recruiting this population for research.

---

> **Method:**  
> **Image acquisition**  
>**3. Please state in the text the plane of acquisition for each sequence.**

We now include the plane of acquisition for each sequence in the text of the Image Acquisition section.

---

>**4. Table 1: The phase encoding row (2nd row) does not provide the phase encoding information. This row should be split in two, one for the acquisition plane information (axial, coronal, sagittal), and one for the actual phase encoding direction in plane (ex: right-left, antero-posterior, or head-foot).**

We thank the reviewer for catching this. We have now updated the table to reflect these changes by adding a row for Acquisition plane, and fixing the Phase-encoding direction

---

> **Image analysis:**  
> **5. Briefly explain why the fifth echo was selected to generate the preliminary brain mask.**

The last (fifth) echo was used to create the brain mask as the last echo will have the greatest signal decay from air-bone-tissue interfaces. This is important to use in order to properly avoid introducing aliasing in the inverse problem. We now include an explanation of this in our methods for readers:

> The last echo is used to generate the brain mask as it reliably removes artifacts from air-tissue and bone-tissue interfaces, e.g., sinuses, without the need for manual erosion. The reason for this is at longer echo times, tissues with rapid signal decay (such as bone, air, and sinuses) lose their MRI signal due to dephasing caused by greater magnetic field inhomogeneities.

---

> **6. Please clarify if all echoes were used for the preliminary QSM calculation.**

We apologise for the confusing way our methods section was written. The methods now describe that the fifth echo was used for mask creation; the STI suite used all five echoes to create five separate QSM images; the 3rd, 4th and 5th QSM maps were then averaged to create the final QSM image for which chi values are taken. The $\chi$-separation toolbox outputs a single map for diamagnetic, paramagnetic, and QSM. Please see the tracked changes in the methods section, and our response to #8 below.

---

> **7. Can the authors expand on the motivation for choosing the STI software? What does ‘cleanest images’ refer to exactly (accuracy, artifact free, etc.)? Please also add a reference to support this statement.**

The STI software appeared to give the best images, based on a visual check by the authors. As this was based on our opinion, we cannot provide a reference for this.  
We now include an explanation of this in the methods:

> STI Suite (v. 3.0), was used to process the final QSM images as it produced the images with the least amount of artifacts (based on a visual assessment by the authors) without eroding the SSS.

---

> **8. Can the authors explain why only the last three echoes were used for the STI QSM calculation?**

This was an error in reporting we have now fixed. Please see the response below.

---

> **9. The methods description mentions that the last three echoes only were used for the STI calculation, but then in the parameters listed in the text, it is mentioned that a TE1=5 ms was used, which is the first TE of the sequence. Can the authors clarify this?**

We apologise for this. The way it was written was not clear. We used all five echoes to produce a five echo QSM image, but then averaged the last three echoes to produce the final QSM image. The section now reads:

> The finalized brain mask along with the magnitude and phase images were used in STI Suite along with the following parameters: 0.9766 x 0.9766 x 1 mm3 voxel size, 5 ms TE1, 5.24 ms TE, and 77.4 ms sum TE, B0 strength = 3, and B0 direction = [0, 0, 1]. The 3D GRE data option was selected for the phase processing stage, and STAR-QSM was selected for the QSM stage. The last three echoes of the QSM were then averaged to create the final QSM image using fslmaths, as the accumulation of phase due to susceptibility is small in early echoes and artifacts dominate the phase.

---

> **10. In the paragraph describing the methods for the paramagnetic source separation, please list the TEs in ms unit instead of s for consistency with the rest of the manuscript. Also in the same paragraph, please use the acronym introduce earlier in the text for delta TE for consistency.**

We thank the reviewer for catching these. We originally stated the TEs in seconds in this section as that is what the paramagnetic source software uses when entering units. However, we believe that using consistent units throughout the manuscript would be best. We have made this correction in this section.

---

> **11. Can the authors include an illustration of the final SSS and CCV masks used to calculate mean susceptibility in these regions?**

Yes of course. We agree that this would be a good addition to the paper.

Please see the new figure now numbered Figure 2.

---

> **Results:**  
> **12. Figure 2: Please add a label to the colorbars (ex: chi (ppm)). Also, please clarify in the caption what the x/y/z label on each figure refers to.**

"$\chi$ (ppm)" has now been added to the two colorbars. In addition, we now include the following in the caption:

> The color bars in (b) and \(c\) indicates the range of susceptibility $\chi$ values in parts per million.

The caption also now includes a sentence at the end that reads:

> y, x and z values represent the slice number in each plane (coronal, sagittal, and axial, respectively).

---

> **13. Figure 3: In the text referring to fig 3, the graphs presented are referred to as boxplots, versus in the figure caption, they are referred to as violin plots. Please make sure to consistently use the same term.**

We apologise for missing this. The text has been changed to reflect the appropriate plot name: 'violin plots'.

> Violin plots of the comparisons are shown in Figure 3.

---

> **Discussion**  
> **14. The authors mention being the first to explore the susceptibility source separation technique in neonates, while it has been done in adult population. Please discuss the challenges of transferring this technique to a neonate population and how these were address in this work.**

Often times new techniques will first be used in adults before eventually being used in neonatal populations. Part of this is due to the relative difficulty in acquiring ethics and recruiting subjects for neonatal research. My lab has a focus on neonatal brain development and I happen to be interested in QSM. We believe we are the first to perform this type of analysis in a neonatal population simply due to a combination of luck, access to data, and timing. However, please see our response to #2 above where we now include some of the challenges in scanning neonates in the introduction.

---

> **15. Can the authors discuss why, based on their results, QSM underestimates the susceptibility compared to the paramagnetic component separation technique?**

Reading over the manuscript, we realise this is a very important concept that should have been included in the introduction. We now include the following sentence near the end of the introduction:

> Furthermore, QSM tends to underestimates parametric components due to the inclusion of diamagnetic tissue, and vice versa, as the opposing magnetic susceptibilities effectively subtract from one another.

### 2025-02-15 index.qmd
yaml adjustments
final

@@
     name: Pediatrics, The University of British Columbia, Vancouver, BC, Canada
+  - id: 4
+    name: Physics and Astronomy, The University of British Columbia, Vancouver, BC, Canada
 
@@
 
-abstract: |
-  **Background**: Quantitative susceptibility mapping (QSM), an magnetic resonance imaging (MRI) modality sensitive to deoxyhemoglobin, is a promising method for measuring cerebral oxygenation in human neonates.
+filters:
+  - authors-block
+
+notebook-links: false
+number-sections: false
+citation:
+  container-title: Unpublished
+# date: last-modified
+bibliography: Gavin_Thesis_Ref.bib
+csl: nature.csl
+
+crossref:
+  fig-title: '**Figure**'
+  tbl-title: '**Table**'
+  title-delim: "**.**"
+---
+
+## Impact Statement
+
+-   This study evaluated the use of QSM and its paramagnetic components to measure cerebral oxygenation in neonates.
+-   By comparing susceptibility-derived oxygen saturation (SvO~2~) in the superior sagittal sinus (SSS) and central cerebral veins (CCV), it adds to the field of neonatal cerebral oxygenation measurement.
+-   Decomposing QSM into paramagnetic components shows potential for improving SvO~2~ accuracy, particularly in the SSS, though variability remains a challenge.
+-   The results suggest no significant oxygenation difference between the SSS and CCV, contrasting with previous findings, indicating a need for further research on neonatal venous oxygenation.
+
+**Category of Study:** basic science
+
+{{< pagebreak >}}
+
+## Abstract:
+
+**Background**: Quantitative susceptibility mapping (QSM), a magnetic resonance imaging (MRI) modality sensitive to deoxyhemoglobin, is a promising method for measuring cerebral oxygenation in human neonates.
 Paramagnetic sources, like deoxyhemoglobin, however, can be obscured by diamagnetic sources such as water and myelin.
@@
 
-  **Conclusion**: While paramagnetic component decomposition yielded more accurate SSS values, it increased variability.
+**Conclusion**: While paramagnetic component decomposition yielded SSS values closer to those found in the literature, it increased variability.
 No significant oxygenation differences were found between the SSS and CCV, contrasting with prior studies.
-
-filters:
-  - authors-block
-
-bibliography: Gavin_Thesis_Ref.bib
-csl: pediatric-research.csl
-
-crossref:
-  fig-title: '**Figure**'
-  tbl-title: '**Table**'
-  title-delim: "**.**"
----
-
-**Category of Study:** basic science
-
-## Impact Statement
-
-- This study evaluated the accuracy of using QSM and its paramagnetic components to measure cerebral oxygenation in neonates.
-- By comparing susceptibility-derived oxygen saturation (SvO~2~) in the superior sagittal sinus (SSS) and central cerebral veins (CCV), it adds to the field of neonatal cerebral oxygenation measurement.
-- Decomposing QSM into paramagnetic components shows potential for improving SvO~2~ accuracy, particularly in the SSS, though variability remains a challenge.
-- The results suggest no significant oxygenation difference between the SSS and CCV, contrasting with previous findings, indicating a need for further research on neonatal venous oxygenation.
 
@@
 
-## Acknowledgments
-
-We wish to acknowledge the work of and thank Victoria Tapics (Research Nurse); Vicki Goh (Research Nurse); Chacko Anil (Neonatologist); and Michael A. Sargent (Radiologist).
-
-## Funding
-
-Authors AMW and REG were co-primary applicants for a BC Children’s Hospital Research Institute - Brain, Behaviour and Development Catalyst Grant ($20,000).
-AMW was supported by an establishment award from BCCHRI.
-Scanning was partly funded through a special award to AMW from BCCHRI .
-
-## Author contributions
-
-TGC wrote the original draft, performed the formal analysis, and contributed to methodology, validation, and visualization.
-AR helped with writing, reviewing & editing.
-REG helped with writing, reviewing & editing, and with initial funding acquisition.
-AMW was involved in project administration, supervision, validation, visualization, resources, methodology, formal analysis, funding acquisition, writing, reviewing & editing, conceptualization, data curation, and investigation.
-
-## Competing interests
-
-The authors have no competing interests to declare.
-
-## Consent statement
-
-The study was approved by the Clinical Research Ethics Board at the University of British Columbia and Children's & Women's Hospital (H21-00655) and written informed consent was obtained from the parent/guardian for each infant.
-
 {{< pagebreak >}}
@@
 For the preceding reasons, non-invasive MRI-based techniques are actively being explored to assess regional and whole-brain blood oxygenation.
-MRI based methods have been developed for adults [@jainRapidMagneticResonance2011; @luQuantitativeEvaluationOxygenation2008; @xuNoninvasiveQuantificationWholebrain2009] and are only beginning to be explored in neonates [@devisNoninvasiveMRIMeasurements2014; @liuQuantitativeAssessmentGlobal2014; @qiHemodynamicMetabolicAssessment2018; @jainCerebralOxygenMetabolism2014; @jiangVesselspecificQuantificationNeonatal2019].
-These methods almost all rely on T2 relaxation to estimate CSvO~2~ [@devisNoninvasiveMRIMeasurements2014; @liuQuantitativeAssessmentGlobal2014; @qiHemodynamicMetabolicAssessment2018; @jiangVesselspecificQuantificationNeonatal2019] with the exception of @jainCerebralOxygenMetabolism2014 which used susceptometry [@jainMRIEstimationGlobal2010].
+While MRI-based methods have been developed for adults [@jainRapidMagneticResonance2011; @luQuantitativeEvaluationOxygenation2008; @xuNoninvasiveQuantificationWholebrain2009], their application in neonates is only beginning to be explored [@devisNoninvasiveMRIMeasurements2014; @liuQuantitativeAssessmentGlobal2014; @qiHemodynamicMetabolicAssessment2018; @jainCerebralOxygenMetabolism2014; @jiangVesselspecificQuantificationNeonatal2019].
+This delay is likely due to the unique challenges posed by neonates, including their smaller anatomies, distinct hemodynamic profiles, susceptibility to motion artifacts, and the difficulties associated with recruiting this population for research.
+These methods have almost all relied on T2 relaxation to estimate CSvO~2~ [@devisNoninvasiveMRIMeasurements2014; @liuQuantitativeAssessmentGlobal2014; @qiHemodynamicMetabolicAssessment2018; @jiangVesselspecificQuantificationNeonatal2019] with the exception of @jainCerebralOxygenMetabolism2014 which used susceptometry [@jainMRIEstimationGlobal2010].
 One limitation of these T2 relaxation methods, however, is the fact that SvO~2~ is often measured using a single imaging slice, averaging values across several voxels, and only in the superior sagittal sinus (SSS).
 In the case of @jainCerebralOxygenMetabolism2014, they obtained regional and whole-brain data, but with thick slices (5mm), and still only estimated CSvO~2~ in the SSS.
-An alternative MRI method using quantitative susceptibility mapping (QSM) has been proposed, which can measure SvO~2~ regionally and across the whole-brain at high resolution (< 1mm^3^ per voxel) [@weberQuantitativeSusceptibilityMapping2021].
+An alternative MRI method using quantitative susceptibility mapping (QSM) has been proposed, which can measure SvO~2~ regionally and across the whole-brain at high resolution (\< 1mm^3^ per voxel) [@weberQuantitativeSusceptibilityMapping2021].
 However, this method left room for improvement, as it removed the SSS (averaging CSvO~2~ across the internal veins), and required an arbitrary threshold value of 0.15 ppm in order to acquire realistic results [@weberQuantitativeSusceptibilityMapping2021].
+Furthermore, QSM tends to underestimates parametric components due to the inclusion of diamagnetic tissue, and vice versa, as the opposing magnetic susceptibilities effectively subtract from one another [@kimHSeparationImagingDiagnosis2023].
 
@@
 
-Participant data comes from a previous study *** <!-- [@zhu-etal-cmro2] (submitted) -->.
-Participants consisted of preterm neonates born between 25- and 31-weeks gestational age (GA) who were admitted to the level III NICU at *** <!-- BC Women’s Hospital -->.
+Participant data comes from a previous study \*\*\* <!-- [@zhu-etal-cmro2] (submitted) -->.
+Participants consisted of preterm neonates born between 25- and 31-weeks gestational age (GA) who were admitted to the level III NICU at \*\*\* <!-- BC Women’s Hospital -->.
 Recruitment took place over a span of one year, from February 2021 to January 2022, facilitated by a dedicated research nurse.
@@
 
-MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750 scanner (scanner software version DV26.0_R03) equipped with a SREE Medical Systems (Cleveland, OH, USA) single-channel neonatal head coil (@tbl-mri).
-The scans were conducted at the *** <!-- BC Children -->’s MRI Research Facility.
+MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750 scanner (scanner software version DV26.0_R03) equipped with a SREE Medical Systems (Cleveland, OH) single-channel neonatal head coil (@tbl-mri).
+The scans were conducted at the \*\*\* <!-- BC Children -->’s MRI Research Facility.
 Prior to the scanning procedure, subjects were carefully prepared by a research nurse.
@@
 df[1,] <- c("Sequence", "3D FSPGR", "3D CUBE", "Multi-shot 3D fast spin-echo", "3D spoiled GRE flow-compensated")
-df[2,] <- c("Phase-encoding direction", "Coronal", "Sagittal", "Axial", "Axial")
-df[3,] <- c("TR (ms)", "7.74", "2,300", "4,680", "30.9")
-df[4,] <- c("TE (ms)", "2.97", "66.29", "10.55", "5 echoes; first echo: 5; echo spacing: 5.24")
-df[5,] <- c("Flip angle", "12\U00B0", "90\U00B0", "111\U00B0", "20\U00B0")
-df[6,] <- c("FOV (cm)", "20", "20", "24", "25")
-df[7,] <- c("Acquisition matrix", "512 x 512", "256 x 256", "128 x 128", "256 x 256")
-df[8,] <- c("In-plane resolution (mm)", "0.39 x 0.39", "0.78 x 0.78", "1.875 x 1.875", "0.977 x 0.977")
-df[9,] <- c("Slice thickness (mm)", "1", "1", "4", "2, reconstructed to 1 with zero filling (ZIP2)")
-df[10,] <- c("Number of slices", "126", "106", "50", "92")
-df[11,] <- c("Additional parameters", "n/a", "n/a", "1,450 ms label period;\n 2,025 ms pulse label;\n 24 control-label pairs", "n/a")
-df[12,] <- c("Scan duration", "4 min 39 s", "5 min 1 s", "5 min 26 s", "5 min 29 s")
+df[2,] <- c("Acquisition plane", "Coronal", "Sagittal", "Axial", "Axial")
+df[3,] <- c("Phase-encoding direction", "Left-Right", "Posterior-Anterior", "Posterior-Anterior", "Left-Right")
+df[4,] <- c("TR (ms)", "7.74", "2,300", "4,680", "30.9")
+df[5,] <- c("TE (ms)", "2.97", "66.29", "10.55", "5 echoes; first echo: 5; echo spacing: 5.24")
+df[6,] <- c("Flip angle", "12\U00B0", "90\U00B0", "111\U00B0", "20\U00B0")
+df[7,] <- c("FOV (cm)", "20", "20", "24", "25")
+df[8,] <- c("Acquisition matrix", "512 x 512", "256 x 256", "128 x 128", "256 x 256")
+df[9,] <- c("In-plane resolution (mm)", "0.39 x 0.39", "0.78 x 0.78", "1.875 x 1.875", "0.977 x 0.977")
+df[10,] <- c("Slice thickness (mm)", "1", "1", "4", "2, reconstructed to 1 with zero filling (ZIP2)")
+df[11,] <- c("Number of slices", "126", "106", "50", "92")
+df[12,] <- c("Additional parameters", "n/a", "n/a", "1,450 ms label period;\n 2,025 ms pulse label;\n 24 control-label pairs", "n/a")
+df[13,] <- c("Scan duration", "4 min 39 s", "5 min 1 s", "5 min 26 s", "5 min 29 s")
 
@@
 if (knitr::is_latex_output()) {
-colnames(df)[1] <- ""
-df[11,] <- linebreak(df[11,])
+#colnames(df)[1] <- ""
+colnames(df) <- c("", "\\textbf{T1w}", "\\textbf{T2w}", "\\textbf{pcASL}", "\\textbf{SWI}")
+df[12,] <- linebreak(df[12,])
 df |>
@@
     gt(rowname_col = "Scan") |> tab_footnote(footnotetext) |>
-    fmt_markdown(columns = TRUE) |>
+    fmt_markdown(columns = everything()) |>
+    cols_label(
+      T1w = md("**T1w**"),
+      T2w = md("**T2w**"),
+      pcASL = md("**pcASL**"),
+      SWI = md("**SWI**")
+    ) |>
     tab_options(quarto.disable_processing = TRUE)
@@
 
-The MRI scan protocol comprised of the following sequences: a T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin labeling (ASL) scan [@alsopRecommendedImplementationArterial2015], a multi-echo susceptibility-weighted imaging scan [@denkSusceptibilityWeightedImaging2010], and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging (EPI) sequence.
+The MRI scan protocol comprised of the following sequences (plane of acquisition in parentheses): a T1-weighted scan (coronal), a T2-weighted scan (sagittal), a pseudo-continuous arterial spin labeling (ASL) scan [@alsopRecommendedImplementationArterial2015] (axial), a multi-echo susceptibility-weighted imaging scan [@denkSusceptibilityWeightedImaging2010] (axial), and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging (EPI) sequence (axial).
 The DWI sequence was not used for the present study.
@@
 The raw DICOM files acquired from the scanning procedure were converted to NIfTI (Neuroimaging Informatics Technology Initiative) format using Chris Rorden’s `dcmniix` tool [@liFirstStepNeuroimaging2016].
-SWI magnitude data files were then used to create subject-specific brain masks that would not erode the SSS during QSM processing, an issue faced by our group in the past *** <!-- [@weberQuantitativeSusceptibilityMapping2021] -->.
+SWI magnitude data files were then used to create subject-specific brain masks that would not erode the SSS during QSM processing, an issue faced by our group in the past \*\*\* <!-- [@weberQuantitativeSusceptibilityMapping2021] -->.
 A step-by-step summary of the pipeline used is shown in @fig-graph.
-
-{{< embed notebooks/Figures.ipynb#fig-graph >}}
 
@@
 6.0.7.3) [@woolrichBayesianAnalysisNeuroimaging2009] `fslroi`, `fslmaths`, and `bet` [@smithFastRobustAutomated2002] to create a preliminary brain mask, similar to our previous efforts, which does not contain the SSS.
+The last echo is used to generate the brain mask as it reliably removes artifacts from air-tissue and bone-tissue interfaces, e.g., sinuses, without the need for manual erosion.
+The reason for this is at longer echo times, tissues with rapid signal decay (such as bone, air, and sinuses) lose their MRI signal due to dephasing caused by greater magnetic field inhomogeneities.
 `Fslroi` was used to isolate the fifth echo of the magnitude data, which was then squared using `fslmaths` and the option `-sqr`.
 Squaring the magnitude image was found to dramatically improve subsequent brain extraction.
-The resulting image was then used to create the preliminary brain mask using bet with the options `-m` and `-R`.
+The resulting image was then used to create the preliminary brain mask using `bet` with the options `-m` and `-R`.
 The former flag generated a binary brain mask, while the latter performed a more robust brain centre estimation.
@@
 STI Suite (v.
-3.0) [@liIntegratedLaplacianbasedPhase2014], was used to process the final QSM images as it produced the cleanest images without eroding the SSS.
-The finalized brain mask and the last three echoes of the magnitude and phase images were used in STI Suite along with the following parameters: 0.9766 x 0.9766 x 1 mm^3^ voxel size, 5 ms TE1, 5.3 ms $\Delta$TE, and 77.4 ms sum TE, B0 strength = 3, and B0 direction = \[0, 0, 1\].
+3.0) [@liIntegratedLaplacianbasedPhase2014], was used to process the final QSM images as it produced the images with the least amount of artifacts (based on a visual assessment by the authors) without eroding the SSS.
+The finalized brain mask along with all five echoes of the magnitude and phase images were used in STI Suite along with the following parameters: 0.9766 x 0.9766 x 1 mm^3^ voxel size, 5 ms TE1, 5.24 ms $\Delta$TE, and 77.4 ms sum TE, B0 strength = 3, and B0 direction = \[0, 0, 1\].
 The 3D GRE data option was selected for the phase processing stage, and STAR-QSM was selected for the QSM stage.
-Finally, the \`select by intensity' tool in `FSLeyes` was then used to semi-automatically make vascular masks of the SSS and CCV from each subject’s QSM image.
+STAR-QSM outputs a single QSM map for each echo (i.e.
+five total).
+The last three echoes of the QSM maps were then averaged to create the final QSM image using `fslmaths`, as the accumulation of phase due to susceptibility is small in early echoes and artifacts dominate the phase [@zhangQuantitativeAnalysisPunctate2019].
+Finally, the \`select by intensity' tool in `FSLeyes` was then used to semi-automatically make vascular masks of the SSS and CCV from each subject’s QSM image (@fig-masks).
 The vascular masks were used to calculate the mean susceptibility of each subject’s SSS and CCV from their QSM image with `fslstats`.
@@
 To isolate the paramagnetic component of subjects’ QSM data, the $\chi$-separation toolbox [@shinHseparationMagneticSusceptibility2021] from the Laboratory for Imaging Science and Technology was used.
-Each subject’s magnitude and phase SWI data were used along with the following parameters: 0.9766 x 0.9766 x 1 mm^3^ voxel size; TE (s) = \[0.005, 0.0102, 0.0155, 0.0207, 0.026\]; delta TE (s) = 0.0052; B0 strength = 3; B0 direction = \[0, 0, 1\].
+All five echoes of each subject’s magnitude and phase SWI data were used along with the following parameters: 0.9766 x 0.9766 x 1 mm^3^ voxel size; TE (ms) = \[5, 10.24, 15.48, 20.72, 26.96\]; $\Delta$TE (ms) = 5.24; B0 strength = 3; B0 direction = \[0, 0, 1\].
+The $\chi$-separation toolbox outputs a single negative (diamagnetic), positive (paramagnetic), and total $\chi$ map (QSM).
 The mean susceptibility of each subject’s SSS and CCV in their paramagnetic maps was calculated with the same vascular masks used for the QSM images.
 Sample images showing the magnitude, final QSM, and final paramagnetic component images are shown in @fig-sample.
-
-{{< embed notebooks/Figures.ipynb#fig-sample >}}
 
@@
 if (knitr::is_latex_output()) {
+  colnames(df) <- c("\\textbf{Variable}", "\\textbf{Subject data (n=19)}")
 df |>
@@
     fmt_markdown(columns = TRUE) |>
+    cols_label(
+      Variable = md("**Variable**"),
+      "Subject data (n = 19)" = md("**Subject data (n = 19)**")
+    )|>
     tab_options(quarto.disable_processing = TRUE)
@@
 if (knitr::is_latex_output()) {
-df <- df |> rename("Paramagnetic map" = pmap, "p-value" = pvalue, "95\\\U0025 CI" = CI)
+df <- df |> rename("\\textbf{Region}" = Region, "\\textbf{Measure}" = Measure, "\\textbf{QSM}" = QSM, "\\textbf{Paramagnetic map}" = pmap, "\\textbf{p-value}" = pvalue, "\\textbf{95\\\U0025 CI}" = CI)
 df |>
@@
     fmt_markdown(columns = TRUE) |>
+    cols_label(
+      Region = md("**Region**"),
+      Measure = md("**Measure**"),
+      QSM = md("**QSM**"),
+      "Paramagnetic map" = md("**Paramagnetic map**"),
+      "p-value" = md("**p-value**"),
+      "95\U0025 CI" = md("**95\U0025 CI**")
+    )|>
     tab_options(quarto.disable_processing = TRUE)
@@
 Region-specific $\chi$ and SvO~2~ values acquired from QSM were compared to values acquired from paramagnetic maps.
-In both the SSS and CCV, it was found that a significant difference existed between values acquired ($\chi$ and SvO~2~) from QSM and paramagnetic maps (p < 0.05).
-A boxplot showing the comparisons made is shown in @fig-methodplot.
-
-{{< embed notebooks/gavin_thesis_markdown.qmd#fig-methodplot >}}
+In both the SSS and CCV, it was found that a significant difference existed between values acquired ($\chi$ and SvO~2~) from QSM and paramagnetic maps (p \< 0.05).
+Violin plots of the comparisons are shown in @fig-methodplot.
 
 The acquired $\chi$ and SvO~2~ values were additionally compared between veins.
-In data created from QSM, a significant difference was found between the CCV and SSS in mean $\chi$ (p < 0.05; 95% CI [`r round(RegionQSMChi.t$conf.low,3)`, `r round(RegionQSMChi.t$conf.high,3)`]) and mean SvO~2~ (p < 0.05; 95% CI [`r round(RegionQSMSvO2.t$conf.low,3)`, `r round(RegionQSMSvO2.t$conf.high,3)`]).
-In data acquired from paramagnetic maps, no significant difference was observed between the CCV and the SSS in either mean $\chi$ (p = `r round(RegionChisepChi.t$p.value,3)`; 95% CI [`r round(RegionChisepChi.t$conf.low,3)`, `r round(RegionChisepChi.t$conf.high,3)`]) or mean SvO~2~ (p = `r round(RegionChisepSvO2.t$p.value,3)`; 95% CI [`r round(RegionChisepSvO2.t$conf.low,3)`, `r round(RegionChisepChi.t$conf.high,3)`]).
+In data created from QSM, a significant difference was found between the CCV and SSS in mean $\chi$ (p \< 0.05; 95% CI \[`r round(RegionQSMChi.t$conf.low,3)`, `r round(RegionQSMChi.t$conf.high,3)`\]) and mean SvO~2~ (p \< 0.05; 95% CI \[`r round(RegionQSMSvO2.t$conf.low,3)`, `r round(RegionQSMSvO2.t$conf.high,3)`\]).
+In data acquired from paramagnetic maps, no significant difference was observed between the CCV and the SSS in either mean $\chi$ (p = `r round(RegionChisepChi.t$p.value,3)`; 95% CI \[`r round(RegionChisepChi.t$conf.low,3)`, `r round(RegionChisepChi.t$conf.high,3)`\]) or mean SvO~2~ (p = `r round(RegionChisepSvO2.t$p.value,3)`; 95% CI \[`r round(RegionChisepSvO2.t$conf.low,3)`, `r round(RegionChisepChi.t$conf.high,3)`\]).
 A summary of these comparisons is represented in @fig-regionplot.
-
-{{< embed notebooks/gavin_thesis_markdown.qmd#fig-regionplot >}}
 
@@
 df[5,] <- c("SSS        ", "Jiang et al. \n(2019)   ", "Healthy neonates \n(n=4)                     ", "MRI: \naTRUPC  ", "64.8 \U00B1 2.9  ")
-df[6,] <- c("SSS        ", "De Vis et al. \n(2014)  ", "PT-TEA \neonates (n=18)                      ", "MRI: \nT2-TRIR ", "52.0 \U00B1 12.0 ")
+df[6,] <- c("SSS        ", "De Vis et al. \n(2014)  ", "PT-TEA \nneonates (n=18)                      ", "MRI: \nT2-TRIR ", "52.0 \U00B1 12.0 ")
 df[7,] <- c("SSS", "Yadav et al. \n(2020)", "Late third \ntrimester fetuses \n(n=33)", "MRI: \nSusceptometry", "58.6 \U00B1 4.8")
-df[8,] <- c("SSS", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \nQSM", paste0(rndmean(newdata$Gavin_SSSVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Gavin_SSSVein_CSvO2*100)))
-df[9,] <- c("SSS", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \n$\\chi$-separation", paste0(rndmean(newdata$Chisep_SSSVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Chisep_SSSVein_CSvO2*100)))
+df[8,] <- c("SSS", "**This study**", "PT-TEA \nneonates \nn=19", "MRI: \nQSM", paste0(rndmean(newdata$Gavin_SSSVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Gavin_SSSVein_CSvO2*100)))
+df[9,] <- c("SSS", "**This study**", "PT-TEA \nneonates \nn=19", "MRI: \n \U03C7-separation", paste0(rndmean(newdata$Chisep_SSSVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Chisep_SSSVein_CSvO2*100)))
 df[10,] <- c("CCV", "Weber et al. (2021)", "Preterm neonates \nwith HIE (n=8)", "MRI: \nQSM", "72.2 \U00B1 6.0")
@@
 df[12,] <- c("CCV", "Jiang et al. (2019)", "Healthy neonates \n(n=4)", "MRI: \naTRUPC", "70.2 \U00B1 3.3")
-df[13,] <- c("CCV", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \nQSM", paste0(rndmean(newdata$Gavin_IntVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Gavin_IntVein_CSvO2)))
-df[14,] <- c("CCV", "\\textbf{This study}", "PT-TEA \nneonates \nn=19", "MRI: \n$\\chi$-separation", paste0(rndmean(newdata$Chisep_IntVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Chisep_IntVein_CSvO2*100)))
+df[13,] <- c("CCV", "**This study**", "PT-TEA \nneonates \nn=19", "MRI: \nQSM", paste0(rndmean(newdata$Gavin_IntVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Gavin_IntVein_CSvO2)))
+df[14,] <- c("CCV", "**This study**", "PT-TEA \nneonates \nn=19", "MRI: \n \U03C7-separation", paste0(rndmean(newdata$Chisep_IntVein_CSvO2*100), " \U00B1 ", rndsd(newdata$Chisep_IntVein_CSvO2*100)))
 footnotetext="PT-TEA = born preterm and scanned at term-equivalent age; late third trimester = >35 weeks gestational age; HIE = hypoxic-ischemic encephalopathy; TRUST = T2-relaxation-under-spin tagging; aTRUPC = accelerated T2-relaxation-under-phase-contrast; T2-TRIR = T2-tissue-relaxation-inversion-recovery; SWI = susceptibility weighted imaging."
@@
 if (knitr::is_latex_output()) {
-df <- df |> rename("SvO$_{2}$ (\\\U0025)" = SvO2)
+df <- df |> rename("\\textbf{SvO$_{2}$ (\\\U0025)}" = SvO2, "\\textbf{Region}" = Region, "\\textbf{Study}" = Study, "\\textbf{Subjects}" = Subjects, "\\textbf{Method}" = Method)
+df[8,2] <- "\\textbf{This study}"
+df[9,2] <- "\\textbf{This study}"
+df[13,2] <- "\\textbf{This study}"
+df[14,2] <- "\\textbf{This study}"
 df[1,] <- linebreak(df[1,])
@@
 } else {
-df <- df |> rename("SvO~2~ (\\\U0025)" = SvO2)
 df |> 
@@
     fmt_markdown(columns = TRUE) |>
+    cols_label(
+      Region = md("**Region**"),
+      Study = md("**Study**"),
+      Subjects = md("**Subjects**"),
+      Method = md("**Method**"),
+      SvO2 = md("**SvO~2~ (%)**")
+    )|>
     tab_options(quarto.disable_processing = TRUE)
@@
 ```
+
 As shown in @tbl-litvalues, the infants observed in MRI studies investigating cerebral vein oxygenation noticeably differ in clinical status, with three studies involving healthy neonates [@weberQuantitativeSusceptibilityMapping2021; @gouAutomaticRejectionBased2024; @jiangVesselspecificQuantificationNeonatal2019], three studies (including the present study) involving preterm neonates [@weberQuantitativeSusceptibilityMapping2021; @devisNoninvasiveMRIMeasurements2014], and one study involving late third trimester fetuses [@portnoyHumanUmbilicalCord2018].
@@
 
-This study aimed to evaluate how the use of susceptibility separation on preterm neonatal QSM images in order to  provide greater accuracy in determining the oxygenation of cerebral venous vessels.
+This study aimed to evaluate how the use of susceptibility separation on preterm neonatal QSM images could be used in determining the oxygenation of cerebral venous vessels.
 We compared venous specific SvO~2~ values obtained from QSM images and their respective paramagnetic components to SvO~2~ values from neonatal MRI studies.
@@
 
+{{< pagebreak >}}
+
+## Data availability
+
+The manuscript was written in a 'reproducible manner'.
+The entire manuscript, including statistics reported, figures, and tables, can be reproduced here: [https://github.com/WeberLab/Chisep_CSVO2_Manuscript](https://github.com/WeberLab/Chisep_CSVO2_Manuscript)
+
+Unfortunately, we can not upload our MRI images to an open repository as we did not obtain permission in our consent forms.
+
 # References {.unnumbered}
@@
 :::
+
+## Acknowledgments
+
+We wish to acknowledge the work of and thank Victoria Tapics (Research Nurse); Vicki Goh (Research Nurse); Chacko Anil (Neonatologist); and Michael A. Sargent (Radiologist).
+
+## Author contributions
+
+TGC wrote the original draft, performed the formal analysis, and contributed to methodology, validation, and visualization.
+AR helped with writing, reviewing & editing.
+REG helped with writing, reviewing & editing, and with initial funding acquisition.
+AMW was involved in project administration, supervision, validation, visualization, resources, methodology, formal analysis, funding acquisition, writing, reviewing & editing, conceptualization, data curation, and investigation.
+
+## Funding
+
+Authors AMW and REG were co-primary applicants for a BC Children’s Hospital Research Institute - Brain, Behaviour and Development Catalyst Grant (\$20,000).
+AMW was supported by an establishment award from BCCHRI.
+Scanning was partly funded through a special award to AMW from BCCHRI .
+
+## Competing interests
+
+The authors have no competing interests to declare.
+
+## Consent statement
+
+The study was approved by the Clinical Research Ethics Board at the University of British Columbia and Children's & Women's Hospital (H21-00655) and written informed consent was obtained from the parent/guardian for each infant.
+
+{{< pagebreak >}}
+
+{{< embed notebooks/Figures.ipynb#fig-graph >}}
+
+{{< embed notebooks/Figures.ipynb#fig-masks >}}
+
+{{< embed notebooks/Figures.ipynb#fig-sample >}}
+
+{{< embed notebooks/gavin_thesis_markdown.qmd#fig-methodplot >}}
+
+{{< embed notebooks/gavin_thesis_markdown.qmd#fig-regionplot >}}

### 2025-02-15 notebooks/gavin_thesis_markdown.qmd
final

@@
 ```
-
+```{r}
+ggsave("Figure4.png", plot = patch1, dpi = 300, width = 8, height = 5, units = "in")
+```
 # Region analysis: Chi and CSvO2
@@
 ```
+```{r}
+ggsave("Figure5.png", plot = patch2, dpi = 300, width = 8, height = 5, units = "in")
+```
 

### 2025-02-16 _tex/index.tex
Built site for gh-pages

@@
 Vancouver, BC, Canada}}
+\affil[4]{, \orgname{Physics and Astronomy, The University of British
+Columbia, Vancouver, BC, Canada}}
 
@@
 
-\abstract{\textbf{Background}: Quantitative susceptibility mapping
-(QSM), an magnetic resonance imaging (MRI) modality sensitive to
-deoxyhemoglobin, is a promising method for measuring cerebral
-oxygenation in human neonates. Paramagnetic sources, like
-deoxyhemoglobin, however, can be obscured by diamagnetic sources such as
-water and myelin. This study evaluated whether QSM images, or isolated
-paramagnetic components, are more accurate for measuring oxygenation of
-cerebral veins of preterm neonates, and explored oxygenation differences
-between the major cerebral veins.
+% keywords
+\keywords{Quantitative Susceptbility
+Mapping,  Preterm,  Newborn,  Cerebral Venous Oxygen Saturation}
+
+\begin{document}
+\maketitle
+
+\textsuperscript{1} Integrated Sciences, The University of British
+Columbia, Vancouver, BC, Canada\\
+\textsuperscript{2} BC Children's Hospital Research Institute, The
+University of British Columbia, Vancouver, BC, Canada\\
+\textsuperscript{3} Pediatrics, The University of British Columbia,
+Vancouver, BC, Canada\\
+\textsuperscript{4} Physics and Astronomy, The University of British
+Columbia, Vancouver, BC, Canada
+
+\textsuperscript{*} Correspondence:
+\href{mailto:aweber@bcchr.ca}{Alexander Mark Weber
+\textless{}aweber@bcchr.ca\textgreater{}}
+
+\subsection{Impact Statement}\label{impact-statement}
+
+\begin{itemize}
+\tightlist
+\item
+  This study evaluated the use of QSM and its paramagnetic components to
+  measure cerebral oxygenation in neonates.
+\item
+  By comparing susceptibility-derived oxygen saturation
+  (SvO\textsubscript{2}) in the superior sagittal sinus (SSS) and
+  central cerebral veins (CCV), it adds to the field of neonatal
+  cerebral oxygenation measurement.
+\item
+  Decomposing QSM into paramagnetic components shows potential for
+  improving SvO\textsubscript{2} accuracy, particularly in the SSS,
+  though variability remains a challenge.
+\item
+  The results suggest no significant oxygenation difference between the
+  SSS and CCV, contrasting with previous findings, indicating a need for
+  further research on neonatal venous oxygenation.
+\end{itemize}
+
+\textbf{Category of Study:} basic science
+
+\newpage{}
+
+\subsection{Abstract:}\label{abstract}
+
+\textbf{Background}: Quantitative susceptibility mapping (QSM), a
+magnetic resonance imaging (MRI) modality sensitive to deoxyhemoglobin,
+is a promising method for measuring cerebral oxygenation in human
+neonates. Paramagnetic sources, like deoxyhemoglobin, however, can be
+obscured by diamagnetic sources such as water and myelin. This study
+evaluated whether QSM images, or isolated paramagnetic components, are
+more accurate for measuring oxygenation of cerebral veins of preterm
+neonates, and explored oxygenation differences between the major
+cerebral veins.
 
@@
 \textbf{Conclusion}: While paramagnetic component decomposition yielded
-more accurate SSS values, it increased variability. No significant
-oxygenation differences were found between the SSS and CCV, contrasting
-with prior studies.}
-
-% keywords
-\keywords{Quantitative Susceptbility
-Mapping,  Preterm,  Newborn,  Cerebral Venous Oxygen Saturation}
-
-\begin{document}
-\maketitle
-
-\textsuperscript{1} Integrated Sciences, The University of British
-Columbia, Vancouver, BC, Canada\\
-\textsuperscript{2} BC Children's Hospital Research Institute, The
-University of British Columbia, Vancouver, BC, Canada\\
-\textsuperscript{3} Pediatrics, The University of British Columbia,
-Vancouver, BC, Canada
-
-\textsuperscript{*} Correspondence:
-\href{mailto:aweber@bcchr.ca}{Alexander Mark Weber
-\textless{}aweber@bcchr.ca\textgreater{}}
-
-\textbf{Category of Study:} basic science
-
-\subsection{Impact Statement}\label{impact-statement}
-
-\begin{itemize}
-\tightlist
-\item
-  This study evaluated the accuracy of using QSM and its paramagnetic
-  components to measure cerebral oxygenation in neonates.
-\item
-  By comparing susceptibility-derived oxygen saturation
-  (SvO\textsubscript{2}) in the superior sagittal sinus (SSS) and
-  central cerebral veins (CCV), it adds to the field of neonatal
-  cerebral oxygenation measurement.
-\item
-  Decomposing QSM into paramagnetic components shows potential for
-  improving SvO\textsubscript{2} accuracy, particularly in the SSS,
-  though variability remains a challenge.
-\item
-  The results suggest no significant oxygenation difference between the
-  SSS and CCV, contrasting with previous findings, indicating a need for
-  further research on neonatal venous oxygenation.
-\end{itemize}
-
-\subsection{Acknowledgments}\label{acknowledgments}
-
-We wish to acknowledge the work of and thank Victoria Tapics (Research
-Nurse); Vicki Goh (Research Nurse); Chacko Anil (Neonatologist); and
-Michael A. Sargent (Radiologist).
-
-\subsection{Funding}\label{funding}
-
-Authors AMW and REG were co-primary applicants for a BC Children's
-Hospital Research Institute - Brain, Behaviour and Development Catalyst
-Grant (\$20,000). AMW was supported by an establishment award from
-BCCHRI. Scanning was partly funded through a special award to AMW from
-BCCHRI .
-
-\subsection{Author contributions}\label{author-contributions}
-
-TGC wrote the original draft, performed the formal analysis, and
-contributed to methodology, validation, and visualization. AR helped
-with writing, reviewing \& editing. REG helped with writing, reviewing
-\& editing, and with initial funding acquisition. AMW was involved in
-project administration, supervision, validation, visualization,
-resources, methodology, formal analysis, funding acquisition, writing,
-reviewing \& editing, conceptualization, data curation, and
-investigation.
-
-\subsection{Competing interests}\label{competing-interests}
-
-The authors have no competing interests to declare.
-
-\subsection{Consent statement}\label{consent-statement}
-
-The study was approved by the Clinical Research Ethics Board at the
-University of British Columbia and Children's \& Women's Hospital
-(H21-00655) and written informed consent was obtained from the
-parent/guardian for each infant.
+SSS values closer to those found in the literature, it increased
+variability. No significant oxygenation differences were found between
+the SSS and CCV, contrasting with prior studies.
 
@@
 actively being explored to assess regional and whole-brain blood
-oxygenation. MRI based methods have been developed for adults
-\citep{jainRapidMagneticResonance2011, luQuantitativeEvaluationOxygenation2008, xuNoninvasiveQuantificationWholebrain2009}
-and are only beginning to be explored in neonates
+oxygenation. While MRI-based methods have been developed for adults
+\citep{jainRapidMagneticResonance2011, luQuantitativeEvaluationOxygenation2008, xuNoninvasiveQuantificationWholebrain2009},
+their application in neonates is only beginning to be explored
 \citep{devisNoninvasiveMRIMeasurements2014, liuQuantitativeAssessmentGlobal2014, qiHemodynamicMetabolicAssessment2018, jainCerebralOxygenMetabolism2014, jiangVesselspecificQuantificationNeonatal2019}.
-These methods almost all rely on T2 relaxation to estimate
-CSvO\textsubscript{2}
+This delay is likely due to the unique challenges posed by neonates,
+including their smaller anatomies, distinct hemodynamic profiles,
+susceptibility to motion artifacts, and the difficulties associated with
+recruiting this population for research. These methods have almost all
+relied on T2 relaxation to estimate CSvO\textsubscript{2}
 \citep{devisNoninvasiveMRIMeasurements2014, liuQuantitativeAssessmentGlobal2014, qiHemodynamicMetabolicAssessment2018, jiangVesselspecificQuantificationNeonatal2019}
@@
 arbitrary threshold value of 0.15 ppm in order to acquire realistic
-results \citep{weberQuantitativeSusceptibilityMapping2021}.
+results \citep{weberQuantitativeSusceptibilityMapping2021}. Furthermore,
+QSM tends to underestimates parametric components due to the inclusion
+of diamagnetic tissue, and vice versa, as the opposing magnetic
+susceptibilities effectively subtract from one another
+\citep{kimHSeparationImagingDiagnosis2023}.
 
@@
 scanner (scanner software version DV26.0\_R03) equipped with a SREE
-Medical Systems (Cleveland, OH, USA) single-channel neonatal head coil
+Medical Systems (Cleveland, OH) single-channel neonatal head coil
 (Table~\ref{tbl-mri}). The scans were conducted at the *** 's MRI
@@
 \toprule
- & T1w & T2w & pcASL & SWI\\
+ & \textbf{T1w} & \textbf{T2w} & \textbf{pcASL} & \textbf{SWI}\\
 \midrule
 Sequence & 3D FSPGR & 3D CUBE & Multi-shot 3D fast spin-echo & 3D spoiled GRE flow-compensated\\
-Phase-encoding direction & Coronal & Sagittal & Axial & Axial\\
+Acquisition plane & Coronal & Sagittal & Axial & Axial\\
+Phase-encoding direction & Left-Right & Posterior-Anterior & Posterior-Anterior & Left-Right\\
 TR (ms) & 7.74 & 2,300 & 4,680 & 30.9\\
@@
 
-The MRI scan protocol comprised of the following sequences: a
-T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin
-labeling (ASL) scan \citep{alsopRecommendedImplementationArterial2015},
-a multi-echo susceptibility-weighted imaging scan
-\citep{denkSusceptibilityWeightedImaging2010}, and a diffusion-weighted
-imaging (DWI) spin-echo echo planar imaging (EPI) sequence. The DWI
-sequence was not used for the present study.
+The MRI scan protocol comprised of the following sequences (plane of
+acquisition in parentheses): a T1-weighted scan (coronal), a T2-weighted
+scan (sagittal), a pseudo-continuous arterial spin labeling (ASL) scan
+\citep{alsopRecommendedImplementationArterial2015} (axial), a multi-echo
+susceptibility-weighted imaging scan
+\citep{denkSusceptibilityWeightedImaging2010} (axial), and a
+diffusion-weighted imaging (DWI) spin-echo echo planar imaging (EPI)
+sequence (axial). The DWI sequence was not used for the present study.
 
@@
 
-\begin{figure}[H]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-graph-output-1.png}}
-
-}
-
-\caption{\label{fig-graph}\textbf{Pipeline for generating
-subject-specific brain masks that include the superior sagittal sinus
-(SSS).} Initial steps involved (1) creating brain a mask from the
-magnitude of the fifth echo of the susceptibility weighted scan.
-Subsequently, the brain mask is dilated and then (2) utilized in
-conjunction with a quantitative susceptibility mapping (QSM) script to
-generate a preliminary QSM image. Further refinement involved (3)
-segmenting the SSS from the QSM image manually to create a tissue mask
-of the SSS region. Finally, (4) the vascular mask of the SSS is
-integrated with the initial brain mask, forming the comprehensive brain
-mask essential for obtaining susceptibility data that includes the SSS.}
-
-\end{figure}%
-
 First, the fifth echo (TE = 25.96 ms) SWI magnitude file was processed
@@
 to create a preliminary brain mask, similar to our previous efforts,
-which does not contain the SSS. \texttt{Fslroi} was used to isolate the
-fifth echo of the magnitude data, which was then squared using
-\texttt{fslmaths} and the option \texttt{-sqr}. Squaring the magnitude
-image was found to dramatically improve subsequent brain extraction. The
-resulting image was then used to create the preliminary brain mask using
-bet with the options \texttt{-m} and \texttt{-R}. The former flag
-generated a binary brain mask, while the latter performed a more robust
-brain centre estimation. The brain mask was then dilated by 7 voxels
-using \texttt{fslmaths} and the options \texttt{-kernel\ boxv} and
-\texttt{-dilM} in order for the dilated mask to contain the SSS (along
-with unwanted tissue as well). This mask was then used, along with the
-phase images, in a MATLAB script for QSM calculation from Christian
-Kames \citep{kamesRapidTwostepDipole2018} to produce a preliminary QSM
-image that contained the SSS, albeit with fairly low signal-to-noise
-ratio and other unwanted tissue. Given the high contrast in voxel
-intensity between the SSS and surrounding tissue, the select by
-intensity tool in \texttt{FSLeyes} \citep{mccarthyFSLeyes2023} was then
-used to segment the SSS from the QSM image and create a 3D mask of the
-selected region. Using \texttt{fslmaths} and the options \texttt{-add}
-and \texttt{-bin}, the SSS mask was then combined with the original
-brain mask of the fifth echo. This resulted in a brain mask that
-contained only brain and SSS signal. Finally, this mask was used in a
-final QSM post-processing step to create a QSM image that includes the
-SSS while maintaining a high signal-to-noise ratio, making it suitable
-to obtain accurate susceptibility values.
+which does not contain the SSS. The last echo is used to generate the
+brain mask as it reliably removes artifacts from air-tissue and
+bone-tissue interfaces, e.g., sinuses, without the need for manual
+erosion. The reason for this is at longer echo times, tissues with rapid
+signal decay (such as bone, air, and sinuses) lose their MRI signal due
+to dephasing caused by greater magnetic field inhomogeneities.
+\texttt{Fslroi} was used to isolate the fifth echo of the magnitude
+data, which was then squared using \texttt{fslmaths} and the option
+\texttt{-sqr}. Squaring the magnitude image was found to dramatically
+improve subsequent brain extraction. The resulting image was then used
+to create the preliminary brain mask using \texttt{bet} with the options
+\texttt{-m} and \texttt{-R}. The former flag generated a binary brain
+mask, while the latter performed a more robust brain centre estimation.
+The brain mask was then dilated by 7 voxels using \texttt{fslmaths} and
+the options \texttt{-kernel\ boxv} and \texttt{-dilM} in order for the
+dilated mask to contain the SSS (along with unwanted tissue as well).
+This mask was then used, along with the phase images, in a MATLAB script
+for QSM calculation from Christian Kames
+\citep{kamesRapidTwostepDipole2018} to produce a preliminary QSM image
+that contained the SSS, albeit with fairly low signal-to-noise ratio and
+other unwanted tissue. Given the high contrast in voxel intensity
+between the SSS and surrounding tissue, the select by intensity tool in
+\texttt{FSLeyes} \citep{mccarthyFSLeyes2023} was then used to segment
+the SSS from the QSM image and create a 3D mask of the selected region.
+Using \texttt{fslmaths} and the options \texttt{-add} and \texttt{-bin},
+the SSS mask was then combined with the original brain mask of the fifth
+echo. This resulted in a brain mask that contained only brain and SSS
+signal. Finally, this mask was used in a final QSM post-processing step
+to create a QSM image that includes the SSS while maintaining a high
+signal-to-noise ratio, making it suitable to obtain accurate
+susceptibility values.
 
 STI Suite (v. 3.0) \citep{liIntegratedLaplacianbasedPhase2014}, was used
-to process the final QSM images as it produced the cleanest images
-without eroding the SSS. The finalized brain mask and the last three
+to process the final QSM images as it produced the images with the least
+amount of artifacts (based on a visual assessment by the authors)
+without eroding the SSS. The finalized brain mask along with all five
 echoes of the magnitude and phase images were used in STI Suite along
 with the following parameters: 0.9766 x 0.9766 x 1 mm\textsuperscript{3}
-voxel size, 5 ms TE1, 5.3 ms \(\Delta\)TE, and 77.4 ms sum TE, B0
+voxel size, 5 ms TE1, 5.24 ms \(\Delta\)TE, and 77.4 ms sum TE, B0
 strength = 3, and B0 direction = {[}0, 0, 1{]}. The 3D GRE data option
 was selected for the phase processing stage, and STAR-QSM was selected
-for the QSM stage. Finally, the `select by intensity' tool in
-\texttt{FSLeyes} was then used to semi-automatically make vascular masks
-of the SSS and CCV from each subject's QSM image. The vascular masks
-were used to calculate the mean susceptibility of each subject's SSS and
-CCV from their QSM image with \texttt{fslstats}.
+for the QSM stage. STAR-QSM outputs a single QSM map for each echo
+(i.e.~five total). The last three echoes of the QSM maps were then
+averaged to create the final QSM image using \texttt{fslmaths}, as the
+accumulation of phase due to susceptibility is small in early echoes and
+artifacts dominate the phase
+\citep{zhangQuantitativeAnalysisPunctate2019}. Finally, the `select by
+intensity' tool in \texttt{FSLeyes} was then used to semi-automatically
+make vascular masks of the SSS and CCV from each subject's QSM image
+(Figure~\ref{fig-masks}). The vascular masks were used to calculate the
+mean susceptibility of each subject's SSS and CCV from their QSM image
+with \texttt{fslstats}.
 
@@
 \citep{shinHseparationMagneticSusceptibility2021} from the Laboratory
-for Imaging Science and Technology was used. Each subject's magnitude
-and phase SWI data were used along with the following parameters: 0.9766
-x 0.9766 x 1 mm\textsuperscript{3} voxel size; TE (s) = {[}0.005,
-0.0102, 0.0155, 0.0207, 0.026{]}; delta TE (s) = 0.0052; B0 strength =
-3; B0 direction = {[}0, 0, 1{]}. The mean susceptibility of each
-subject's SSS and CCV in their paramagnetic maps was calculated with the
-same vascular masks used for the QSM images. Sample images showing the
-magnitude, final QSM, and final paramagnetic component images are shown
-in Figure~\ref{fig-sample}.
-
-\begin{figure}[H]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-sample-output-1.png}}
-
-}
-
-\caption{\label{fig-sample}\textbf{An example of subject imaging data.}
-A sample sagittal, coronal, and axial slice is displayed for each image.
-(a) The 1st echo of the magnitude susceptibility weighted imaging
-sequence; (b) the final quantitative susceptibility mapping image; and
-(c) the paramagnetic component isolated from the quantitative
-susceptibility map. The bars in (b) and (c) indicates the range of
-susceptibility \(\chi\) values.}
-
-\end{figure}%
+for Imaging Science and Technology was used. All five echoes of each
+subject's magnitude and phase SWI data were used along with the
+following parameters: 0.9766 x 0.9766 x 1 mm\textsuperscript{3} voxel
+size; TE (ms) = {[}5, 10.24, 15.48, 20.72, 26.96{]}; \(\Delta\)TE (ms) =
+5.24; B0 strength = 3; B0 direction = {[}0, 0, 1{]}. The
+\(\chi\)-separation toolbox outputs a single negative (diamagnetic),
+positive (paramagnetic), and total \(\chi\) map (QSM). The mean
+susceptibility of each subject's SSS and CCV in their paramagnetic maps
+was calculated with the same vascular masks used for the QSM images.
+Sample images showing the magnitude, final QSM, and final paramagnetic
+component images are shown in Figure~\ref{fig-sample}.
 
@@
 \toprule
-Variable & Subject data (n = 19)\\
+\textbf{Variable} & \textbf{Subject data (n=19)}\\
 \midrule
@@
 \toprule
-Region & Measure & QSM & Paramagnetic map & p-value & 95\% CI\\
+\textbf{Region} & \textbf{Measure} & \textbf{QSM} & \textbf{Paramagnetic map} & \textbf{p-value} & \textbf{95\% CI}\\
 \midrule
@@
 values acquired (\(\chi\) and SvO\textsubscript{2}) from QSM and
-paramagnetic maps (p \textless{} 0.05). A boxplot showing the
-comparisons made is shown in Figure~\ref{fig-methodplot}.
-
-\begin{figure}[H]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-methodplot-output-1.png}}
-
-}
-
-\caption{\label{fig-methodplot}\textbf{Vein-specific susceptibility and
-oxygen saturation values by method.} (A, B) contains violin plots
-comparing subject chi (ppm) acquired from the cerebral central veins;
-(C, D) contains violin plots comparing subject SvO2 (\%) acquired from
-the superior sagittal sinus. Raw data points from paramagnetic maps are
-shown as filled green circles and raw data points from QSM are shown as
-filled orange circles. Each line connects the raw data points of a
-single subject. (***) indicates P\textless0.05.}
-
-\end{figure}%
+paramagnetic maps (p \textless{} 0.05). Violin plots of the comparisons
+are shown in Figure~\ref{fig-methodplot}.
 
@@
 represented in Figure~\ref{fig-regionplot}.
-
-\begin{figure}[H]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-regionplot-output-1.png}}
-
-}
-
-\caption{\label{fig-regionplot}\textbf{Inter-venous comparisons of
-susceptibility and oxygen saturation}. Violin plots comparing (A, C) chi
-(ppm) and (B, D) SvO2 (\%) between the CCV and the SSS. Panels (A) and
-(B) used data acquired from QSM, and its raw data points are shown as
-filled orange circles. Panels (C) and (D) used data acquired from
-paramagnetic maps, and its raw data points are shown as filled green
-circles. Each line connects the raw data points of a single subject.
-(***) indicates p\textless0.05; (n.s.) indicates no significant
-difference.}
-
-\end{figure}%
 
@@
 \toprule
-Region & Study & Subjects & Method & SvO$_{2}$ (\%)\\
+\textbf{Region} & \textbf{Study} & \textbf{Subjects} & \textbf{Method} & \textbf{SvO$_{2}$ (\%)}\\
 \midrule
@@
 SSS & \makecell[l]{Jiang et al. \\(2019)   } & \makecell[l]{Healthy neonates \\(n=4)                     } & \makecell[l]{MRI: \\aTRUPC  } & 64.8 ± 2.9\\
-SSS & \makecell[l]{De Vis et al. \\(2014)  } & \makecell[l]{PT-TEA \\eonates (n=18)                      } & \makecell[l]{MRI: \\T2-TRIR } & 52.0 ± 12.0\\
+SSS & \makecell[l]{De Vis et al. \\(2014)  } & \makecell[l]{PT-TEA \\neonates (n=18)                      } & \makecell[l]{MRI: \\T2-TRIR } & 52.0 ± 12.0\\
 SSS & \makecell[l]{Yadav et al. \\(2020)} & \makecell[l]{Late third \\trimester fetuses \\(n=33)} & \makecell[l]{MRI: \\Susceptometry} & 58.6 ± 4.8\\
 SSS & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\QSM} & 72.46 ± 3.43\\
-SSS & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\$\chi$-separation} & 58.14 ± 7.3\\
+SSS & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\ χ-separation} & 58.14 ± 7.3\\
 CCV & Weber et al. (2021) & \makecell[l]{Preterm neonates \\with HIE (n=8)} & \makecell[l]{MRI: \\QSM} & 72.2 ± 6.0\\
@@
 CCV & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\QSM} & 68.71 ± 0.03\\
-CCV & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\$\chi$-separation} & 57.69 ± 6.97\\
+CCV & \textbf{This study} & \makecell[l]{PT-TEA \\neonates \\n=19} & \makecell[l]{MRI: \\ χ-separation} & 57.69 ± 6.97\\
 \bottomrule
@@
 This study aimed to evaluate how the use of susceptibility separation on
-preterm neonatal QSM images in order to provide greater accuracy in
-determining the oxygenation of cerebral venous vessels. We compared
-venous specific SvO\textsubscript{2} values obtained from QSM images and
-their respective paramagnetic components to SvO\textsubscript{2} values
-from neonatal MRI studies. We found that susceptibility separation
-provided SvO\textsubscript{2} values of the SSS that were comparable to
-values found in the literature, providing evidence that this processing
+preterm neonatal QSM images could be used in determining the oxygenation
+of cerebral venous vessels. We compared venous specific
+SvO\textsubscript{2} values obtained from QSM images and their
+respective paramagnetic components to SvO\textsubscript{2} values from
+neonatal MRI studies. We found that susceptibility separation provided
+SvO\textsubscript{2} values of the SSS that were comparable to values
+found in the literature, providing evidence that this processing
 technique may be a valid tool for measuring regional cerebral oxygen
@@
 
+\newpage{}
+
+\subsection{Data availability}\label{data-availability}
+
+The manuscript was written in a `reproducible manner'. The entire
+manuscript, including statistics reported, figures, and tables, can be
+reproduced here:
+\url{https://github.com/WeberLab/Chisep_CSVO2_Manuscript}
+
+Unfortunately, we can not upload our MRI images to an open repository as
+we did not obtain permission in our consent forms.
+
 \section*{References}\label{references}
@@
 
+\subsection{Acknowledgments}\label{acknowledgments}
+
+We wish to acknowledge the work of and thank Victoria Tapics (Research
+Nurse); Vicki Goh (Research Nurse); Chacko Anil (Neonatologist); and
+Michael A. Sargent (Radiologist).
+
+\subsection{Author contributions}\label{author-contributions}
+
+TGC wrote the original draft, performed the formal analysis, and
+contributed to methodology, validation, and visualization. AR helped
+with writing, reviewing \& editing. REG helped with writing, reviewing
+\& editing, and with initial funding acquisition. AMW was involved in
+project administration, supervision, validation, visualization,
+resources, methodology, formal analysis, funding acquisition, writing,
+reviewing \& editing, conceptualization, data curation, and
+investigation.
+
+\subsection{Funding}\label{funding}
+
+Authors AMW and REG were co-primary applicants for a BC Children's
+Hospital Research Institute - Brain, Behaviour and Development Catalyst
+Grant (\$20,000). AMW was supported by an establishment award from
+BCCHRI. Scanning was partly funded through a special award to AMW from
+BCCHRI .
+
+\subsection{Competing interests}\label{competing-interests}
+
+The authors have no competing interests to declare.
+
+\subsection{Consent statement}\label{consent-statement}
+
+The study was approved by the Clinical Research Ethics Board at the
+University of British Columbia and Children's \& Women's Hospital
+(H21-00655) and written informed consent was obtained from the
+parent/guardian for each infant.
+
+\newpage{}
+
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-graph-output-1.png}}
+
+}
+
+\caption{\label{fig-graph}\textbf{Pipeline for generating
+subject-specific brain masks that include the superior sagittal sinus
+(SSS).} Initial steps involved (1) creating brain a mask from the
+magnitude of the fifth echo of the susceptibility weighted scan.
+Subsequently, the brain mask is dilated and then (2) utilized in
+conjunction with a quantitative susceptibility mapping (QSM) script to
+generate a preliminary QSM image. Further refinement involved (3)
+segmenting the SSS from the QSM image manually to create a tissue mask
+of the SSS region. Finally, (4) the vascular mask of the SSS is
+integrated with the initial brain mask, forming the comprehensive brain
+mask essential for obtaining susceptibility data that includes the SSS.}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-masks-output-1.png}}
+
+}
+
+\caption{\label{fig-masks}\textbf{Sample venous masks.} A sample
+superior sagittal sinus (red) and central cerebral vein mask (blue)
+displayed in coronal, sagittal, and axial view. The QSM image is used as
+the underlay. y, x and z values represent the slice number in each plane
+(coronal, sagittal, and axial, respectively).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-sample-output-1.png}}
+
+}
+
+\caption{\label{fig-sample}\textbf{An example of subject imaging data.}
+A sample coronal, sagittal, and axial slice is displayed for each image.
+(a) The 1st echo of the magnitude susceptibility weighted imaging
+sequence; (b) the final quantitative susceptibility mapping image; and
+(c) the paramagnetic component isolated from the quantitative
+susceptibility map. The color bars in (b) and (c) indicates the range of
+susceptibility \(\chi\) values in parts per million. y, x and z values
+represent the slice number in each plane (coronal, sagittal, and axial,
+respectively).}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-methodplot-output-1.png}}
+
+}
+
+\caption{\label{fig-methodplot}\textbf{Vein-specific susceptibility and
+oxygen saturation values by method.} (A, B) contains violin plots
+comparing subject chi (ppm) acquired from the cerebral central veins;
+(C, D) contains violin plots comparing subject SvO2 (\%) acquired from
+the superior sagittal sinus. Raw data points from paramagnetic maps are
+shown as filled green circles and raw data points from QSM are shown as
+filled orange circles. Each line connects the raw data points of a
+single subject. (***) indicates P\textless0.05.}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-regionplot-output-1.png}}
+
+}
+
+\caption{\label{fig-regionplot}\textbf{Inter-venous comparisons of
+susceptibility and oxygen saturation}. Violin plots comparing (A, C) chi
+(ppm) and (B, D) SvO2 (\%) between the CCV and the SSS. Panels (A) and
+(B) used data acquired from QSM, and its raw data points are shown as
+filled orange circles. Panels (C) and (D) used data acquired from
+paramagnetic maps, and its raw data points are shown as filled green
+circles. Each line connects the raw data points of a single subject.
+(***) indicates p\textless0.05; (n.s.) indicates no significant
+difference.}
+
+\end{figure}%
+
 \end{document}

### 2025-02-19 index.qmd
published

@@
 citation:
-  container-title: Unpublished
+  container-title: Pediatric Research
+  volume: 97
+  issue: 2
+  doi:
 # date: last-modified

### 2025-02-19 index.qmd
doi

--- 2025-02-15 index.qmd
@@
 citation:
-  container-title: Unpublished
+  container-title: Pediatric Research
+  volume: 97
+  issue: 2
+  doi: "not yet assigned"
 # date: last-modified

### 2025-02-21 index.qmd
added back notebook links, cleaned up _quarto.yml, and added figures back to the main doc

--- 2025-02-15 index.qmd
@@
 
-notebook-links: false
 number-sections: false
 citation:
-  container-title: Unpublished
+  container-title: Pediatric Research
+  volume: 97
+  issue: 2
+  doi: "not yet assigned"
 # date: last-modified
@@
 
+{{< embed notebooks/Figures.ipynb#fig-graph >}}
+
 First, the fifth echo (TE = 25.96 ms) SWI magnitude file was processed using FSL’s (v.
@@
 
+{{< embed notebooks/Figures.ipynb#fig-masks >}}
+
 To isolate the paramagnetic component of subjects’ QSM data, the $\chi$-separation toolbox [@shinHseparationMagneticSusceptibility2021] from the Laboratory for Imaging Science and Technology was used.
@@
 Sample images showing the magnitude, final QSM, and final paramagnetic component images are shown in @fig-sample.
+
+{{< embed notebooks/Figures.ipynb#fig-sample >}}
 
@@
 
+{{< embed notebooks/gavin_thesis_markdown.qmd#fig-methodplot >}}
+
 The acquired $\chi$ and SvO~2~ values were additionally compared between veins.
@@
 A summary of these comparisons is represented in @fig-regionplot.
+
+{{< embed notebooks/gavin_thesis_markdown.qmd#fig-regionplot >}}
 
@@
 The manuscript was written in a 'reproducible manner'.
-The entire manuscript, including statistics reported, figures, and tables, can be reproduced here: [https://github.com/WeberLab/Chisep_CSVO2_Manuscript](https://github.com/WeberLab/Chisep_CSVO2_Manuscript)
+The entire manuscript, including statistics reported, figures, and tables, can be reproduced here: [weberlab.github.io/Chisep_CSVO2_Manuscript](https://weberlab.github.io/Chisep_CSVO2_Manuscript)
 
@@
 The study was approved by the Clinical Research Ethics Board at the University of British Columbia and Children's & Women's Hospital (H21-00655) and written informed consent was obtained from the parent/guardian for each infant.
-
-{{< pagebreak >}}
-
-{{< embed notebooks/Figures.ipynb#fig-graph >}}
-
-{{< embed notebooks/Figures.ipynb#fig-masks >}}
-
-{{< embed notebooks/Figures.ipynb#fig-sample >}}
-
-{{< embed notebooks/gavin_thesis_markdown.qmd#fig-methodplot >}}
-
-{{< embed notebooks/gavin_thesis_markdown.qmd#fig-regionplot >}}

### 2025-02-21 _tex/index.tex
Built site for gh-pages

@@
 
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
+Notebook}}
+
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
+Notebook}}
+
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
+Notebook}}
+
 \newpage{}
@@
 
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
+Notebook}}
+
 The MRI scan protocol comprised of the following sequences (plane of
@@
 Figure~\ref{fig-graph}.
+
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-graph-output-1.png}}
+
+}
+
+\caption{\label{fig-graph}\textbf{Pipeline for generating
+subject-specific brain masks that include the superior sagittal sinus
+(SSS).} Initial steps involved (1) creating brain a mask from the
+magnitude of the fifth echo of the susceptibility weighted scan.
+Subsequently, the brain mask is dilated and then (2) utilized in
+conjunction with a quantitative susceptibility mapping (QSM) script to
+generate a preliminary QSM image. Further refinement involved (3)
+segmenting the SSS from the QSM image manually to create a tissue mask
+of the SSS region. Finally, (4) the vascular mask of the SSS is
+integrated with the initial brain mask, forming the comprehensive brain
+mask essential for obtaining susceptibility data that includes the SSS.}
+
+\end{figure}%
+
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures-preview.html\#cell-fig-graph}{Python
+Figures}}
 
@@
 
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-masks-output-1.png}}
+
+}
+
+\caption{\label{fig-masks}\textbf{Sample venous masks.} A sample
+superior sagittal sinus (red) and central cerebral vein mask (blue)
+displayed in coronal, sagittal, and axial view. The QSM image is used as
+the underlay. y, x and z values represent the slice number in each plane
+(coronal, sagittal, and axial, respectively).}
+
+\end{figure}%
+
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures-preview.html\#cell-fig-masks}{Python
+Figures}}
+
 To isolate the paramagnetic component of subjects' QSM data, the
@@
 
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-sample-output-1.png}}
+
+}
+
+\caption{\label{fig-sample}\textbf{An example of subject imaging data.}
+A sample coronal, sagittal, and axial slice is displayed for each image.
+(a) The 1st echo of the magnitude susceptibility weighted imaging
+sequence; (b) the final quantitative susceptibility mapping image; and
+(c) the paramagnetic component isolated from the quantitative
+susceptibility map. The color bars in (b) and (c) indicates the range of
+susceptibility \(\chi\) values in parts per million. y, x and z values
+represent the slice number in each plane (coronal, sagittal, and axial,
+respectively).}
+
+\end{figure}%
+
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures-preview.html\#cell-fig-sample}{Python
+Figures}}
+
 Once the mean susceptibility values of the SSS and CCV were obtained
@@
 are shown in Figure~\ref{fig-methodplot}.
+
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-methodplot-output-1.png}}
+
+}
+
+\caption{\label{fig-methodplot}\textbf{Vein-specific susceptibility and
+oxygen saturation values by method.} (A, B) contains violin plots
+comparing subject chi (ppm) acquired from the cerebral central veins;
+(C, D) contains violin plots comparing subject SvO2 (\%) acquired from
+the superior sagittal sinus. Raw data points from paramagnetic maps are
+shown as filled green circles and raw data points from QSM are shown as
+filled orange circles. Each line connects the raw data points of a
+single subject. (***) indicates P\textless0.05.}
+
+\end{figure}%
+
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/gavin_thesis_markdown-preview.html\#cell-fig-methodplot}{gavin\_thesis\_R}}
 
@@
 represented in Figure~\ref{fig-regionplot}.
+
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-regionplot-output-1.png}}
+
+}
+
+\caption{\label{fig-regionplot}\textbf{Inter-venous comparisons of
+susceptibility and oxygen saturation}. Violin plots comparing (A, C) chi
+(ppm) and (B, D) SvO2 (\%) between the CCV and the SSS. Panels (A) and
+(B) used data acquired from QSM, and its raw data points are shown as
+filled orange circles. Panels (C) and (D) used data acquired from
+paramagnetic maps, and its raw data points are shown as filled green
+circles. Each line connects the raw data points of a single subject.
+(***) indicates p\textless0.05; (n.s.) indicates no significant
+difference.}
+
+\end{figure}%
+
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/gavin_thesis_markdown-preview.html\#cell-fig-regionplot}{gavin\_thesis\_R}}
 
@@
 \endgroup{}
+
+\textsubscript{Source:
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
+Notebook}}
 
@@
 reproduced here:
-\url{https://github.com/WeberLab/Chisep_CSVO2_Manuscript}
+\href{https://weberlab.github.io/Chisep_CSVO2_Manuscript}{weberlab.github.io/Chisep\_CSVO2\_Manuscript}
 
@@
 
-\newpage{}
-
-\begin{figure}[H]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-graph-output-1.png}}
-
-}
-
-\caption{\label{fig-graph}\textbf{Pipeline for generating
-subject-specific brain masks that include the superior sagittal sinus
-(SSS).} Initial steps involved (1) creating brain a mask from the
-magnitude of the fifth echo of the susceptibility weighted scan.
-Subsequently, the brain mask is dilated and then (2) utilized in
-conjunction with a quantitative susceptibility mapping (QSM) script to
-generate a preliminary QSM image. Further refinement involved (3)
-segmenting the SSS from the QSM image manually to create a tissue mask
-of the SSS region. Finally, (4) the vascular mask of the SSS is
-integrated with the initial brain mask, forming the comprehensive brain
-mask essential for obtaining susceptibility data that includes the SSS.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-masks-output-1.png}}
-
-}
-
-\caption{\label{fig-masks}\textbf{Sample venous masks.} A sample
-superior sagittal sinus (red) and central cerebral vein mask (blue)
-displayed in coronal, sagittal, and axial view. The QSM image is used as
-the underlay. y, x and z values represent the slice number in each plane
-(coronal, sagittal, and axial, respectively).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-sample-output-1.png}}
-
-}
-
-\caption{\label{fig-sample}\textbf{An example of subject imaging data.}
-A sample coronal, sagittal, and axial slice is displayed for each image.
-(a) The 1st echo of the magnitude susceptibility weighted imaging
-sequence; (b) the final quantitative susceptibility mapping image; and
-(c) the paramagnetic component isolated from the quantitative
-susceptibility map. The color bars in (b) and (c) indicates the range of
-susceptibility \(\chi\) values in parts per million. y, x and z values
-represent the slice number in each plane (coronal, sagittal, and axial,
-respectively).}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-methodplot-output-1.png}}
-
-}
-
-\caption{\label{fig-methodplot}\textbf{Vein-specific susceptibility and
-oxygen saturation values by method.} (A, B) contains violin plots
-comparing subject chi (ppm) acquired from the cerebral central veins;
-(C, D) contains violin plots comparing subject SvO2 (\%) acquired from
-the superior sagittal sinus. Raw data points from paramagnetic maps are
-shown as filled green circles and raw data points from QSM are shown as
-filled orange circles. Each line connects the raw data points of a
-single subject. (***) indicates P\textless0.05.}
-
-\end{figure}%
-
-\begin{figure}[H]
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-regionplot-output-1.png}}
-
-}
-
-\caption{\label{fig-regionplot}\textbf{Inter-venous comparisons of
-susceptibility and oxygen saturation}. Violin plots comparing (A, C) chi
-(ppm) and (B, D) SvO2 (\%) between the CCV and the SSS. Panels (A) and
-(B) used data acquired from QSM, and its raw data points are shown as
-filled orange circles. Panels (C) and (D) used data acquired from
-paramagnetic maps, and its raw data points are shown as filled green
-circles. Each line connects the raw data points of a single subject.
-(***) indicates p\textless0.05; (n.s.) indicates no significant
-difference.}
-
-\end{figure}%
-
 \end{document}

### 2025-02-21 index.qmd
added github and binder link

@@
   doi: "not yet assigned"
-# date: last-modified
+date: last-modified
 bibliography: Gavin_Thesis_Ref.bib

### 2025-02-21 index.qmd
removed anonymizing information

@@
 
-Participant data comes from a previous study \*\*\* <!-- [@zhu-etal-cmro2] (submitted) -->.
-Participants consisted of preterm neonates born between 25- and 31-weeks gestational age (GA) who were admitted to the level III NICU at \*\*\* <!-- BC Women’s Hospital -->.
+Participant data comes from a previous study [@zhu-etal-cmro2] (under review).
+Participants consisted of preterm neonates born between 25- and 31-weeks gestational age (GA) who were admitted to the level III NICU at BC Women’s Hospital.
 Recruitment took place over a span of one year, from February 2021 to January 2022, facilitated by a dedicated research nurse.
@@
 MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750 scanner (scanner software version DV26.0_R03) equipped with a SREE Medical Systems (Cleveland, OH) single-channel neonatal head coil (@tbl-mri).
-The scans were conducted at the \*\*\* <!-- BC Children -->’s MRI Research Facility.
+The scans were conducted at the BC Children’s MRI Research Facility.
 Prior to the scanning procedure, subjects were carefully prepared by a research nurse.
@@
 The raw DICOM files acquired from the scanning procedure were converted to NIfTI (Neuroimaging Informatics Technology Initiative) format using Chris Rorden’s `dcmniix` tool [@liFirstStepNeuroimaging2016].
-SWI magnitude data files were then used to create subject-specific brain masks that would not erode the SSS during QSM processing, an issue faced by our group in the past \*\*\* <!-- [@weberQuantitativeSusceptibilityMapping2021] -->.
+SWI magnitude data files were then used to create subject-specific brain masks that would not erode the SSS during QSM processing, an issue faced by our group in the past [@weberQuantitativeSusceptibilityMapping2021].
 A step-by-step summary of the pipeline used is shown in @fig-graph.

### 2025-02-21 _tex/index.tex
Built site for gh-pages

@@
 
-Participant data comes from a previous study *** . Participants
-consisted of preterm neonates born between 25- and 31-weeks gestational
-age (GA) who were admitted to the level III NICU at *** . Recruitment
-took place over a span of one year, from February 2021 to January 2022,
-facilitated by a dedicated research nurse. Parents of eligible infants
-were approached by the research nurse prior to discharge from the NICU
-to explain the study objectives and seek their consent for
-participation. Infants meeting the criteria for inclusion were scanned
-for the study if they had already been discharged from the NICU, were in
-stable condition, and had reached a term equivalent age of 37 to 44
-weeks GA. However, certain exclusion criteria were applied to ensure the
-homogeneity and integrity of the study sample: infants were excluded if
-there was clinical evidence of a congenital malformation or syndrome, a
-TORCH infection, or ultrasound evidence of large parenchymal hemorrhagic
-infarction (\textgreater2 cm, Grade 4 intraventricular hemorrhage).
+Participant data comes from a previous study \citep{zhu-etal-cmro2}
+(under review). Participants consisted of preterm neonates born between
+25- and 31-weeks gestational age (GA) who were admitted to the level III
+NICU at BC Women's Hospital. Recruitment took place over a span of one
+year, from February 2021 to January 2022, facilitated by a dedicated
+research nurse. Parents of eligible infants were approached by the
+research nurse prior to discharge from the NICU to explain the study
+objectives and seek their consent for participation. Infants meeting the
+criteria for inclusion were scanned for the study if they had already
+been discharged from the NICU, were in stable condition, and had reached
+a term equivalent age of 37 to 44 weeks GA. However, certain exclusion
+criteria were applied to ensure the homogeneity and integrity of the
+study sample: infants were excluded if there was clinical evidence of a
+congenital malformation or syndrome, a TORCH infection, or ultrasound
+evidence of large parenchymal hemorrhagic infarction (\textgreater2 cm,
+Grade 4 intraventricular hemorrhage).
 
@@
 Medical Systems (Cleveland, OH) single-channel neonatal head coil
-(Table~\ref{tbl-mri}). The scans were conducted at the *** 's MRI
+(Table~\ref{tbl-mri}). The scans were conducted at the BC Children's MRI
 Research Facility. Prior to the scanning procedure, subjects were
@@
 used to create subject-specific brain masks that would not erode the SSS
-during QSM processing, an issue faced by our group in the past *** . A
-step-by-step summary of the pipeline used is shown in
-Figure~\ref{fig-graph}.
+during QSM processing, an issue faced by our group in the past
+\citep{weberQuantitativeSusceptibilityMapping2021}. A step-by-step
+summary of the pipeline used is shown in Figure~\ref{fig-graph}.
 

### 2025-03-31 index.qmd
added citation metadata

@@
   container-title: Pediatric Research
-  volume: 97
-  issue: 2
-  doi: "not yet assigned"
+  issued: 2025
+  doi: 10.1038/s41390-025-03966-6
 date: last-modified

### 2025-03-31 _tex/index.tex
Built site for gh-pages

@@
 \makeatletter
-\newsavebox\pandoc@box
-\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
-  \sbox\pandoc@box{#1}%
-  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
-  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
-  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
-  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
-  \else\usebox{\pandoc@box}%
-  \fi%
-}
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 % Set default figure placement to htbp
+\makeatletter
 \def\fps@figure{htbp}
@@
 \makeatother
-
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
 \usepackage[]{natbib}
@@
 \textsubscript{Source:
-\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index.qmd.html}{Article
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/index-preview.html}{Article
 Notebook}}
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-graph-output-1.png}}
+\includegraphics{index_files/figure-latex/notebooks-Figures-fig-graph-output-1.png}
 
@@
 \textsubscript{Source:
-\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures-preview.html\#cell-fig-graph}{Python
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures.ipynb.html\#cell-fig-graph}{Python
 Figures}}
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-masks-output-1.png}}
+\includegraphics{index_files/figure-latex/notebooks-Figures-fig-masks-output-1.png}
 
@@
 \textsubscript{Source:
-\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures-preview.html\#cell-fig-masks}{Python
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures.ipynb.html\#cell-fig-masks}{Python
 Figures}}
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-Figures-fig-sample-output-1.png}}
+\includegraphics{index_files/figure-latex/notebooks-Figures-fig-sample-output-1.png}
 
@@
 \textsubscript{Source:
-\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures-preview.html\#cell-fig-sample}{Python
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/Figures.ipynb.html\#cell-fig-sample}{Python
 Figures}}
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-methodplot-output-1.png}}
+\includegraphics{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-methodplot-output-1.png}
 
@@
 \textsubscript{Source:
-\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/gavin_thesis_markdown-preview.html\#cell-fig-methodplot}{gavin\_thesis\_R}}
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/gavin_thesis_markdown.qmd.html\#cell-fig-methodplot}{gavin\_thesis\_R}}
 
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-regionplot-output-1.png}}
+\includegraphics{index_files/figure-latex/notebooks-gavin_thesis_markdown-fig-regionplot-output-1.png}
 
@@
 \textsubscript{Source:
-\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/gavin_thesis_markdown-preview.html\#cell-fig-regionplot}{gavin\_thesis\_R}}
+\href{https://WeberLab.github.io/Chisep_CSVO2_Manuscript/notebooks/gavin_thesis_markdown.qmd.html\#cell-fig-regionplot}{gavin\_thesis\_R}}
 
